### 2023-11-06 tex/state-learner-pub-version.tex
new repo

\documentclass[a4,danish]{article}
\usepackage[a4paper, margin=3.5cm]{geometry}
\input{./latex-settings/standard-settings.tex}
\input{./latex-settings/standard-commands.tex}
\input{./latex-settings/theorem-env.tex}
\input{./latex-settings/title-compact.tex}

\fxsetup{status=draft}
\usepackage[colorlinks=true,citecolor=blue]{hyperref} % Doesn't work for org, so included here

%% Specific libraries
\usepackage{booktabs}

%% Specific commands
\newcommand{\data}{\ensuremath{\mathcal{D}}}

\begin{document}

% \articletype{ARTICLE TEMPLATE}% Specify the article type or omit as appropriate

\title{The state learner \\ {-- a super learner for right-censored data}}

\author{Anders Munch and Thomas A. Gerds
  % \name{Anders Munch\textsuperscript{a}\thanks{CONTACT: Anders Munch. Email: a.munch@sund.ku.dk}
  %   and Thomas A. Gerds\textsuperscript{a}}
  % \affil{\textsuperscript{a} Section of Biostatistics, Department of Public Health, University of Copenhagen}
}

\maketitle

\begin{abstract}
  In survival analysis, prediction models are needed as stand-alone tools and in
  applications of causal inference to estimate nuisance parameters. The super
  learner is a machine learning algorithm which combines a library of prediction
  models into a meta learner based on cross-validated loss. Unfortunately, the
  commonly used partial likelihood loss is not suited for super learning, and
  inverse probability of censoring weighted loss functions require a
  pre-specified estimator of the censoring distribution. To relax this, we
  introduce the state learner, a new super learner for survival analysis, which
  evaluates the loss based on the observed data simultaneously using libraries
  of predictions models for the event(s) of interest and the censoring
  distribution. We establish an oracle inequality for the state learner and
  investigate its performance through numerical experiments. We illustrate how
  the state learner allows us to estimate causal effects in a competing risks
  setting without having to pre-specify models for neither the cause-specific
  hazards of interest nor the censoring distribution.
  % In survival analysis, prediction models are needed as stand-alone tools and in
  % applications of causal inference to estimate nuisance parameters. The super
  % learner is a machine learning algorithm which combines a library of prediction
  % models into a meta learner based on cross-validated loss. Unfortunately, the
  % commonly used partial likelihood loss is not suited for super learning, and
  % inverse probability of censoring weighted loss functions require a
  % pre-specified estimator of the censoring distribution. To relax this, we
  % introduce the state learner, a new super learner for survival analysis, which
  % evaluates the loss based on the observed data simultaneously for two libraries
  % of prediction models, one for the event time distribution and one for the
  % censoring distribution. We establish an oracle inequality for the state
  % learner and investigate its performance through numerical experiments.
\end{abstract}

% \begin{keywords}
%   Loss based estimation; right-censored data; competing risks; super learner;
%   cross-validation
% \end{keywords}

% \begin{amscode}
%   % Use 2010 Mathematics Subject Classification keywords
%   62N02
% \end{amscode}

\section{Introduction}
\label{sec:introduction}

A super learner is a machine learning algorithm that combines a finite
set of learners into a meta learner by estimating prediction
performance in hold-out samples using a pre-specified loss function
\citep{van2007super}. When the aim is to make a prediction model,
super learners typically combine strong learners, such as Cox
regression models and random survival forests
\citep{gerds2021medical}.
% For targeted learning of a low dimensional
% target parameter in presence of a high-dimensional nuisance parameter,
% a super learner will usually also include the highly adaptive lasso
% \citep{van2011targeted, benkeser2016highly,van2017generally}.
While the general idea of combining strong learners based on cross-validation
data stems from earlier work \citep{wolpert1992stacked,breiman1996stacked}, the
name super learner is justified by an oracle inequality
\citep{van2003unicv,vaart2006oracle}: The super learner is guaranteed to perform
almost as well as the model which minimizes the expected performance, i.e., the
model we would select if we could evaluate the prediction performance in an
infinite hold-out sample.

We are concerned with the choice of the loss function for super learning in
survival analysis. Existing super learner algorithms for right-censored data use
partial log-likelihood loss or inverse probability of censoring weighted loss
\citep{polley2011-sl-cens,keles2004asymptotically,golmakani2020super,westling2021inference}.
The use of the partial log-likelihood loss restricts the class of learners and
excludes for example simple Kaplan-Meier based learners and also more complex
random survival forest algorithms. For this reason \cite{golmakani2020super}
restrict their learners to Cox proportional hazard models. A lesser known fact
is that a super learner constructed with the negative partial log-likelihood
loss implicitly depends on the censoring distribution
\citep{hjort1992inference,whitney2019comment}. A disadvantage of inverse
probability of censoring weighted loss functions is that they requires a
pre-specified model for the censoring distribution. \cite{westling2021inference}
tackle this challenge by iterating between super learning of the censoring
distribution and the event time distribution.

In this article we define the state learner, a new super learner for
right-censored data, which simultaneously evaluates the loss for learners of the
event time distribution and the censoring distribution. The loss function which
is used to define the state learner is only based on observable
quantities. The state learner can be applied to all types of survival
estimators, works in the presence of competing risks, and does not require a
single pre-specified estimator of the conditional censoring distribution. To
analyze the theoretical properties of the state learner we focus on the
so-called discrete super learner which `combines' the library of
learners by picking the one that minimizes the cross-validated loss. The state
learner uses separate libraries to model each competing event and the censoring
distribution. We show that the oracle selector of the state learner is
consistent if all libraries contain a consistent learner and prove a finite
sample oracle inequality.

The state learner can be used to select a model which predicts the probability
of an event based on covariates in the presence of competing risks. Another
application is in targeted learning where conditional event probabilities occur
as high-dimensional nuisance parameters which need to be estimated at a certain
rate \citep{van2011targeted, rytgaard2021estimation, rytgaard2022targeted}. We
show how a targeted estimator can be obtained from the state learner, and that a
second order product structure for the asymptotic bias term of the targeted
estimator is retained when the state learner is used to estimate nuisance
parameters.

% For machine learning with right-censored data, the negative partial
% log-likelihood is a popular loss function
% \citep[e.g.,][]{li2016regularized,
%   yao2017deep,lee2018deephit,katzman2018deepsurv,
%   gensheimer2019scalable,lee2021boosted, golmakani2020super}.
% However, there are two major problems when attempting to use the partial log-likelihood for super
% learning. The first is that the partial
% log-likelihood cannot be used to evaluate cross-validated loss for
% survival learners which are not absolutely continuous in the time
% argument \citep{golmakani2020super}. The second problem is that the
% expected value of the partial log-likelihood depends on the censoring
% distribution. Both problems are further discussed in Section
% \ref{sec:relat-liter-exist}.

% Inverse probability of censoring weighted (IPCW) estimators are often
% used to resolve the dependence of the expected loss on the censoring
% distribution \citep{graf1999assessment,
%   molinaro2004tree,keles2004asymptotically,hothorn2006survival,gerds2006consistent,gonzalez2021stacked}.
% However, commonly these methods require a consistent estimator of the
% conditional censoring distribution given covariates. An obvious idea
% would be to estimate the conditional censoring distribution by a super
% learner. However, that would require a loss function to evaluate the
% performance of candidate learners of the censoring distribution and
% the expected value of that loss function would in turn depend on the
% event time distribution. To resolve this dilemma
% \cite{westling2021inference} and \cite{han2021inverse} suggested to
% iterate between super learning of the censoring distribution and the
% event time distribution, where both super learners use the respective
% cross-validated IPCW loss. Our approach is different. The state
% learner uses all combinations of learners for the event time
% distribution with a learner for the censoring distribution to
% parameterize a set of `observable' cumulative distribution
% functions. Specifically, the cross-validated integrated Brier score
% for each of the cumulative distribution functions then only depends on
% the observed data distribution and hence can be estimated empirically
% without further modeling.

The article is organized as follows. We introduce our notation and framework in
Section~\ref{sec:framework}. In Section~\ref{sec:super-learning} we define super
learning in general with right-censored data, and in
Section~\ref{sec:super-learner-simple} we introduce the state learner.
Section~\ref{sec:theor-results-prop} provides theoretical guarantees for the
state learner. In Section~\ref{sec:targeted-learning} we discuss the use of the
state learner in the context of targeted learning. We report a numerical study
in Section~\ref{sec:numer-exper}, and analyze a prostate cancer data set in
Section~\ref{sec:real-data-appl}. Finally, we relate the state learner to
existing approaches in Section~\ref{sec:discussion} and discuss some limitations
of our proposal. Appendices~\ref{sec:proof-proposition}
and~\ref{sec:state-learner-with} contain proofs.

\section{Notation and framework}
\label{sec:framework}

In a competing risk framework \citep{andersen2012statistical}, let \( T\) be a
time to event variable, \(D\in\{1,2\}\) the cause of the event, and
$X \in \mathcal{X}$ a vector of baseline covariates taking values in a bounded
subset \( \mathcal{X} \subset \R^p \), \( p\in\N \). Let $\tau< \infty$ be the
maximal length of follow-up. We use \( \mathcal{Q} \) to denote the collection
of all probability measures on \( [0,\tau] \times \{1,2\}\times \mathcal{X} \)
such that \( (T, D, X) \sim Q \) for some unknown \( Q \in \mathcal{Q} \). For
\(j\in\{1,2\}\), the cause-specific conditional cumulative hazard functions are
defined by
\( \Lambda_{j} \colon [0, \tau] \times \mathcal{X} \rightarrow \R_+ \) such that
\begin{equation*}
  % \label{eq:cum-haz}
  \Lambda_{j}(t \mid x) = \int_0^t\frac{  Q(T \in \diff s, D=j \mid X=x )}{Q(T \geq s \mid X=x )}.
\end{equation*}
For ease of notation we assume throughout that \( \Lambda_j(\blank \mid x) \) is
continuous for all \( x \) and \( j \). We denote by \(S\) the conditional
event-free survival function,
\begin{equation}
  \label{eq:surv-def}
  S(t \mid x)=\exp\left\{-\Lambda_{1}(t \mid x)-\Lambda_{2}(t \mid x)\right\}.
\end{equation}
Let \( \mathcal{M} \) denote the space of all conditional cumulative hazard
functions on \( [0,\tau] \times\mathcal{X}\). Any distribution
\( Q \in \mathcal{Q} \) can be characterized by
\begin{equation*}
  \label{eq:parametrizeQ}
  \begin{split}
    Q(\diff t,j,\diff x)&= \left\{S(t- \mid x)\Lambda_1(\diff t \mid x)H(\diff x)\right\}^{1_{\{j=1\}}}\\
                        &\quad\left\{S(t- \mid x)\Lambda_2(\diff t \mid x)H(\diff x)\right\}^{1_{\{j=2\}}},
  \end{split}
\end{equation*}
where \(\Lambda_{j} \in \mathcal{M}\) for \(j=1,2\) and \(H\) is the marginal
distribution of the covariates.

We consider the usual right-censored setting in which we observe data
\(O = (\tilde{T},\tilde D, X)\), where $\tilde T = \min(T,C)$ for a
right-censoring time \(C\), $\Delta = \1{\{T \leq C\}}$, and
\(\tilde D=\Delta D\). Let \(\mathcal{P}\) denote a set of probability measures
on the sample space
\(\mathcal{O} = [0, \tau] \times \{0, 1, 2\} \times \mathcal{X}\) such that
\(O \sim P \) for some unknown \(P\in \mathcal{P}\). We assume that the event
time and the censoring time are conditionally independent given covariates,
\( T \independent C \mid X \). This implies that any distribution
\( P \in \mathcal{P} \) is characterized by a distribution
\( Q \in \mathcal{Q} \) and a conditional cumulative hazard function for \( C \)
given \( X \) \citep[c.f.,][]{begun1983information,gill1997coarsening}. We use
\(\Gamma\in\mathcal M\) to denote the conditional cumulative hazard function for
censoring. We assume that \( \Gamma(\blank \mid x) \) is continuous for all
\( x \), and let \(G(t \mid x)=\exp\left\{-\Gamma(t \mid x)\right\}\) denote the
survival function of the conditional censoring distribution. In our setting with
competing risks, this yields
\begin{equation}\label{eq:parametrizeP}
  \begin{split}
    P(\diff t, j, \diff x) &= \left\{G(t- \mid x)S(t- \mid x)\Lambda_1(\diff t \mid x)H(\diff x)\right\}^{\1{{\{j=1\}}}}\\
                           &\quad\left\{G(t- \mid x)S(t- \mid x)\Lambda_2(\diff t \mid x)H(\diff x)\right\}^{\1{{\{j=2\}}}}\\
                           &\quad\left\{G(t- \mid x)S(t- \mid x)\Gamma(\diff t \mid x)H(\diff x)\right\}^{\1{{\{j=0\}}}}\\
                           &=\left\{G(t- \mid x)Q(\diff t,j,\diff x)\right\}^{\1{{\{j\ne 0\}}}}\\    
                           &\quad\left\{G(t- \mid x)S(t- \mid x)\Gamma(\diff t \mid x)H(\diff x)\right\}^{\1{{\{j=0\}}}}.
  \end{split}
\end{equation}
Hence, we may write
\( \mathcal{P} = \{ P_{Q, \Gamma} : Q \in \mathcal{Q}, \Gamma \in
\mathcal{G} \} \) for some \( \mathcal{G} \subset \mathcal{M} \). We
also have
\begin{equation*}
P(\tilde T>t \mid X=x) = S(t \mid x)G(t \mid x) = \exp\left\{-\Lambda_{1}(t \mid x)-\Lambda_{2}(t \mid x)-\Gamma(t \mid x) \right\}.
\end{equation*}
We further assume that there exists \(\kappa<\infty\) such that
\(\Lambda_{j}(\tau- \mid x)<\kappa \), for \(j\in\{1,2\}\), and
\(\Gamma(\tau- \mid x)<\kappa\) for almost all \(x\in\mathcal X\). Note that this
implies that \(G(\tau- \mid x)\) is bounded away from zero for almost all \(x\in\mathcal X\).
Under these assumptions, the conditional cumulative hazard functions
\(\Lambda_{j}\) and \(\Gamma\) can be identified from \(P\) by
\begin{align}
  \Lambda_{j}(t \mid x) &= \int_0^t\frac{  P(\tilde T \in \diff s, \tilde D=j \mid X=x )}{P(\tilde T \geq s \mid X=x )}, \label{eq:lambdaj}\\
  \Gamma(t \mid x) &= \int_0^t\frac{  P(\tilde T \in \diff s, \tilde D=0 \mid X=x )}{P(\tilde T \geq s \mid X=x )}\label{eq:gamma}.
\end{align}
Thus, we can consider $\Lambda_j$ and \(\Gamma\) as operators which map from
\( \mathcal{P} \) to \(\mathcal M\).
% and we shall sometimes write
% \( \Lambda^{P} \) to emphasize that \( \Gamma^{P} \) is associated with a
% particular data-generating distribution \( P \).

\section{Super learning with right-censored survival data}
\label{sec:super-learning}

A super learner estimates a parameter $\Psi$ which can be identified from the
observed data distribution \(P\in\mathcal P\). In this section, to introduce the
concept of super learning, we simply consider estimation of the function
\( \Lambda_{j}\). The parameter \(\Psi:\mathcal P\to\mathcal M\) is then
identified via equation \eqref{eq:lambdaj} by \(\Psi(P)=\Lambda_j\).

As input to the super learner we need a sample \( \data_n=\{O_i\}_{i=1}^n \) of
i.i.d.\ observations from some unknown \( P \in \mathcal{P} \) and a finite
collection of candidate learners $\mathcal{A}$. Each learner
\(a \in \mathcal{A}\) is a map
\( a \colon \mathcal{O}^n \rightarrow \mathcal{M}\) which takes a data set as
input and returns an estimate $a(\data_n) \in \mathcal{M}$ of $\Lambda_{j}$.
% Formally, the domain of $a$ depends on \(n\) but we suppress this in the notation.
In what follows, we use the short-hand notation
\(P[f] = \int f(o) P(\diff o) \). A super learner evaluates the
performance of \(a \in \mathcal{A}\) using a loss function
\(L\colon \mathcal{M} \times \mathcal{O} \rightarrow \R_+\) by
estimating the expected loss \(P[L(a(\data_n), \blank)]\) using
cross-validation. Specifically, the expected loss of $a\in\mathcal A$
is estimated by splitting the data set $\data_n$ into $K$ disjoint
approximately equally sized subsets
\(\data_n^1, \data_n^2, \dots, \data_n^K \) and then calculating the
cross-validated loss
\begin{equation*}
  % \label{eq:cv-risk-est}
  \hat{R}_n(a; L) =
  \frac{1}{K}\sum_{k=1}^{K}
  % \empmeas^k{[L {(a{ (\data_n^{-k})} , \blank) }]},
  \frac{1}{| \data_n^{k} |}\sum_{O_i \in \data_n^{k}}
  L
  {
    \left(
      a{ (\data_n^{-k})}
      , O_i
    \right)
  },
  \quad \text{with} \quad
  \data_n^{-k} = \data_n \setminus \data_n^{k}.
\end{equation*}
% where \( \empmeas^{k} \) is the empirical measure of \( \data_n^{k} \). 
The subset \(\data_n^{-k}\) is referred to as the \(k\)'th training
sample, while \(\data_n^{k}\) is referred to as the \(k\)'th test or
hold-out sample.
The discrete super learner is defined as
\begin{equation*}
\hat{a}_n = \argmin_{a\in\mathcal A}\hat{R}_n(a; L).
\end{equation*}
The final estimator of \(\Psi(P)=\Lambda_j\) is then the selected
learner applied to the full data set, i.e., \(\hat{a}_n(\data_n)\).
The oracle learner is defined as the learner that minimizes the
average loss according to the data-generating distribution \( P \),
i.e.,
\begin{equation*}
  \tilde{a}_n =
  \argmin_{a \in \mathcal{A}}
  \tilde{R}_n(a; L),
  \quad \text{with} \quad 
  \tilde{R}_n(a; L)=
  \frac{1}{K}\sum_{k=1}^{K} 
  P{
    \left[
      L
      {
        \left(
          a{ (\data_n^{-k})}
          , \blank
        \right)
      }
    \right]}
  .
\end{equation*}
Note that both the discrete super learner and the oracle learner depend on the
library of learners and on the number of folds \(K\), and that the oracle
learner is a function of the data and the unknown data-generating distribution.
These dependencies are suppressed in the notation.

\section{The state learner}
\label{sec:super-learner-simple}

The problem with most existing super learners for right-censored data is that
they depend on a pre-specified estimator of the censoring distribution. The main
idea of the state learner is to jointly use learners of \( \Lambda_1 \),
\( \Lambda_2 \), and \( \Gamma \), and the relations in
equation~(\ref{eq:parametrizeP}), to learn a feature of the observed data
distribution \( P \). The discrete state learner ranks a tuple of learners of
\( (\Lambda_1, \Lambda_2, \Gamma) \) based on how well they jointly model the
observed data. To formally introduce the state learner, we define the
multi-state process
\begin{equation*}
  \eta(t) = \1\{\tilde{T} \leq t, \tilde D=1\} + 2\,\1\{\tilde{T} \leq t, \tilde D=2\}+ 3\,\1\{\tilde{T} \leq t, \tilde D=0\},
  \quad \text{for} \quad t \in [0, \tau].
\end{equation*}
Note that at time \(t\), we observe that each observation is in one of four
mutually exclusive states (Figure \ref{fig:multi-state-process}).
\begin{figure}[h]
  \centering
  \includegraphics[width=.5\textwidth]{./figures/figure-multi-state-process.pdf}
  \caption{Illustration of the multi-state process \(\eta\) used by
    the state learner. Note that `censored' is a state, hence the
    process is always observed at any time.}
  \label{fig:multi-state-process}
\end{figure}
The conditional distribution of \( \eta(t) \) given \( X=x \) is
determined by the function
\begin{equation}
  \label{eq:F-def}
  F(t, k, x) = P(\eta(t) = k \mid X=x),
  \quad \text{for all} \quad
  t \in [0,\tau],
  k \in \{0,1,2,3\},
  x \in \mathcal{X}.
\end{equation}
The function \( F \) describes the conditional state occupation
probabilities corresponding to the observed multi-state process
\(\eta\).

We propose to construct a super learner for \( F \), i.e., the target of this
super learner is $\Psi(P) = F$ where the parameter is identified through
equation~(\ref{eq:F-def}). Because each quadruple
$(\Lambda_{1}, \Lambda_{2}, \Gamma, H)$ characterizes a \(P\in\mathcal P\) which
in turn determines \( (F, H) \), a learner for \( F \) can be constructed from
learners of \( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$ as follows:
\begin{equation}\label{eq:transition}
  \begin{split}
  F(t, 1, x)
  & = P(\tilde{T} \leq t, \Delta=1 \mid X=x)
    = \int_0^t e^{\{-\Lambda_{1}(s \mid x)-\Lambda_{2}(s \mid x) - \Gamma(s \mid x)\} }  \Lambda_{1}(\diff s \mid x),
  \\
  F(t, 2, x)
  & = P(\tilde{T} \leq t, \Delta=2 \mid X=x)
    = \int_0^t e^{\{-\Lambda_{1}(s \mid x)-\Lambda_{2}(s \mid x) - \Gamma(s \mid x)\} }  \Lambda_{2}(\diff s \mid x),
  \\
  F(t, 3, x)
  & =
    P(\tilde{T} \leq t, \Delta=0 \mid X=x)
    = \int_0^t e^{\{-\Lambda_{1}(s \mid x)-\Lambda_{2}(s \mid x) - \Gamma(s \mid x)\} }  \Gamma(\diff s \mid x),
  \\
  F(t, 0, x)
  &
    = P(\tilde{T} > t \mid X= x)
    = 1- F(t, 1, x) - F(t, 2, x)- F(t, 3, x).
  \end{split}
\end{equation}
The state learner requires three libraries of learners, \(\mathcal{A}_1\),
\( \mathcal{A}_2 \), and \( \mathcal{B} \), where \(\mathcal{A}_1\) and
\( \mathcal{A}_2\) contain learners of the conditional cause-specific cumulative
hazard functions of the event time distribution \(\Lambda_1\) and
\( \Lambda_2\), respectively, and \(\mathcal{B}\) contains learners of the
conditional cumulative hazard function of the censoring
distribution. % We further
% define \(\mathcal{H}\) as the set of all probability distributions on
% \( \mathcal{X} \) and
% \begin{equation*}
%   \mathcal{H}_n=\left\{h\colon\mathcal{O}^n\longrightarrow\mathcal{H} \midd
%     h(\data_n)=H_n = \frac 1 n \sum_{i=1}^n \delta_{X_i} \right\}
% \end{equation*}
% as the library of learners which consists of a single learner, the empirical
% distribution function.
Based on the Cartesian product of
libraries of learners for \((\Lambda_1,\Lambda_2,\Gamma)\) we construct a library
$\mathcal{F}(\mathcal{A}_1, \mathcal{A}_2, \mathcal{B})$ of learners
for \( F \):
\begin{align*}
  \mathcal{F}(\mathcal{A}_1, \mathcal{A}_2, \mathcal{B})
  &= \{ \phi_{a_1,a_2, b} : a_1 \in \mathcal{A}_1, a_2 \in \mathcal{A}_2, b \in \mathcal{B}\},
    \intertext{where in correspondance with  the relations in equation \eqref{eq:transition},} 
    \phi_{a_1,a_2, b}(\data_n)(t,1,x) &= \int_0^t e^{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  a_1(\data_n)(\diff s \mid x),\\
  \phi_{a_1,a_2, b}(\data_n)(t,2,x) &= \int_0^t e^{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  a_2(\data_n)(\diff s \mid x),\\
  \phi_{a_1,a_2, b}(\data_n)(t,3,x) &= \int_0^t e^{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  b(\data_n)(\diff s \mid x),\\
  \phi_{a_1,a_2, b}(\data_n)(t,0,x) &= 1-  \sum_{j=1}^3\phi_{a_1,a_2, b}(\data_n)(t,j,x).
\end{align*}
To evaluate how well a function \( F \) predicts the observed
multi-state process we use the integrated Brier score
\( \bar B_\tau( F,O) = \int_0^{\tau} B_t(F,O) \diff t \), where \( B_t \) is the
Brier score \citep{brier1950verification} at time \( t \in [0, \tau] \),
\begin{equation*}
  B_t(F,O) = \sum_{j=0}^{3}
  \left(
      F(t,j,X) - \1{\{\eta(t)=j\}}
  \right)^2.
\end{equation*}
As described in Section~\ref{sec:super-learning}, each learner
\( \phi_{a_1, a_2, b} \) in the library
\( \mathcal{F}(\mathcal{A}_1, \mathcal{A}_2, \mathcal{B}) \) is evaluated using
the cross-validated loss,
\begin{equation*}
  \hat{R}_{n}(\phi_{a_1,a_2,b} ; \bar{B}_{\tau}) =
  \frac{1}{K}\sum_{k=1}^{K}
  \frac{1}{| \data_n^{k} |}\sum_{O_i \in \data_n^{k}}
  \bar B_\tau
  {
    \left(
      \phi_{a_1,a_2,b}{ (\data_n^{-k})}
      , O_i
    \right)
  },
\end{equation*}
and the discrete state learner is
\begin{align*}\label{eq:discrete-state-learner}
  \hat{\phi}_n
  &=  \argmin_{(a_1,a_2,b)\in \mathcal{A}_1\times\mathcal{A}_2\times\mathcal{B}}
    \hat{R}_{n}(\phi_{a_1,a_2,b} ; \bar{B}_{\tau}).
\end{align*}
% Note that depending on the target parameter of the learning task, the
% state learner may or may not depend on the distribution of the
% covariates \(H_n\). For example, for the parameter
% \(\tilde\Psi(\Lambda_{1}, \Lambda_{2}, \Gamma, H)=\Lambda_{1}\) the
% state learner would not depend on \(H_n\) but for the parameter
% \(\tilde\Psi(\Lambda_{1}, \Lambda_{2}, \Gamma, H)=\int S(t \mid x)H(\diff
% x)\) it would.

\section{Theoretical results for the state learner}
\label{sec:theor-results-prop}

In this section we establish theoretical guarantees for the state learner.
% We let \( \E_P \) denote
% expectation under \( P \) and use \( \Theta \) to denote the
% collection of all conditional state-occupation probability functions.
Proposition~\ref{prop:stric-prop} can be derived from the fact that the
integrated Brier score (also called the continuous ranked probability score) is
a strictly proper scoring rule \citep{gneiting2007strictly}. This implies that
if we minimize the average loss of the integrated Brier score, we recover the
parameters of the data-generating distribution. In particular, the oracle of a
state learner will be consistent if the library of learners contains at least
one learner that is consistent for estimation of \( F \). Recall that the
function \(F\) implicitly depends on the data generating probability measure
\(P\in\mathcal P\) but that this was suppressed in the notation. We now make
this dependence explicit by writing \(F_0\) for the function which is obtained
by substituting a specific \(P_0\in\mathcal{P}\) for \(P\) in equation
\eqref{eq:transition}.

\begin{proposition}
  \label{prop:stric-prop}
  For \(P_0\in\mathcal{P}\) define
  \begin{equation*}
    F^* = \argmin_{F} P_0{[\bar{B}_\tau(F, \blank)]},
  % F^* = \argmin_{(\Lambda_1,\Lambda_2,\Gamma) \in \mathcal{M}\otimes\mathcal{M}\otimes\mathcal{M}} P_0{[\bar B_\tau(\phi_{(\Lambda_1,\Lambda_2,\Gamma)}, \blank)]},
\end{equation*}
where the minimum is taken over all \( F \), such that \( F \) is a conditional
state occupation probability function for some measure \( P \) as defined in equation~(\ref{eq:F-def}).
Then \( F^*(t,j,\blank) = F_0(t,j,\blank) \) \( H \)-almost surely for
any \( j\in \{0,1,2,3\} \) and almost any \( t \in [0, \tau]\).
\end{proposition}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

We establish a finite sample oracle result for the state learner.
Our Corollary~\ref{cor:oracle-prop} is in essence a special case of a
general cross-validation result by \cite{vaart2006oracle}. We assume
that we split the data into equally sized folds, and for simplicity of
presentation we take \( n \) to be such that
\( |\data_n^{-k}| = n/K \) with \( K \) fixed. We will allow the
number of learners to grow with \( n \) and write
\( \mathcal{F}_n=\mathcal{F}(\mathcal{A}_{1,n}, \mathcal{A}_{2,n},
\mathcal{B}_n)\) as short-hand notation and to emphasize the
dependence on \( n \).
% The discrete super $\hat{\phi}_n$ and the
% oracle $\tilde{\phi}_n$ are defined as in Section~\ref{sec:framework}
% but now with respect to the library \( \mathcal{F}_n \) and the loss
% \( B \).
In the following we let \( \| \blank \|_{P_0} \) denote the norm
\begin{equation}
  \label{eq:norm}
  \| F \|_{P_0} = 
  \left\{
    \sum_{j=0}^{3}\int_{\mathcal{X}} \int_0^{\tau} F(t, j, x)^2 \diff t H_0( \diff x)
  \right\}^{1/2}.
\end{equation}
\begin{corollary}
  \label{cor:oracle-prop}
  For all \(P_0\in\mathcal{P}\), \( n \in \N \), \( k \in \{1, \dots, K\} \),
  and $\delta>0$,
  \begin{align*}
    \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    & \leq (1 + 2\delta)
      \E_{P_0}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    \\
    & \quad
      + (1+ \delta) 16   K \tau
      \left(
      13 + \frac{12}{\delta}
      \right)
      \frac{\log(1 + |\mathcal{F}_n|)}{n}.
  \end{align*}
\end{corollary}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

Corollary~\ref{cor:oracle-prop} has the following asymptotic consequences.

\begin{corollary}
  \label{cor:asymp-cons}
  Assume that \( |\mathcal{F}_n| = \bigO(n^q)\), for some \( q \in \N \)
  and that there exists a sequence \( \phi_n \in \mathcal{F}_n \),
  \( n \in \N \), such that
  \( \E_{P_0}{\left[ \Vert \phi_n(\data_n^{-k}) - F_{0} \Vert_{P_0}^2 \right]} =
  \bigO(n^{-\alpha}) \), for some \( \alpha\leq 1 \).
  \begin{enumerate}[label=(\roman*), topsep=0pt]
  \item If $\alpha=1$ then
    \( \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2
      \right]} = \bigO(\log(n)n^{-1}) \).
  \item If $\alpha<1$ then
    \( \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]} =
    \bigO(n^{-\alpha}) \).
  \end{enumerate}
\end{corollary}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

In Section~\ref{sec:targeted-learning} we demonstrate how the state learner can
be used for targeted learning. A targeted estimator is typically obtained from
estimators of the nuisance parameters $\Lambda_1$, $\Lambda_2$, and $\Gamma$. By
equations~(\ref{eq:lambdaj}) and~(\ref{eq:gamma}) and the definition of \( F \),
we have
\begin{equation}
  \label{eq:7}
  \Gamma(t , x) 
  = \int_0^t  \frac{F(\diff s, 3, x )}{F(s-, 0, x )},
  \quad \text{and} \quad
  \Lambda_j(t , x) 
  = \int_0^t  \frac{F(\diff s, j, x )}{F(s-, 0, x )},
  \quad j \in \{1,2\},
\end{equation}
and thus a targeted estimator can also be obtained from an estimator of $F$
using equation~(\ref{eq:7}). The key feature of a targeted estimator is that it
is asymptotically equivalent to a sum of i.i.d.\ random variables plus a second
order remainder term \citep{bickel1993efficient,fisher2021visually}. For our
setting of competing risks, the remainder term is typically dominated by terms
of the form
\begin{equation}
  \label{eq:dr-term}
  P{\left[
      \int_0^{\tau} w_n(s, \blank)
      \hat{M}_{1,n}(s \mid  \blank)
      \hat{M}_{2,n}(\diff s \mid  \blank)
    \right]},
\end{equation}
where \( (\hat{M}_{1,n}, \hat{M}_{2,n}) \) is any of the nine combinations of
\( \hat{M}_{1,n} \in \{[\Gamma -\hat{\Gamma}_n], [\Lambda_1
-\hat{\Lambda}_{1,n}], [\Lambda_2 -\hat{\Lambda}_{2,n}]\} \) and
\( \hat{M}_{2,n} \in \{[\Gamma -\hat{\Gamma}_n], [\Lambda_1
-\hat{\Lambda}_{1,n}], [\Lambda_2 -\hat{\Lambda}_{2,n}]\} \), and \( w_n \) is
some data-dependent function with domain \([0,\tau]\times\mathcal X \)
\citep{van2003unified}. In particular, a targeted estimator will be
asymptotically linear if the `products' of the estimation errors
\( \hat{M}_{1,n} \) and \( \hat{M}_{2,n} \) in equation~(\ref{eq:dr-term}) are
\( \smallO_P{(n^{-1/2})}\). Proposition~\ref{prop:dr-structure} states that if
equation~(\ref{eq:dr-term}) holds for the a targeted estimator based on
estimator $\hat{\Lambda}_{1,n}$, $\hat{\Lambda}_{2,n}$, and $\hat{\Gamma}_{n}$,
then a similar product structure holds for a targeted estimator based on
\( \hat{F}_n \). We state the result for the special case that
\(\hat{M}_{1,n}= \Gamma-\hat{\Gamma}_n \) and
\(\hat{M}_{2,n} =\Lambda_1-\hat{\Lambda}_{1,n} \), but similar results holds for
any combinations of \( \Gamma-\hat{\Gamma}_n\),
\( \Lambda_1-\hat{\Lambda}_{1,n} \), and \( \Lambda_2-\hat{\Lambda}_{2,n} \).
\begin{proposition}
  \label{prop:dr-structure}
  Assume that \( w(s,x)\leq c \), \( F(s, 0, x) \geq 1/c \) and
  \( \hat{F}_n(s, 0, x) \geq 1/c \) for some \( c>0 \) for all
  \( s \in [0, \tau] \) and \( x \in \mathcal{X} \). Then there are real-valued
  uniformly bounded functions \( w^a_n \), \( w^b_n \), \( w^c_n \), and
  \( w^d_n \) with domain \( [0,\tau]^2 \times \mathcal{X} \) such that
  \begin{align*}
    & P_0{\left[
      \int_0^{\tau} w(s, \blank)
      \left\{
      \Gamma_0(s,\blank) -\hat{\Gamma}_n(s,\blank)
      \right\}
      [\Lambda_0-\hat{\Lambda}_n]
      (\diff s, \blank)
      \right]}
    \\
    & =
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^a_n(s,u,\blank) [F_0 - \hat{F}_n](u-, 0, \blank)[F_0 - \hat{F}_n](s-, 0, \blank) F_0(\diff u, 2, \blank ) F_0 ( \diff s, 1, \blank)
      \right]}
    \\
    & \quad +
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^b_n(s,u,\blank) [F_0 - \hat{F}_n](u-, 0, \blank)
      F_0(\diff u, 2, \blank ) [F_0 - \hat{F}_n](\diff s, 1, \blank)
      \right]}
    \\
    & \quad +
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^c_n(s,u,\blank) [F_0 - \hat{F}_n](\diff u, 2, \blank)
      [F_0 - \hat{F}_n](s-, 0, \blank)
      F_0(\diff s, 1, \blank ) 
      \right]}
    \\
    & \quad +
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^d_n(s,u,\blank) [F_0 - \hat{F}_n](\diff u, 2, \blank)
      [F_0 - \hat{F}_n](\diff s, 1, \blank)
      \right]}.
  \end{align*}
\end{proposition}
\begin{proof}
  See Appendix~\ref{sec:state-learner-with}.
\end{proof}

\section{Targeted learning}
\label{sec:targeted-learning}

Features of the observed data distribution \( P \in \mathcal{P} \) are rarely of
interest. We are instead interested in a parameter
\( \theta \colon \mathcal{Q} \rightarrow \Theta \) that expresses a property of
the uncensored population governed by the measure \( Q \in \mathcal{Q} \). The
parameter space $\Theta$ can be a subset of \(\R^d\) or a subset of a function
space, such as \(\mathcal{M}\). In subsection~\ref{sec:cause-spec-aver} we give
a concrete example from causal inference where $\theta$ is the average treatment
effect and \( \Theta = [-1,1] \). Under the assumption of conditional
independent censoring and positivity, $\theta$ is identifiable from
\( \mathcal{P} \) which means that there exists an operator
\( \Psi \colon \mathcal{P} \rightarrow \Theta \) such that
\( \theta(Q) = \Psi(P_{Q, \Gamma}) \) for all $\Gamma \in \mathcal{M}$. By
equation~(\ref{eq:parametrizeP}) this imply that we may write
\begin{equation*}
  \theta(Q) = \Psi(P) = \tilde{\Psi}^0(\Lambda_1, \Lambda_2, H),
\end{equation*}
for some operator \( \tilde{\Psi}^0 \). The state learner provides a ranking of
tuples
\( (a_1, a_2, b) \in \mathcal{A}_1 \times \mathcal{A}_2 \times \mathcal{B} \),
and we use \( \hat{a}_{1,n} \), \( \hat{a}_{2,n} \), and \( \hat{b}_n \) to
denote the learners corresponding to the discrete state learner
\( \hat{\phi}_n \). Letting \( H(\data_n) \) denote the empirical measure of
\( \{X_1, \dots, X_n\} \), we can obtain a simple plug-in an estimator of
$\theta$ as
\begin{equation}
  \label{eq:2}
  \hat{\Psi}^0(\data_n) =
  \tilde{\Psi}^0(\hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n), H(\data_n)). 
\end{equation}

The asymptotic distribution of \( \hat{\Psi}_{\tau, j}^0 \) is difficult to
analyze due to the model selection step involved when estimating the nuisance
parameters $\Lambda_1$ and $\Lambda_2$. In addition, the estimator will
typically have an asymptotic bias that vanishes at a sub-optimal rate. Using
tools from semi-parametric efficiency theory, it is possible to construct a
so-called targeted or debiased estimator with smaller asymptotic bias and an
asymptotic distribution which we know how to estimate
\citep{bickel1993efficient,van2011targeted,chernozhukov2018double}. A targeted
estimator is based on the efficient influence function for the parameter
$\tilde{\Psi}^0$ and relies on an estimator of $\Gamma$ in addition to
estimators of \( \Lambda_1 \) and $\Lambda_2$. The efficient influence function
is a \( P \)-zero mean and square integrable function indexed by the nuisance
parameters \( (\Lambda_1, \Lambda_2, \Gamma) \), which we denote by
\( \psi(\blank ; \Lambda_1, \Lambda_2, \Gamma) \). The name is justified because
any regular asymptotically linear estimator that has \( \psi \) as its influence
function is efficient, meaning that it has smallest asymptotic variance among
all regular asymptotically linear estimators.

An example of a targeted estimator is the \fxnote*{also TMLE, depending on
  implementation}{one-step estimator}, defined as
\begin{equation}
    \label{eq:one-step-def}
    \hat{\Psi}_{\text{OS}}(\data_n)
    =
    \tilde{\Psi}^0(\hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
    H(\data_n))
    + \empmeas{[\psi(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
      \hat{b}_n(\data_n) )]},
\end{equation}
where \( \empmeas \) is the empirical measure of a sample \(\{O_i\}_{i=1}^n\).
We can make the following asymptotic expansion of the one-step estimator
\citep{pfanzagl1985contributions,fisher2021visually,kennedy2022semiparametric},
\begin{equation*}
  \hat{\Psi}_{\text{OS}}(\data_n)- \Psi(P)
  =  \empmeas{[\psi(\blank ; \Lambda_1, \Lambda_2, \Gamma)]}
  +\mathrm{Rem}{(\hat{\Lambda}_{1,n},\hat{\Lambda}_{2,n},  \hat{\Gamma}_n, P)} + \smallO_{P}(n^{-1/2}),
\end{equation*}
where the remainder term is typically such that
\citep{van2003unified,rytgaard2022continuous}
\begin{equation}
  \label{eq:4}
  \mathrm{Rem}{(\hat{\Lambda}_{1,n},\hat{\Lambda}_{2,n},  \hat{\Gamma}_n, P)}
  = \mathcal{O}_P{
    \left\{
      (\Lambda_1-\hat{\Lambda}_{1,n})^2
      +
      (\Lambda_2-\hat{\Lambda}_{2,n})^2
      +
      (\Gamma-\hat{\Gamma}_{n})^2
    \right\}
  }.
\end{equation}
Hence, under a suitable Donsker class regularity condition
\citep{bickel1993efficient,kennedy2016semiparametric}, when
equation~(\ref{eq:4}) holds and the nuisance parameters $\Lambda_1$,
$\Lambda_2$, and $\Gamma$ are consistently estimated at rate
\( \smallO_P{(n^{-1/4})} \) then
\begin{equation}
  \label{eq:3}
  \sqrt{n}(\hat{\Psi}_{\text{OS}}(\data_n)- \Psi(P)) \weakly \mathcal{N}(0,
  P{[\psi(\blank; \Lambda_1, \Lambda_2, \Gamma)^2]}).
\end{equation}
In particular, equation~(\ref{eq:3}) and Slutsky's lemma imply that we can
obtain asymptotically valid \((1-\alpha)\cdot100\%\) confidence intervals by
calculating
\begin{equation*}
  \left[
    \hat{\Psi}_{\text{OS}}(\data_n) - q_{\alpha/2} \hat{\sigma}(\data_n) ,
    \;
    \hat{\Psi}_{\text{OS}}(\data_n) + q_{\alpha/2} \hat{\sigma}(\data_n)
  \right],
\end{equation*}
where \( q_{\alpha} \) is the \( (1-\alpha) \)-quantile or the standard normal
distribution, and
\begin{equation*}
  \hat{\sigma}(\data_n) = 
  \left(
    \empmeas{
      \left[
        \psi(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n))^2
      \right]}
  \right)^{1/2}.
\end{equation*}

% When the nuisance parameters are consistently estimated at rate
% \( \smallO_P{(n^{-1/4})} \) and
% \begin{equation*}
%   P{\left(
%       \psi(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
%       \hat{b}_n(\data_n))
%       \in \mathcal{R}
%     \right)} \longrightarrow 1,
%   \quad \text{as} \quad   n \longrightarrow \infty , 
% \end{equation*}
% for some Donsker class of functions \( \mathcal{R} \), it holds that
% \citep{bickel1993efficient,kennedy2022semiparametric}

% According to equation~(\ref{eq:parametrizeP}) and equation~(\ref{eq:7}), any
% \( P \) is characterized by both \((\Lambda_1, \Lambda_2, \Gamma, H)\) and
% \( (F, H) \), which means that there exist operators $\tilde\Psi$ and $\bar\Psi$
% such that
% \begin{equation*}
%   % \label{eq:state-learner-identity}
%   \tilde\Psi(\Lambda_{1,0},\Lambda_{2,0}, \Gamma_0, H_0)=\Psi(P_0)=\bar\Psi(F_0,H_0).
% \end{equation*}
% Hence, any parameter $\Psi$ defined on \( \mathcal{P} \) can be estimated using
% either estimators \( (\hat{\Lambda}_n, \hat{\Gamma}_n, \hat{H}_n) \) or
% \( (\hat{F}_n, \hat{H}_n) \). When we use the state learner the final estimator
% of the target parameter is $\bar\Psi(\hat\phi_n(\data_n), \hat{H}_n)$, where
% \( \hat{H}_n \) can often be taken to be the marginal empirical measure.

% For the rest of this section we consider the case where $\Theta = \R$, and
% $\Psi$ is estimable by a regular asymptotically linear estimator. Specifically,
% we assume that it is possible to construct an estimator
% \( \tilde\Psi(\hat\Lambda_n, \hat\Gamma_n, H_n) \) such that the the following
% expansion holds:
% \begin{equation}
%   \label{eq:ral}
%   \tilde\Psi(\hat\Lambda_n, \hat\Gamma_n, H_n) - \Psi(P) = \empmeas{[\psi_P]} +
%   \mathrm{Rem}{(\hat{\Lambda}_{n},  \hat{\Gamma}_n, P)} + \smallO_{P}(n^{-1/2}),
% \end{equation}
% where \( \empmeas \) is the empirical measure of a sample \(\{O_i\}_{i=1}^n\),
% $\psi_P$ a zero-mean function with \( P[\psi_P^2]< \infty \), and
% \( \mathrm{Rem}{(\hat{\Lambda}_{n}, \hat{\Gamma}_n, P)} \) a second order
% remainder term \citep{bickel1993efficient,fisher2021visually}.

% ...

% The \( \smallO_P{(n^{-1/2})} \) bound on the products of estimation errors is
% typically a weaker condition than requiring \( \Gamma \), $\Lambda_1$, and
% $\Lambda_2$ to be estimated independently at rate \( n^{-1/2} \). This is
% particularly important when data-adaptive estimation methods are used as these
% methods rarely provide \( n^{-1/2} \)-rate convergence for function-valued
% nuisance parameters.

\subsection{Cause-specific average treatment effect}
\label{sec:cause-spec-aver}

We now demonstrate how the state learner and the general estimation strategy
outlined above can be used to estimate a concrete parameter of interest in the
competing risk setting. We assume that the covariate vector \( X \in \R^p \) can
be separated into a binary treatment indicator \( A \) and a vector of potential
confounders, \( W \in \mathcal{W}\subset \R^{p-1} \). We abuse notation slightly
by writing \( \Lambda_j(t \mid x) = \Lambda_j(t \mid a, w) \) and
\( S(t \mid x) = S(t \mid a, w) \) when \( x=(a, w) \). We use $\mu$ to denote
the marginal distribution of \( W \) and $\pi$ to denote the conditional
probability of treatment,
\begin{equation*}
  \pi(w) = P(A=1 \mid W=w).
\end{equation*}
We assume throughout that $\pi$ is uniformly bounded away from \( 0 \) and
\( 1 \) on \( \mathcal{W} \). As both \( A \) and \( W \) are fully observed for
all individuals we can use a super learner to estimate $\pi$
\fxnote*{add}{[ref]}, and we denote this estimator by $\hat{\pi}_n$. We use the
empirical measure of \( \{W_1, \dots, W_n\} \) to estimate $\mu$, and denote
this estimator by the $\hat{\mu}_n$. As parameter of interest we consider the
\fxnote*{what is this called?}{standardized difference in absolute risk of cause
  \( 1 \) before time $\tau$},
\begin{equation*}
  \theta_{{\tau}}(Q) = \int_{\mathcal{W}} 
  \left\{
    \int_0^{\tau}
    S(s- \mid w, 1)  \Lambda_1(\diff s \mid w, 1)
    -
    \int_0^{\tau}
    S(s- \mid w, 0)  \Lambda_1(\diff s \mid w, 0)
  \right\}
  \mu(\diff w).
\end{equation*}
Under a set of
additional assumptions, \( \theta_{{\tau}} \) can be given the causal
interpretation
\begin{equation*}
  \theta_{{\tau}}(Q) =
  \E{[T^{A=1} \leq \tau, D^{A=1}=1]}-
  \E{[T^{A=0} \leq \tau, D^{A=0}=1]},
\end{equation*}
where \( (T^A, D^A) \) denotes potential outcomes \citep{hernanRobinsWhatIf}. In
this case, the interpretation of $\theta_{\tau}$ is the difference in the
average risk of cause \( 1 \) occurring before time \( \tau \) in the population
if everyone had been given treatment (\( A=1 \)) compared to if no one had been
given treatment.

Using equation~(\ref{eq:surv-def}), we may write
\( \theta_{{\tau}}(Q) = \tilde{\Psi}_{t}^0(\Lambda_1, \Lambda_2, \mu) \),
where
\begin{equation}
  \label{eq:1}    
  \begin{split}
  \tilde{\Psi}_{t}^0(\Lambda_1, \Lambda_2, \mu) & =
  \int_{\mathcal{W}} 
  \int_0^{\tau}
  e^{-\Lambda_1(s- \mid w, 1)-\Lambda_2(s- \mid w, 1)}  \Lambda_1(\diff s \mid
  w, 1)
  \mu(\diff w)
  \\
  &  \quad
  -\int_{\mathcal{W}} 
  \int_0^{\tau}
  e^{-\Lambda_1(s- \mid w, 0)-\Lambda_2(s- \mid w, 0)}  \Lambda_1(\diff s \mid w, 0)
  \mu(\diff w).
  \end{split}
\end{equation}
The efficient influence function for the parameter $\tilde{\Psi}_{\tau}$
depends on the set \( (\Lambda_1, \Lambda_2, \Gamma, \pi) \) of nuisance
parameters, and is given by \fxnote*{other refs?}{\citep{rytgaard2022targeted}}
\begin{equation}
  \label{eq:5}
  \begin{split}
    & \psi_{\tau}(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
    \\
    & =
      \sum_{l=1}^{2} \int_0^{\tau}
      \left[
      h_{\tau,l,1}(t, O  ; \Lambda_1, \Lambda_2, \Gamma, \pi)
      - h_{\tau,l,0}(t, O  ; \Lambda_1, \Lambda_2, \Gamma, \pi)
      \right]
    \\
    & \quad\qquad  \qquad
      \times
      \left(
      N_l(\diff t) - \1{\{\tilde{T} \geq t\}}\Lambda_l(\diff t \mid A, W)
      \right)
    \\
    & \quad
      -\sum_{a=0}^{1} 
      \left\{
      (-1)^a
      \int_0^{\tau}
      e^{-\Lambda_1(s- \mid W, a)-\Lambda_2(s- \mid W, a)}  \Lambda_1(\diff s \mid
      W, a)
      \right\}
      - \tilde{\Psi}_{t}^0(\Lambda_1, \Lambda_2, \mu),
    % & \quad+ \int_0^{\tau} e^{-\Lambda_1(s- \mid W, 1)-\Lambda_2(s- \mid W,
    % 1)} \Lambda_1(\diff s \mid W, 1)
    % \\
    % & \quad - \int_0^{\tau} e^{-\Lambda_1(s- \mid W, 0)-\Lambda_2(s- \mid W,
    % 0)} \Lambda_1(\diff s \mid W, 0)
  \end{split}
\end{equation}
where
\begin{align*}
  h_{\tau,l,a}(t, O  ; \Lambda_1, \Lambda_2, \Gamma, \pi)
  =&  \frac{\1{\{A=a\}} e^{\Gamma(t-\mid A, W)}}{\pi(W)^{a}(1-\pi(W))^{1-a}}
  \\
   & \quad \times
     \left(
     1- \frac{\int_t^{\tau} e^{-\Lambda_1(s- \mid A, W)-\Lambda_2(s- \mid A, W)}
     \Lambda_1(\diff s \mid A, W)}
     {e^{-\Lambda_1(t \mid A, W)-\Lambda_2(t \mid A, W)}}
     \right)^{\1{\{l=1\}}}
  \\
   & \quad\times
     \left(
     - \frac{\int_t^{\tau} e^{-\Lambda_1(s- \mid A, W)-\Lambda_2(s- \mid A, W)}
     \Lambda_1(\diff s \mid A, W)}
     {e^{-\Lambda_1(t \mid A, W)-\Lambda_2(t \mid A, W)}}
     \right)^{\1{\{l=2\}}}
\end{align*}
Equation~(\ref{eq:1}) and~(\ref{eq:5}) then allow us to construct a one-step
estimator by using the definition given in equation~(\ref{eq:one-step-def}),
which gives the estimator
\begin{equation}
  \label{eq:one-step-comp-ate}
  \tilde{\Psi}_{t, j}^0(\hat{a}_1(\data_n), \hat{a}_2(\data_n), \hat{\mu}_n(\data_n))
\end{equation}

\section{Numerical experiments}
\label{sec:numer-exper}

We compare the state learner with two other discrete super learners and an
oracle selector in a simulation study without a competing event. The two other
super learners are based on inverse probability of censoring weighted (IPCW)
Brier scores \citep{graf1999assessment,gerds2006consistent}, and we refer to
these as IPCW super learners. These super learners depend on an estimator of the
censoring distribution, and we consider IPCW super learners that use either the
Kaplan-Meier estimator (IPCW(KM)) or a correctly specified Cox model (IPCW(Cox))
to estimate the censoring distribution. Both IPCW super learners are fitted
using the \texttt{R}-package \texttt{riskRegression}
\citep{Gerds_Ohlendorff_Ozenne_2023}. Each discrete super learner provides a
learner for the cumulative hazard function for the outcome of interest, and from
this a risk prediction model can be obtained. We measure performance of each
super learner in terms of the Brier score of the provided risk prediction model
at a specific time horizon. The Brier score is approximated using a large
(\( n = 20,000 \)) independent data set of uncensored data. The oracle selector
uses the large data set of uncensored event times to select the learner with the
lowest expected Brier score. The expected Brier score of the oracle selector
serves as a lower benchmark value. For all super learners we split the data into
five folds for training and testing.

Note that given a learner for the cumulative hazard function of the
outcome event, we can typically use the same method to construct a
learner of the cumulative hazard function of the censoring
distribution. This would typically work by training the learner on the
data set \( \data_n^c \), where \( \data_n^c = \{O_i^c\}_{i=1}^n \)
with \( O_i^c = (\tilde{T}_i, 1-\Delta, X_i) \). When we say that we
use a learner for the cumulative hazard function of the outcome to
learn the cumulative hazard function of the censoring time, we mean
that the learner is trained on \( \data_n^c \).

\paragraph{Scenario 1}

We first generate a simple dataset to demonstrate that an IPCW super learner can
perform poorly when the censoring model is misspecified. We start by generating a
binomial baseline covariate \( A \) with success probability \( 30\% \). We then
generate outcome and censoring variables according to a Cox-Weibull distribution
with hazard rates of approximately \( 0.5 \) and \( 2.5 \), respectively. We use
a library \( \mathcal{A} \) consisting of two learners, the (marginal)
Kaplan-Meier estimator and the Kaplan-Meier estimator stratified on the binary
covariate. For the state learner we use the same learners to construct the
library \( \mathcal{B} \).

The results are shown in Figure~\ref{fig:ipcw-fail}. We see that the IPCW super
learner based on a misspecified censoring model (IPCW(KM)) has a larger Brier
score than the other super learners and that its performance does not improve
much with sample size. % When the censoring model is misspecified, the
% IPCW(KM) super learner fails to identify the learner corresponding to the
% data-generating mechanism, even though this learner is included in the
% library.
The performance of the state learner is comparable to or slightly better than
the IPCW super learner based on a correctly pre-specified censoring model. The
performance of both the IPCW(Cox) super learner and the state learner are close
to that of the oracle for most sample sizes.

\begin{figure}
  \centerline{\includegraphics[width=1\linewidth]{./figures/ipcw-fail.pdf}}
  \caption[]{Results for scenario~1 of the simulation study. For the learner
    selected by each of the four discrete super learners, the Brier score
    calculated in a large independent data set without censoring is plotted
    against sample size. The results are based on 200 repetitions.}
  \label{fig:ipcw-fail}
\end{figure}

\paragraph{Scenario 2}

We next generate data according to a more complex distribution motivated from a
real dataset in which censoring depends on the baseline covariates. We simulate
data based on a prostate cancer study described in
\citep{kattan2000pretreatment}. The outcome of interest was the time to tumor
recurrence, and five baseline covariates were used to predict outcome:
prostate-specific antigen (PSA, ng/mL), Gleason score sum (GSS, values between 6
and 10), radiation dose (RD), hormone therapy (HT, yes/no) and clinical stage
(CS, six values). The study was designed such that a patient's radiation dose
depended on when the patient entered the study \citep{gerds2013estimating}. This
in turn implied that the time of censoring depended on the radiation dose.
% It was found that almost all covariates were predictive of censoring.
The data were re-analyzed in \citep{gerds2013estimating} where a sensitivity
analysis was conducted based on simulated data. We use the same simulation
setup, where event and censoring times are generated according to parametric
Cox-Weibull models estimated from the original data, and the covariates are
generated according to either marginal Gaussian normal or binomial distributions
estimated from the original data
\citep[c.f.,][Section~4.6]{gerds2013estimating}. We use the library consisting
of the nine learners described in Table~\ref{tab:zel-library}. For the state
learner we use the same library to learn the censoring distribution.

\begin{table}[h]
  \centering
  \begin{tabular}{ lll }
    \toprule
    Family & Model & Description \\
    \midrule
    Marginal & \texttt{KM} & The Kaplan-Meier estimator \\ 
    Cox & \texttt{Cox} & All five covariates included with additive effects \\
           & \( \texttt{Cox strata CS} \)  & Cox model stratified on CS  \\
           & \( \texttt{Cox strata HT} \)  & Cox model stratified on HT \\
           & \( \texttt{Cox spline} \)  & PSA and RD modeled with splines \\ 
    Penalized Cox & \texttt{Lasso} & Cox model with \( L_1 \)-norm penalty   \\
           & \texttt{Ridge} & Cox model with \( L_2 \)-norm penalty \\
           & \texttt{Elastic} & Cox model with \( L_1 \)- and \( L_2 \)-norm penalty \\
    Random forest & \texttt{RF} & Random forest with 50 trees and default settings \\
    \bottomrule 
  \end{tabular}
  \caption[]{Overview of the nine learners used in scenario~2 of the simulation
    study. The Kaplan-Meier estimator was fitted using the package
    \texttt{prodlim} \citep{Gerds_2019prodlim}. All Cox models included all five
    covariates in the model and were fitted using the package \texttt{survival}
    \citep{survival-package}. All penalized Cox models included all five
    covariates as linear predictors and were fitted using the package
    \texttt{glmnet} \citep{glmnet-cox,glmnet-glm}. The random forest was fitted
    with the package \texttt{randomForestSRC} \citep{rfsrc-paclage}.}
  \label{tab:zel-library}
\end{table}

The results are shown in Figure~\ref{fig:zelefski}. We see that the Brier score
of the IPCW super learner based on a misspecified censoring model (IPCW(KM))
decreased with sample size, but is larger than that of the other super learners
for all sample sizes. The state learner and the IPCW super learner based on a
correctly pre-specified censoring model demonstrate similar performance for all
sample sizes. The performance of both the IPCW(Cox) super learner and the state
learner approaches the benchmark provided by the oracle selector for large sample
sizes.
\begin{figure}
  \centerline{\includegraphics[width=1\linewidth]{./figures/zelefski-sim.pdf}}
  \caption[]{Results for scenario~2 of the simulation study. For the learner
    selected by each of the four discrete super learners, the Brier score
    calculated in a large independent data set without censoring is plotted
    against sample size. The results are based on 200 repetitions.}
  \label{fig:zelefski}
\end{figure}

\section{Real data application}
\label{sec:real-data-appl}

The original prostate cancer data analyzed by \cite{kattan2000pretreatment},
which we introduced in Section~\ref{sec:numer-exper}, include a competing event
in the form of death without tumor recurrence. To illustrate our method we fit
the state learner to the original data set consisting of 1,042 patients. We
consider death without tumor recurrence and recurrence of tumor as two competing
events of interest. We include the fives learner \texttt{KM}, \texttt{Cox strata
  CS}, \texttt{Lasso}, \texttt{Elastic}, and \texttt{RF} which are described in
Table~\ref{tab:zel-library}. We use the same library of learners to learn
\( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$. In this case, \( \Lambda_1 \)
denotes the cause-specific cumulative hazard function of tumor recurrence, and
\( \Lambda_2 \) denotes the cause-specific cumulative hazard function of death
without tumor recurrence.

This gives a library consisting of \( 5^3 = 125 \) learners for the conditional
state occupation probability function \( F \) defined in
equation~(\ref{eq:F-def}). We use five folds for training and testing the
models, and we repeat training and evaluation five times with different splits.
The integrated Brier score for all learners are shown in
Figure~\ref{fig:zelefski-real}, and the top 10 combinations of learners are
displayed in Table~\ref{tab:zelefski-real}. We see that the prediction
performance is mostly affected by the choice of learner for the censoring
distribution. Several combinations of learners give similar performance as
measured by the integrated Brier score, as long as a random forest is used to
model the censoring distribution.

\begin{figure}
  \centerline{\includegraphics[width=1\linewidth]{./figures/zelefski-real-data.pdf}}
  \caption[]{The results of applying the 125 combinations of learners to the
    prostate cancer data set. The learners are \texttt{KM} (KM), \texttt{Cox
      strata CS} (strata), \texttt{Lasso} (lasso), \texttt{Elastic} (elastic),
    and \texttt{RF} (RF) as described Table~\ref{tab:zel-library}. The error
    bars are based on five repetitions using different splits. We refer to
    learners of \( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$ as `Tumor
    learner', `Death learner', and `Censoring learner', respectively.}
  \label{fig:zelefski-real}
\end{figure}

\begin{table}[ht]
  \centering \input{./figures/zel-tab.tex}
  \caption{The 10 best performing models in terms of integrated Brier score. The
    reported standard errors are based on five repetitions using different
    splits. The models are described in Table~\ref{tab:zel-library}. We refer to
    learners of \( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$ as `Tumor
    learner', `Death leaner', and `Censoring learner', respectively.}
  \label{tab:zelefski-real}
\end{table}
\section{Discussion}
\label{sec:discussion}

We have proposed a new super learner that can be used with right-censored data
and competing events. In this section, we compare our proposal to existing super
learners and discuss avenues for further research.

\subsection{Existing super learners for right-censored data}
\label{sec:relat-liter-exist}

Machine learning based on right-censored data commonly use the negative partial
log-likelihood as loss function
\citep[e.g.,][]{li2016regularized,yao2017deep,lee2018deephit,katzman2018deepsurv,gensheimer2019scalable,lee2021boosted,kvamme2021continuous}.
However, this loss function is unsuited for super learning, because many
canonical survival learners (e.g., the Kaplan-Meier estimator, random survival
forest, and semi-parametric Cox models) provide cumulative hazard functions that
are piece-wise constant in the time argument, and hence assign zero probability
to event times not observed in the training data. This implies that when data
are observed in continuous time, any of these learners will almost surely have
infinite loss in any independent hold-out sample according to the negative
partial log-likelihood loss. When a proportional hazards model is assumed, the
baseline hazard function can be profiled out of the likelihood to give a new
partial log-likelihood loss \citep{cox1972regression}, which has been suggested
as a loss function for super learning
\citep{golmakani2020super,verweij1993cross}. While this allows the library of
learners to include Cox' proportional hazard models, the drawback is that the
library is in fact \textit{only} allowed to include these models. The advantage
of the state learner is that it does not require evaluation of the density of
\( F(\blank, j, x) \) and does not assume a particular semi-parametric structure
for $\Lambda_j$ but can be used with any library of learners.

Another approach for super learning with right-censored data is to use an
inverse probability of censoring weighted (IPCW) loss function
\citep{graf1999assessment,van2003unicv,molinaro2004tree,keles2004asymptotically,hothorn2006survival,gerds2006consistent,gonzalez2021stacked}.
An IPCW loss function is attractive because the associated risk does not depend
on the censoring distribution but describes a feature of the population of
interested governed by the measure \( Q \in \mathcal{Q} \). Similar results can
be obtained using censoring unbiased transformations
\citep{fan1996local,steingrimsson2019censoring} or pseudo-values
\citep{andersen2003generalised,mogensen2013random,sachs2019ensemble}. All these
methods rely on an estimator of the censoring distribution, and their drawback
is that this estimator has to be pre-specified. When the data-generating
mechanism is complex and not well-understood, pre-specification of the censoring
distribution is a challenge. The advantage of using the state learner is that a
censoring distribution need not be pre-specified but is selected automatically
based on the provided library \( \mathcal{B} \).

To the best of our knowledge, the only existing attempt at avoiding the need to
pre-specify a censoring model is a recent proposal suggested independently by
\cite{han2021inverse} and \cite{westling2021inference}. The authors do not
consider competing risks but suggest to iterate between learning \( \Lambda \)
and $\Gamma$ using IPCW loss functions and select the final learner when the
iterative procedure has converged.
% Using the estimator \( \hat{b}^{(0)}(\data_n^c) \) of $\Gamma$ based on some
% initial learner \( \hat{b}^{(0)} \in \mathcal{B} \), we calculate the
% estimated IPCW risk for each leaner in \( \mathcal{A} \). Picking
% \( a^{(1)} \) as the discrete super learner according to this IPCW loss
% function, we then estimate weight
No general theoretical guarantees exist for this procedure, but it would be
interesting to compare its performance to that of the state learner in a
simulation study.

% In a non-survival targeted learning setting \cite{sun2022selective}, based on
% ideas presented by \cite{robins2007comment}, proposed to use as a model
% selection criteria the so-called `doubly robustness' property that some
% targeted estimators enjoy. In many special cases the remainder term
% \( \mathrm{Rem}{(\hat{\Lambda}_{n}, \hat{\Gamma}_n, P)} \) defined in
% equation~(\ref{eq:ral}) has the property that
% \begin{equation}
%   \label{eq:dr-property}
%   \mathrm{Rem}{(\Lambda^*, \Gamma_0, P_0)}
%   = \mathrm{Rem}{(\Lambda_0, \Gamma^*, P_0)} = 0,
%   \quad \text{for any }
%   \Lambda^*, \Gamma^* \in \mathcal{M}.
% \end{equation}
% The property in equation~(\ref{eq:dr-property}) is known as `doubly
% robustness' because it ensures that
% $\tilde{\Psi}(\hat{\Lambda}_{n}, \hat{\Gamma}_n, H_n)$ defined in
% equation~(\ref{eq:ral}) is consistent if just one of $\Lambda$ or $\Gamma$ is
% estimated consistently
% \citep{van2003unified,bang2005doubly,kang2007demystifying}. For
% $a \in \mathcal{A}$ and \( b \in \mathcal{B} \), define
% \( \hat{\Psi}^k_n(a, b) = \tilde{\Psi}(a(\data_n^k), b(\data_n^k), \empmeas^k) \)
% where \( \empmeas^{k} \) is the empirical measure of \( \data_n^{k} \).
% \cite{sun2022selective} suggest to select $a \in \mathcal{A}$ as the minimizer
% of the estimated `fluctuation pseudo-risk',
% \begin{equation*}
%   \hat R_{\mathrm{f}}(a) =
%   \max_{b_1, b_2 \in \mathcal{B}}
%   \frac{1}{K}\sum_{k=1}^{K}
%   \left\{
%     \hat{\Psi}^k_n(a, b_1)
%     - \hat{\Psi}^k_n(a, b_2)
%   \right\}^2.
% \end{equation*}
% The idea is that the doubly robustness property
% (equation~(\ref{eq:dr-property})) implies that \(\hat{R}_{\mathrm{f}}(a)\) will
% be zero (asymptotically) if the learner \(a\) correctly estimates $\Lambda$. The
% authors establish finite sample results guarantying that the selected learner
% \(a\) will be robust against changing the learner of the other nuisance
% parameter across $\mathcal{B}$. It seems unclear to what extend this robustness
% property also guarantees consistency of the estimator of the target parameters
% when we use the fluctuation risk \(\hat{R}_{\mathrm{f}}\) to select the learners
% of the nuisance parameters. It would be interesting to explore the performance
% of this procedure in a survival setting and compare it to targeted estimators
% obtained using the state learner as outlined in
% Section~\ref{sec:targeted-learning}.

\subsection{A performance measure of interest}
\label{sec:perf-meas-inter}

A major advantage of the state learner is that performance of each combination
of learners is measured in terms of observable quantities. This means that no
additional nuisance parameters need to be estimated to evaluate the loss. The
drawback with this approach is that we are rarely interested in features of the
observed data distribution when the data are right-censored. The finite sample
oracle inequality in Corollary~\ref{cor:oracle-prop} concerns the function
\( F \), which is a feature of \( P \in \mathcal{P} \), while what we are
typically interested in is \( \Lambda_j \) or \( S \), which are features of
\( Q \in \mathcal{Q} \). We emphasize that while the state learner provides us
with estimates of \( \Lambda_j \) and $\Gamma$ based on libraries
\( \mathcal{A}_j \) and \( \mathcal{B} \), performance are not assessed directly
for these parameters, but only jointly for estimation of the parameter \( F \).
For settings without a competing risk, our numerical studies suggest that
measuring performance with respect to estimation of \( F \) also leads to good
performance for estimation of \( S \). Further research on this topic, both
numerical and theoretical, is warranted.

\fxnote*{mention this?}{[Double robustness maybe lacking]}

% In the context of targeted learning, a drawback of the state learner is that the
% doubly robustness property defined in equation~(\ref{eq:dr-property}) seem to be
% lost because we only use a single nuisance parameter estimator. As defined here,
% however, the state learner is build using libraries for the conditional
% cause-specific cumulative hazard functions, so some doubly robustness might be
% preserved.

\subsection{Implementation}
\label{sec:software}

Our proposed super learner can be implemented with a broad library of learners
using existing software, for instance the \texttt{R}-package
\texttt{riskRegression} \citep{Gerds_Ohlendorff_Ozenne_2023}. Furthermore, while
the library \( \mathcal{F}(\mathcal{A}_1,\mathcal{A}_2,\mathcal{B}) \) consists
of \( |\mathcal{A}_1||\mathcal{A}_2||\mathcal{B}| \) learners, as long as we
have sufficient memory we need only fit
\( |\mathcal{A}_1| +|\mathcal{A}_2| + |\mathcal{B}| \) learners in each fold. To
evaluate the performance of each learner we need to perform
\( |\mathcal{A}_1||\mathcal{A}_2||\mathcal{B}| \) operations to calculate the
integrated Brier score in each hold-out sample, one for each combination of the
fitted models, but these operations are often negligible compared to fitting the
models. Hence the state learner is essentially no more computationally demanding
than any procedure that uses super learning to learn $\Lambda_1$, $\Lambda_2$,
and $\Gamma$ separately. While our proposal is based on constructing the library
\( \mathcal{F} \) from libraries for learning \( \Lambda_1 \), $\Lambda_2$, and
$\Gamma$, it would also be of interest to consider learners that estimate
\( F \) directly.

In our numerical studies, we only considered learners of $\Lambda_j$ and
$\Gamma$ that provide cumulative hazard functions which are piece-wise constant
in the time argument. This simplifies the calculation of \( F \) as the
integrals in equation~(\ref{eq:transition}) reduce to sums. When $\Lambda_j$ or
\( \Gamma \) are absolutely continuous in the time argument, calculating \( F \)
is more involved, but we expect that a good approximation can be achieved by
discretization. In the future, we intend to investigate the performance of the
state learner when using a broader library of learners in more comprehensive
simulation studies.

% \bibliography{bib.bib}

\appendix

\section{Theoretical guarantees for the state learner}
\label{sec:proof-proposition}

In this section we provide proofs of the results stated in
Section~\ref{sec:theor-results-prop}.

Define
\( \bar{B}_{\tau,0}(F, o) = \bar{B}_{\tau}(F, o) - \bar{B}_{\tau}(F_0, o) \) and
\( R_{0}(F) = P_0{[\bar{B}_{\tau,0}(F, \blank)]} \).
\begin{lemma}
  \label{lemma:norm}
  \( R_{0}(F) = \Vert F - F_0 \Vert_{P_0}^2 \), where \( \Vert \blank \Vert_{P_0}\) is defined
  in equation~(\ref{eq:norm}).
\end{lemma}
\begin{proof}
  For any \( t \in [0, \tau] \) and \( k\in \{0,1,2\} \) we have
  \begin{align*}
    & \E_{P_0}{\left[ (F(t, k, X) - \1{\{\eta(t) = k \}})^2 \right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, k, X) - F_0(t, k, X) + F_0(t, k, X) - \1{\{\eta(t) = k
      \}})^2 \right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, k, X) - F_0(t, k, X))^2\right]}
      + \E_{P_0}{\left[ (F_0(t, k, X) - \1{\{\eta(t) = k \}})^2\right]}
    \\
    & \quad
      + 2\E_{P_0}{\left[ (F(t, k, X) - F_0(t, k, X))(F_0(t, k, X) - \1{\{\eta(t) = k
      \}})\right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, k, X) - F_0(t, k, X))^2\right]}
      + \E_{P_0}{\left[ (F_0(t, k, X) - \1{\{\eta(t) = k \}})^2\right]},
  \end{align*}
  where the last equality follows from the tower property. Hence, using Fubini,
  we have
  \begin{equation*}
    P{[\bar{B}_{\tau}(F, \blank)]}
    = \Vert F - F_0 \Vert_{P_0}^2 + P_0{[\bar{B}_{\tau}(F_0, \blank)]}. \qedhere
  \end{equation*}
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:stric-prop}]
  The result follows from Lemma~\ref{lemma:norm}.
\end{proof}

In the following, let $\Theta$ denote the function space consisting of all
conditional state occupation probability functions for some measure \( P \) as
defined in equation~(\ref{eq:F-def}).

\begin{proof}[Proof of Corollary~\ref{cor:oracle-prop}]
  First note that minimising the loss \( \bar{B}_{\tau} \) is equivalent to
  minimising the loss \( \bar{B}_{\tau,0} \), so the discrete super learner and
  oracle according to \( \bar{B}_{\tau} \) and \( \bar{B}_{\tau,0} \) are
  identical. By Lemma~\ref{lemma:norm}, \( R_0(F) \geq 0 \) for any
  \( F \in \Theta \), and so using Theorem 2.3 from \citep{vaart2006oracle} with
  \( p=1 \), we have that for all \( \delta >0 \),
\begin{align*}
  \E_{P_0}{\left[ R_0(\hat{\phi}_n(\data_n^{-k})) \right]}
  \leq
  &(1+2\delta)\E_{P_0}{\left[ R_0(\tilde{\phi}_n(\data_n^{-k})) \right]}
  \\
  & \quad + (1+\delta) \frac{16 K}{n}
    \log(1 + |\mathcal{F}_n|)\sup_{F \in \Theta}
    \left\{
    M(F) + \frac{v(F)}{R_0(F)}
    \left(
    \frac{1}{\delta} + 1
    \right)
    \right\}
\end{align*}
where for each \( F \in \Theta \), \( (M(F), v(F)) \) is some Bernstein pair for
the function \(o \mapsto \bar{B}_{\tau,0}(F, o) \). As
\( \bar{B}_{\tau,0}(F, \blank) \) is uniformly bounded by \( \tau \) for any
\( F \in \Theta \), it follows from section 8.1 in \citep{vaart2006oracle} that
\( (\tau, 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}) \) is a Bernstein pair for
\( \bar{B}_{\tau,0}(F, \blank) \). Now, for any \( a,b,c \in \R \) we have
\begin{align*}
  (a-c)^2 - (b-c)^2
  & = (a-b+b-c)^2 - (b-c)^2
  \\
  & = (a-b)^2 + (b-c)^2 +2(b-c)(a-b) - (b-c)^2
  \\
  & = (a-b)
    \left\{
    (a-b) +  2(b-c)
    \right\}
  \\
  & = (a-b)
    \left\{
     a + b -2c
    \right\},
\end{align*}
so using this with \( a=F(t, k, x) \), \( b=F_0(t, k, x) \), and
\( c = \1{\{\eta(t) = k\}} \), we have by Jensen's inequality
\begin{align*}
  & P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}
  \\
  & \leq
    2\tau\E_{P_0}{\left[
    \sum_{k=0}^{3} \int_0^{\tau}
    \left\{
    \left(
    F(t, k, X) - \1{\{\eta(t) = k\}}
    \right)^2
    -
    \left(
    F_0(t, k, X) - \1{\{\eta(t) = k\}}
    \right)^2
    \right\}^2
    \diff t 
    \right]}
  \\
  & =2\tau
    \E_{P_0}\Bigg[
    \sum_{k=0}^{3} \int_0^{\tau}
    \left(
    F(t, k, X) - F_0(t, k, X)
    \right)^2
  \\
  & \quad \quad \quad\quad \quad \quad \times
    \left\{
    F(t, k, X) +  F_0(t, k, X)-2 \1{\{\eta(t) = k\}}
    \right\}^2
    \diff t 
    \Bigg]
  \\
  & \leq
    8\tau \E_{P_0}{\left[
    \sum_{k=0}^{3} \int_0^{\tau}
    \left(
    F(t, k, X) - F_0(t, k, X)
    \right)^2
    \diff t 
    \right]}.
  \\
  & =
    8\tau \Vert F - F_0 \Vert_{P_0}^2.
\end{align*}
Thus when \( v(F) = 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]} \) we have by
Lemma~\ref{lemma:norm}
\begin{equation*}
  \frac{v(F)}{R_0(F)}
  = 1.5 \frac{P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}}{P_0{[\bar{B}_{\tau,0}(F, \blank)]}}
  \leq 12 \tau,
\end{equation*}
and so using the Bernstein pairs \( (\tau, 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}) \) we have
\begin{equation*}
  \sup_{F \in \Theta}
  \left\{
    M(F) + \frac{v(F)}{R_0(F)}
    \left(
      \frac{1}{\delta} + 1
    \right)
  \right\}
  \leq \tau
  \left(
    13 + \frac{12}{\delta}
  \right),
\end{equation*}
For all $\delta>0$ we thus have
\begin{align*}
  \E_{P_0}{\left[ R_0(\hat{\phi}_n(\data_n^{-k})) \right]}
  \leq
  &(1+2\delta)\E_{P_0}{\left[ R_0(\tilde{\phi}_n(\data_n^{-k})) \right]}
  \\
  & \quad
    + (1+\delta)\log(1 + |\mathcal{F}_n|) \tau \frac{16 K}{n}
    \left(
    13 + \frac{12}{\delta}
    \right),
\end{align*}
and then the final result follows from Lemma~\ref{lemma:norm}.
\end{proof}

\begin{proof}[Proof of Corollary~\ref{cor:asymp-cons}]
  By definition of the oracle and Lemma~\ref{lemma:norm},
  \( \E_{P_0}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
  \leq \E_{P_0}{\left[ \Vert \phi_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]} \) for
  all \( n \in \N \). The results then follows from
  Corollary~\ref{cor:oracle-prop}.
\end{proof}

\section{The state learner with targeted learning}
\label{sec:state-learner-with}

In this section show that a product structure is preserved when the estimator
$\bar\Psi(\hat{F}_n, \hat{H}_n)$ is used instead of
$\tilde\Psi(\hat{\Lambda}_n, \hat{\Gamma}_n, \hat{H}_n)$.

\begin{proof}[Proof of Proposition~\ref{prop:dr-structure}]
  For notational convenience we suppress \( X \) in the following. The final
  result can be obtained by adding the argument \( X \) to all functions and
  averaging. We use the relations from equation~(\ref{eq:7}) to write
  \begin{align*}
    & \int_0^{\tau} w(s) 
      \left\{
      \Gamma(s) - \hat{\Gamma}_n(s)
      \right\}
      [\Lambda - \hat{\Lambda}_n](\diff s)
    \\
    & =
      \int_0^{\tau} w(s) 
      \left\{
      \int_0^s \frac{F(\diff u, 2)}{F(u-, 0)} -
      \int_0^s \frac{\hat{F}_n(\diff u, 2)}{\hat{F}_n(u-, 0)}  -
      \right\}
      \left[
      \frac{F(\diff s, 1)}{F(s-, 0)}
      - \frac{\hat{F}_n(\diff s, 1)}{\hat{F}_n(s-, 0)}
      \right]
    \\
    & =
      \int_0^{\tau} w(s) 
      \Bigg\{
      \int_0^s 
      \left(
      \frac{1}{F(u-, 0)} -  \frac{1}{\hat{F}_n(u-, 0)}
      \right) F(\diff u, 2)
    \\
    & \qquad\qquad \qquad
      +
      \int_0^s \frac{1}{\hat{F}_n(u-, 0)} 
      \left[
      F(\diff u, 2) - \hat{F}_n(\diff u, 2)
      \right]
      \Bigg\}
    \\
    & \qquad\qquad \times
      \left[
      \left(
      \frac{1}{F(s-, 0)} -
      \frac{1}{\hat{F}_n(s-, 0)}
      \right)F(\diff s, 1)
       + \frac{1}{\hat{F}_n(s-, 0)}
      \left(
      F(\diff s, 1) -
      \hat{F}_n(\diff s, 1)
      \right)
      \right]
    \\
    &
      = \int_0^{\tau} 
      \int_0^s
      w(s) 
      \left(
      \frac{1}{F(u-, 0)} -  \frac{1}{\hat{F}_n(u-, 0)}
      \right) 
      \left(
      \frac{1}{F(s-, 0)} -
      \frac{1}{\hat{F}_n(s-, 0)}
      \right)F(\diff u, 2)F(\diff s, 1)
    \\
    & \quad +
      \int_0^{\tau}
      \int_0^s
      w(s) 
      \left(
      \frac{1}{F(u-, 0)} -  \frac{1}{\hat{F}_n(u-, 0)}
      \right) \frac{F(\diff u, 2) }{\hat{F}_n(u-,0)}
      \left(
      F(\diff s, 1) -
      \hat{F}_n(\diff s, 1)
      \right)
    \\
    & \quad +
      \int_0^{\tau} 
      \int_0^s      
      \frac{w(s) }{\hat{F}_n(u-, 0)} 
      \left[
      F(\diff u, 2) - \hat{F}_n(\diff u, 2)
      \right]
      \left(
      \frac{1}{F(s-, 0)} -
      \frac{1}{\hat{F}_n(s-, 0)}
      \right)F(\diff s, 1)
    \\
    & \quad +
      \int_0^{\tau} 
      \int_0^s      
      \frac{w(s) }{\hat{F}_n(u-, 0)} 
      \left[
      F(\diff u, 2) - \hat{F}_n(\diff u, 2)
      \right]
      \frac{1}{\hat{F}_n(s-, 0)}
      \left(
      F(\diff s, 1) -
      \hat{F}_n(\diff s, 1)
      \right).
  \end{align*}
  Consider the first term on the right hand side. Defining
  \begin{equation*}
    w_n^*(t)  = 
    \left(
      F(t-, 0)
      - \hat{F}_n(t-, 0)
    \right)
    \left(
      \frac{1}{F(t-, 0)}
      - \frac{1}{\hat{F}_n(t-, 0)}
    \right),
  \end{equation*}
  we can write
  \begin{align*}
    & \int_0^{\tau} 
      \int_0^s
      w(s) 
      \left(
      \frac{1}{F(u-, 0)} -  \frac{1}{\hat{F}_n(u-, 0)}
      \right)      
      \left(
      \frac{1}{F(s-, 0)} -
      \frac{1}{\hat{F}_n(s-, 0)}
      \right)F(\diff u, 2)F(\diff s, 1)
    \\
    & =
      \int_0^{\tau} 
      \int_0^s
      w(s)
      w_n^*(u) 
      \left(
      F(u-, 0) - \hat{F}_n(u-, 0)
      \right)
    \\
    & \qquad \qquad \quad
      \times
      w_n^*(s) 
      \left(
      F(s-, 0) - \hat{F}_n(s-, 0)
      \right)       
      F(\diff u, 2)F(\diff s, 1)
    \\
    & =
      \int_0^{\tau} 
      \int_0^s
      w_n^a(s,u)
      \left(
      F(u-, 0) - \hat{F}_n(u-, 0)
      \right)
      \left(
      F(s-, 0) - \hat{F}_n(s-, 0)
      \right)       
      F(\diff u, 2)F(\diff s, 1),
  \end{align*}
  where we have defined \( w_n^a(s,u) = w(s)w^*_n(s)w^*_n(u) \). By assumption,
  \( w_n^a(s,u) \) is uniformly bounded. The same approach can be applied to the
  three remaining terms which gives the result.
\end{proof}

\bibliography{bib.bib}

\end{document}