### 2023-11-29 experiments/tmp-summarize-results.tex
simulation results extracted and simplified

% Created 2023-11-29 Wed 11:31
% Intended LaTeX compiler: pdflatex
\documentclass[a4paper,danish]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\usepackage{amsmath}
\usepackage{array}
\usepackage[T1]{fontenc}
\usepackage{natbib}
\usepackage[margin=4cm]{geometry}
\usepackage{dsfont, pgfpages, tikz,amssymb, amsmath,xcolor, caption, subcaption}
\input{./latex-settings/standard-settings.tex}
\input{./latex-settings/standard-commands.tex}
\input{./latex-settings/org-settings.tex}
\input{./latex-settings/title-compact.tex}
\author{Anders Munch}
\date{\today}
\title{Summary of some simulation results}
\begin{document}

\maketitle
\section{The state learner vs IPCW super learners}
\label{sec:org4c83351}

We compare the state learner to two IPCW-based super learners using different
estimators of the censoring distribution. As benchmark we use the (discrete)
oracle. This gives the following 4 super learners.

\begin{enumerate}
\item The state learner (referred to as \texttt{statelearner}).
\item A super learner based on the estimated integrated Brier score, where the
censoring mechanism is estimated with the Kaplan-Meier estimator (referred to
as \texttt{ipcw\_km}).
\item A super learner based on the estimated integrated Brier score, where the
censoring mechanism is estimated with a Cox model that includes all available
covariates as main effects (referred to as \texttt{ipcw\_cox}).
\item The (discrete) oracle super learner, which picks the model that minimizes the
Brier score in an independent data set of 10.000 uncensored samples (referred to
as \texttt{oracle}).
\end{enumerate}

We used data generated in two different ways:
\begin{enumerate}
\item Data as generated in \citep{gerds2013estimating} (referred to as \texttt{original}).
\item As in 1., but where censoring in completely independent of covariates
(referred to as \texttt{indep\_cens}).
\end{enumerate}

To evaluate performance super learner, we use an independent data set of 10.000
uncensored samples and calculate the integrated Brier score in this data set.
All results are based on 500 simulated data sets.

\begin{center}
\includegraphics[width=.9\linewidth]{/tmp/tmp-fig-sl-plot1.pdf}
\end{center}

\clearpage

\section{The state learner vs SurvSL}
\label{sec:org09707c3}

We compare the state learner to the super learner proposed by
\cite{westling2021inference}. As benchmark we use again the (discrete) oracle.
This gives the following 3 super learners.

\begin{enumerate}
\item The state learner (referred to as \texttt{statelearner}).
\item The super learner proposed in \citep{westling2021inference} (referred to as
\texttt{survSL}).
\item The (discrete) oracle super learner, which picks the model that minimizes the
Brier score in an independent data set of 10.000 uncensored samples (referred to
as \texttt{oracle}).
\end{enumerate}

We use the same data-generating mechanisms and method of evaluation as described
in the section above.

\begin{center}
\includegraphics[width=.9\linewidth]{/tmp/tmp-fig-sl-plot2.pdf}
\end{center}

\section{References}
\label{sec:orgdc8dbd8}
\renewcommand{\section}[2]{} 
\bibliography{./latex-settings/default-bib.bib}
\end{document}