### 2024-04-29 tex/jrss-version/state-learner-jrss-style.tex

\documentclass{statsoc}
\usepackage[a4paper]{geometry}

\usepackage{natbib, dsfont, enumitem,
  amssymb,soul,xcolor,amsmath,graphicx,verbatim,pgfplots,tikz,prodint,booktabs}

\usepackage[tablesonly,heads,nolists]{endfloat}

%% Theorem environment
% \theoremstyle{plain} % plain italic style
\newtheorem{theorem}{Theorem}
% \numberwithin{theorem}{section}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}

%% Only for comments:
\usepackage[author=]{fixme}
\fxusetheme{color}
\definecolor{fxtarget}{rgb}{.5,.5,.5}
\definecolor{fxnote}{rgb}{.5,.5,.5}
\fxsetup{status=draft}

\bibliographystyle{rss}

%% Toggle with or without figures
% \def\nofig{1}

%% TODO: Update figures with name 'figure-'.
%% Write macro for including figure or not.

% \newcommand{\includeFigCond}[2]{
%   \ifx\nofig\undefined %
%     \includegraphics[#1]{#2} %
%   \else %
%     \texttt{#2} %
%   \fi %
% }
\newcommand{\includeFigCond}[2][]{
  \ifx\nofig\undefined %
    \includegraphics[#1]{#2} %
  \else %
    \texttt{#2} %
  \fi %
}

% New operators and commands
\DeclareMathOperator{\E}{\mathbb{E}} % expectation
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\C}{\mathbb{C}}
\renewcommand{\S}{\mathbb{S}}
\newcommand{\blank}{\makebox[1ex]{\textbf{$\cdot$}}}
\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
\renewcommand{\phi}{\varphi}
\renewcommand{\epsilon}{\varepsilon}
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand{\weakly}{\rightsquigarrow}
\newcommand\smallO{\textit{o}}
\newcommand\bigO{\textit{O}}
\newcommand{\midd}{\; \middle|\;}
\newcommand{\1}{\mathds{1}}
\usepackage{ifthen} %% Empirical process with default argument
% \newcommand{\G}[1][]{%
%    \ifthenelse{ \equal{#1}{} }
%       {\ensuremath{\mathbb{G}_n}}
%       {\ensuremath{\mathbb{G}_{#1}}}
% }
% New version:
\newcommand{\G}[2][n]{
{\ensuremath{\mathbb{G}_{#1}}{\left[#2\right]}}
}
\DeclareMathOperator*{\argmin}{\arg\!\min}
\DeclareMathOperator*{\argmax}{\arg\!\max}
\newcommand{\empmeas}{\ensuremath{\mathbb{P}_n}} % empirical measure

\newcommand{\data}{\ensuremath{\mathcal{D}}}

\title[The state learner]{The state learner \\ {-- a super learner for right-censored data}}
\author{Anders Munch}
\address{Section of Biostatistics, University of Copenhagen, Copenhagen, Denmark}
\email{a.munch@sund.ku.dk}
\author[Munch {\it et al.}]{Thomas A.\ Gerds}
\address{Section of Biostatistics, University of Copenhagen, Copenhagen, Denmark}

\begin{document}

\begin{abstract}
  In survival analysis, prediction models are needed as stand-alone
  tools and in applications of causal inference to estimate nuisance
  parameters. The super learner is a machine learning algorithm which
  combines a library of prediction models into a meta learner based on
  cross-validated loss. In right censored data, loss needs to be estimated somehow.
  % based on a model for the censoring distribution.
  % Unfortunately, the
  % commonly used partial likelihood loss is not suited for super learning, and
  % inverse probability of censoring weighted loss functions require a
  % pre-specified estimator of the censoring distribution. To relax this,
  We introduce the state learner, a new super learner for survival analysis, which
  simultaneously evaluates libraries
  of prediction models for the event of interest and the censoring
  distribution. The state learner can
be applied to all types of survival models, works in the presence
of competing risks, and does not require a single pre-specified
estimator of the conditional censoring distribution.
  We establish an oracle inequality for the state learner and
  investigate its performance through numerical experiments. We illustrate how
  the state learner allows us to estimate causal effects in a competing risks
  setting without having to pre-specify models for neither the cause-specific
  hazards of interest nor the censoring distribution.
\end{abstract}

\keywords{Competing risks, cross-validation, loss based estimation,
  right-censored data, super learner}

\section{Introduction}
\label{sec:introduction}

Medical decision making is traditionally based on the discrepancy
between intuitive jugdment and the law of chance
\citep{redelmeier1995probability}. We are concerned with the
statistical models that provide probabilistic predictions about the
medical future of a patient. We have two different applications of
these models in mind. One is as a stand-alone tool where the model
provides information for a freshly diagnosed patient. The other is as
input to an algorithm which estimates a causal target parameter such
as an average treatment effect. The models are learned on data from
previous patients. A super learner is a machine learning algorithm
that combines a finite set of learners into a meta learner by
estimating prediction performance in hold-out samples using a
pre-specified loss function \citep{van2007super}. When the aim is to
make a prediction model, super learners combine strong learners, such
as Cox regression models and random survival forests
\citep{gerds2021medical}.
% For targeted learning of a low dimensional
% target parameter in presence of a high-dimensional nuisance parameter,
% a super learner will usually also include the highly adaptive lasso
% \citep{van2011targeted, benkeser2016highly,van2017generally}.

While the general idea of combining strong learners based on
cross-validation data stems from earlier work
\citep{wolpert1992stacked,breiman1996stacked}, the name super learner
is justified by an oracle inequality
\citep{van2003unicv,vaart2006oracle}: The super learner is guaranteed
to perform almost as well as the model which minimises the expected
performance, i.e., the model we would select if we could evaluate the
prediction performance in an infinite hold-out sample. The choice of
the loss function for super learning in survival analysis deserves a
closer look. Existing super learner algorithms for right-censored data
use partial log-likelihood loss or inverse probability of censoring
weighted loss
\citep{polley2011-sl-cens,keles2004asymptotically,golmakani2020super,westling2021inference}.
The use of the partial log-likelihood loss, however, restricts the
class of learners and excludes for example simple Kaplan-Meier based
learners and also more complex random survival forest algorithms. For
this reason \cite{golmakani2020super} restrict their learners to Cox
proportional hazard models. A lesser known fact is that a super
learner constructed with the negative partial log-likelihood loss
implicitly depends on the censoring distribution
\citep{hjort1992inference,whitney2019comment}. Inverse probability of
censoring weighted loss has been studied for decades
\citep{graf1999assessment, hothorn2006survival, gerds2006consistent}. A disadvantage of inverse probability of
censoring weighted loss estimators is that they require a
pre-specified model for the censoring
distribution. In \cite{westling2021inference} this problem is tackled by
iterating between super learning of the censoring distribution and the
event time
distribution. % ANDERS: mangler afrundning efter sidste saetning. eller maaske kan saetningen flyttes laenger ned?

In this article we define the state learner, a new super learner for
right-censored data, which simultaneously estimates the expected loss
of learners of the event time distribution and the censoring
distribution. The loss function which is used to define the state
learner is based on the observable
quantities. % ANDERS: hvad mener vi med observable quantities? distribution of the observed random variables?
The state learner can be applied to all types of survival models,
works in the presence of competing risks, and does not require a
single pre-specified estimator of the conditional censoring
distribution.  To analyze the theoretical properties of the state
learner we focus on the discrete super learner which combines the
library of learners by picking the one that minimises the
cross-validated loss
\citep{van2011targeted}. % ANDERS: er der bedre referencer fra Mark paper/bog for discrete super learner?
In the presence of competing risks, our algorithm uses separate
libraries to model the hazard rates of the competing risks and the
censoring distribution. We show that the oracle selector of the state
learner is consistent if all libraries contain a consistent learner
and prove a finite sample oracle inequality. The result of the state
learner algorithm is a model which predicts the probability of an
event based on covariates in the presence of competing risks
\citep{gerds2021medical}. The other application is in targeted
learning where conditional event probabilities occur as
high-dimensional nuisance parameters which need to be estimated at a
certain rate \citep{van2011targeted, rytgaard2021estimation,
  rytgaard2022targeted}. We show for a targeted estimator, which uses
the state learner to estimate nuisance parameters, that a second order
product structure holds for the asymptotic bias term.

The article is organised as follows. We introduce our notation and
framework in Section~\ref{sec:framework}. In
Section~\ref{sec:super-learning} we define general super learning
for right-censored data, and in Section~\ref{sec:relat-liter-exist}
we review existing methods.  Section~\ref{sec:super-learner-simple}
introduces our novel proposal, the state learner, and
Section~\ref{sec:theor-results-prop} provides theoretical guarantees
for the state learner. In Section~\ref{sec:targeted-learning} we
discuss the use of the state learner in the context of targeted
learning. We report results of our numerical experiments in
Section~\ref{sec:numer-exper} and analyze a prostate cancer data set
in Section~\ref{sec:real-data-appl}.  Section~\ref{sec:discussion}
contains a discussion of the merits and limitations of our
proposal. Appendices~\ref{sec:proof-proposition}
and~\ref{sec:state-learner-with} contain proofs.

\section{Notation and framework}
\label{sec:framework}

In a competing risk framework \citep{andersen2012statistical}, let \( T\) be a
time to event variable, \(D\in\{1,2\}\) the cause of the event, and
$X \in \mathcal{X}$ a vector of baseline covariates taking values in a bounded
subset \( \mathcal{X} \subset \R^p \), \( p\in\N \). Let $\tau< \infty$ be the
maximal length of follow-up. We use \( \mathcal{Q} \) to denote the collection
of all probability measures on \( [0,\tau] \times \{1,2\}\times \mathcal{X} \)
such that \( (T, D, X) \sim Q \) for some unknown \( Q \in \mathcal{Q} \). For
\(j\in\{1,2\}\), the cause-specific conditional cumulative hazard functions are
defined by
\( \Lambda_{j} \colon [0, \tau] \times \mathcal{X} \rightarrow \R_+ \) such that
\begin{equation*}
  % \label{eq:cum-haz}
  \Lambda_{j}(t \mid x) = \int_0^t\frac{  Q(T \in \diff s, D=j \mid X=x )}{Q(T \geq s \mid X=x )}.
\end{equation*}
For ease of notation we assume throughout that \( \Lambda_j(\blank \mid x) \) is
continuous for all \( x \) and \( j \). %ANDERS: need a note here which says that this is not a limitation?
We denote by \(S\) the conditional
event-free survival function:
\begin{equation}
  \label{eq:surv-def}
  S(t \mid x)=\exp\left\{-\Lambda_{1}(t \mid x)-\Lambda_{2}(t \mid x)\right\}.
\end{equation}
% ANDERS: I prefer to have the \tau explicit
Let \( \mathcal{M}_{\tau}\) denote the space of all conditional cumulative hazard
functions on \( [0,\tau] \times\mathcal{X}\). Any distribution
\( Q \in \mathcal{Q} \) can be characterised by
\begin{equation*}
  \label{eq:parametrizeQ}
  \begin{split}
    Q(\diff t,j,\diff x)&= \left\{S(t- \mid x)\Lambda_1(\diff t \mid x)H(\diff x)\right\}^{1_{\{j=1\}}}\\
                        &\quad\left\{S(t- \mid x)\Lambda_2(\diff t \mid x)H(\diff x)\right\}^{1_{\{j=2\}}},
  \end{split}
\end{equation*}
where \(\Lambda_{j} \in \mathcal{M}_{\tau}\) for \(j=1,2\) and \(H\) is the marginal
distribution of the covariates.

We consider the right-censored setting in which we observe data
\(O = (\tilde{T},\tilde D, X)\), where $\tilde T = \min(T,C)$ for a
right-censoring time \(C\), $\Delta = \1{\{T \leq C\}}$, and
\(\tilde D=\Delta D\). Let \(\mathcal{P}\) denote a set of probability
measures on the sample space
\(\mathcal{O} = [0, \tau] \times \{0, 1, 2\} \times \mathcal{X}\) such
that \(O \sim P \) for some unknown \(P\in \mathcal{P}\). We assume
that the event times and the censoring times are conditionally
independent given covariates, \( T \independent C \mid X \). This
implies that any distribution \( P \in \mathcal{P} \) is characterised
by a distribution \( Q \in \mathcal{Q} \) and a conditional cumulative
hazard function for \( C \) given \( X \)
\citep[c.f.,][]{begun1983information,gill1997coarsening}. We use
\(\Gamma\in\mathcal M\) to denote the conditional cumulative hazard
function for censoring. % ANDERS: can we not allow Gamma to be discrete?
For notational convenience, we assume that \( \Gamma(\blank \mid x) \) is
continuous for all \( x \), and let
\(G(t \mid x)=\exp\left\{-\Gamma(t \mid x)\right\}\) denote the
survival function of the conditional censoring distribution. The distribution 
is characterized by
\begin{equation}\label{eq:parametrizeP}
  \begin{split}
    P(\diff t, j, \diff x) &= \left\{G(t- \mid x)S(t- \mid x)\Lambda_1(\diff t \mid x)H(\diff x)\right\}^{\1{{\{j=1\}}}}\\
                           &\quad\left\{G(t- \mid x)S(t- \mid x)\Lambda_2(\diff t \mid x)H(\diff x)\right\}^{\1{{\{j=2\}}}}\\
                           &\quad\left\{G(t- \mid x)S(t- \mid x)\Gamma(\diff t \mid x)H(\diff x)\right\}^{\1{{\{j=0\}}}}\\
                           &=\left\{G(t- \mid x)Q(\diff t,j,\diff x)\right\}^{\1{{\{j\ne 0\}}}}\\    
                           &\quad\left\{G(t- \mid x)S(t- \mid x)\Gamma(\diff t \mid x)H(\diff x)\right\}^{\1{{\{j=0\}}}}.
  \end{split}
\end{equation}
% ANDERS: H is missing in the model?
Hence, we may write
\( \mathcal{P} = \{ P_{Q, \Gamma} : Q \in \mathcal{Q}, \Gamma \in
\mathcal{G} \} \) for some \( \mathcal{G} \subset \mathcal{M}_{\tau} \). We
also have
\begin{equation*}
P(\tilde T>t \mid X=x) = S(t \mid x)G(t \mid x) = \exp\left\{-\Lambda_{1}(t \mid x)-\Lambda_{2}(t \mid x)-\Gamma(t \mid x) \right\}.
\end{equation*}
We further assume that there exists \(\kappa<\infty\) such that
\(\Lambda_{j}(\tau- \mid x)<\kappa \), for \(j\in\{1,2\}\), and
\(\Gamma(\tau- \mid x)<\kappa\) for almost all \(x\in\mathcal X\). Note that this
implies that \(G(\tau- \mid x)\) is bounded away from zero for almost all \(x\in\mathcal X\).
Under these assumptions, the conditional cumulative hazard functions
\(\Lambda_{j}\) and \(\Gamma\) can be identified from \(P\) by
\begin{align}
  \Lambda_{j}(t \mid x) &= \int_0^t\frac{  P(\tilde T \in \diff s, \tilde D=j \mid X=x )}{P(\tilde T \geq s \mid X=x )}, \label{eq:lambdaj}\\
  \Gamma(t \mid x) &= \int_0^t\frac{  P(\tilde T \in \diff s, \tilde D=0 \mid X=x )}{P(\tilde T \geq s \mid X=x )}\label{eq:gamma}.
\end{align}
Thus, we can consider $\Lambda_j$ and \(\Gamma\) as operators which map from
\( \mathcal{P} \) to \(\mathcal M\).

\section{The concept of super learning}
\label{sec:super-learning}

In survival analysis, a super learner can be used to estimate a
parameter $\Psi$ which can be identified from the observed data
distribution \(P\in\mathcal P\). In this section, to introduce the
discrete super learner and the oracle learner, we consider estimation of the
function-valued parameter \(\Psi:\mathcal P\to\mathcal M\), given by
\(\Psi(P)=\Lambda_j\). This parameter is identified via equation
\eqref{eq:lambdaj} on \([0,\tau]\).

As input to the super learner we need a data set
\( \data_n=\{O_i\}_{i=1}^n \) of i.i.d.\ observations from \( P \in \mathcal{P} \) and a finite collection of candidate
learners $\mathcal{A}$. Each learner \(a \in \mathcal{A}\) is a map
\( a \colon \mathcal{O}^n \rightarrow \mathcal{M}_{\tau}\) which takes a data
set as input and returns an estimate $a(\data_n) \in \mathcal{M}_{\tau}$ of
$\Lambda_{j}$.
% Formally, the domain of $a$ depends on \(n\) but we suppress this in the notation.
In what follows, we use the short-hand notation
\(P[f] = \int f(o) P(\diff o) \). A super learner evaluates the
performance of \(a \in \mathcal{A}\) with a loss function
\(L\colon \mathcal{M}_{\tau} \times \mathcal{O} \rightarrow \R_+\) and
estimates the expected loss \(P[L(a(\data_n), \blank)]\) using
cross-validation. Specifically, the expected loss of $a\in\mathcal A$
is estimated by splitting the data set $\data_n$ into $K$ disjoint
approximately equally sized subsets
\(\data_n^1, \data_n^2, \dots, \data_n^K \) and then calculating the
cross-validated loss
\begin{equation*}
  % \label{eq:cv-risk-est}
  \hat{R}_n(a; L) =
  \frac{1}{K}\sum_{k=1}^{K}
  % \empmeas^k{[L {(a{ (\data_n^{-k})} , \blank) }]},
  \frac{1}{| \data_n^{k} |}\sum_{O_i \in \data_n^{k}}
  L
  {
    \left(
      a{ (\data_n^{-k})}
      , O_i
    \right)
  },
  \quad \text{with} \quad
  \data_n^{-k} = \data_n \setminus \data_n^{k}.
\end{equation*}
% where \( \empmeas^{k} \) is the empirical measure of \( \data_n^{k} \). 
The subset \(\data_n^{-k}\) is referred to as the \(k\)'th training
sample, while \(\data_n^{k}\) is referred to as the \(k\)'th test or
hold-out sample.
The discrete super learner is defined as
\begin{equation*}
\hat{a}_n = \argmin_{a\in\mathcal A}\hat{R}_n(a; L).
\end{equation*}
% The final estimator of \(\Psi(P)=\Lambda_j\) is then the selected
% learner applied to the full data set, i.e., \(\hat{a}_n(\data_n)\).
The oracle learner is defined as the learner that minimises the
average loss according to the data-generating distribution \( P \),
i.e.,
\begin{equation*}
  \tilde{a}_n =
  \argmin_{a \in \mathcal{A}}
  \tilde{R}_n(a; L),
  \quad \text{with} \quad 
  \tilde{R}_n(a; L)=
  \frac{1}{K}\sum_{k=1}^{K} 
  P{
    \left[
      L
      {
        \left(
          a{ (\data_n^{-k})}
          , \blank
        \right)
      }
    \right]}
  .
\end{equation*}
Note that both the discrete super learner and the oracle learner
depend on the library of learners and on the number of folds \(K\),
and that the oracle learner is a function of the data and the unknown
data-generating distribution. However, these dependencies are
suppressed in the notation.

\section{Existing methods}
\label{sec:relat-liter-exist}

Machine learning based on right-censored data commonly use the
negative partial log-likelihood as loss function
\citep[e.g.,][]{li2016regularized,yao2017deep,lee2018deephit,katzman2018deepsurv,gensheimer2019scalable,lee2021boosted,kvamme2021continuous}.
It seems however that this loss function is not perfectly suited for
super learning, because it cannot be applied to important survival
learners: The Kaplan-Meier estimator, the random survival forest, and
the semi-parametric Cox regression model provide cumulative hazard
functions that are piece-wise constant in the time argument, and hence
assign zero probability to event times not observed in the training
data. This implies that when event times are observed in continuous
time, any of these learners will almost surely have infinite loss in
an independent test set according to the negative partial
log-likelihood loss.
% ANDERS: would be good to have a formula for a the negative partial log-likelihood loss showing that the result is infitity

When a proportional hazards model is assumed, the baseline hazard
function can be profiled out of the likelihood to give a new partial
log-likelihood loss \citep{cox1972regression}, which has been
suggested as a loss function for super learning
\citep{golmakani2020super,verweij1993cross}. While this allows the
library of learners to include Cox' proportional hazard models, the
drawback is that the library is now restricted to include only these
models. % ANDERS: can the library really include multiple proportional hazards models? only one of them can be correct ...
% also, would be good to merge this paragraph with the previous somehow

Another approach for super learning with right-censored data is to use an
inverse probability of censoring weighted (IPCW) loss function
\citep{graf1999assessment,van2003unicv,molinaro2004tree,keles2004asymptotically,hothorn2006survival,gerds2006consistent,gonzalez2021stacked}.
An IPCW loss function is attractive because the associated risk does not depend
on the censoring distribution but describes a feature of the population of
interested governed by the measure \( Q \in \mathcal{Q} \). Similar results can
be obtained using censoring unbiased transformations
\citep{fan1996local,steingrimsson2019censoring} or pseudo-values
\citep{andersen2003generalised,mogensen2013random,sachs2019ensemble}. All these
methods rely on an estimator of the censoring distribution, and their drawback
is that this estimator has to be pre-specified. When the data-generating
mechanism is complex and not well-understood, pre-specification of the censoring
distribution is a challenge.

An attempt to avoid a pre-specified censoring model was proposed
independently by \cite{han2021inverse} and
\cite{westling2021inference}. In both articles, the authors suggest to
iterate between learning of \( \Lambda \) and learning of $\Gamma$
using IPCW loss functions. However, no general theoretical guarantees
seem to exist for this procedure, and it has not yet been extended to
the situation with competing risks.

\section{The state learner}
\label{sec:super-learner-simple}

% The problem with most existing super learners for right-censored data is that
% they depend on a pre-specified estimator of the censoring distribution.
The main idea of the state learner is to jointly use learners of
\( \Lambda_1 \), \( \Lambda_2 \), and \( \Gamma \), and the relations
in equation~(\ref{eq:parametrizeP}), to learn a feature of the
observed data distribution \( P \). The discrete state learner ranks a
tuple of learners of \( (\Lambda_1, \Lambda_2, \Gamma) \) based on how
well they jointly model the observed data. %ANDERS: would be good to add one sentence about how to obtain a prediction model from such a ranking.
To formally introduce the
state learner, we define the multi-state process
\begin{equation*}
  \eta(t) = \1\{\tilde{T} \leq t, \tilde D=1\} + 2\,\1\{\tilde{T} \leq t, \tilde
  D=2\} - \1\{\tilde{T} \leq t, \tilde D=0\},
  \quad \text{for} \quad t \in [0, \tau].
\end{equation*}
At time \(t\), we observe that each individual is in one of four
mutually exclusive states (Figure \ref{fig:multi-state-process}).
\begin{figure}[h]
  \centering %
  \includeFigCond[width=.5\textwidth]{figure-multi-state-process.pdf}
  \caption{Illustration of the multi-state process \(\eta\) used by
    the state learner. Note that `censored' is a state, hence the
    process is always observed at any time.}
  \label{fig:multi-state-process}
\end{figure}
% ANDERS: Figuren er lavet med prodlim, ik? du kan angive nummer for boksene (0 = initial, 1 = cause 1, 2= cause 2 og jeg ville foretraekke 3 i stedet for -1 for censored)
The conditional distribution of \( \eta(t) \) given \( X \) is determined by
the function
\begin{equation}
  \label{eq:F-def}
  F(t, k, x) = P(\eta(t) = k \mid X=x),
  \quad \text{for all} \quad
  t \in [0,\tau],
  k \in \{-1,0,1,2\},
  x \in \mathcal{X}.
\end{equation}
The function \( F \) describes the conditional state occupation
probabilities of the multi-state process \(\eta\). We construct a
super learner for \( F \). The target of this super learner is the
function-valued parameter $\Psi(P) = F$ which is identified through
equation~(\ref{eq:F-def}).
% ANDERS: the next sentence requires conditional independent censoring and restriction to [0,\tau], right?
Because each quadruple
$(\Lambda_{1}, \Lambda_{2}, \Gamma, H)$ characterizes a distribution
\(P\in\mathcal P\) which in turn determines \( (F, H) \),
% ANDERS: and we should refer to equations (2) and (6) 
a learner for \( F \) can be constructed from learners of \( \Lambda_1 \),
\( \Lambda_2 \), and $\Gamma$ as follows:
\begin{equation}\label{eq:transition}
  \begin{split}
  F(t, 1, x)
  & = P(\tilde{T} \leq t, \Delta=1 \mid X=x)
    = \int_0^t e^{\{-\Lambda_{1}(s \mid x)-\Lambda_{2}(s \mid x) - \Gamma(s \mid x)\} }  \Lambda_{1}(\diff s \mid x),
  \\
  F(t, 2, x)
  & = P(\tilde{T} \leq t, \Delta=2 \mid X=x)
    = \int_0^t e^{\{-\Lambda_{1}(s \mid x)-\Lambda_{2}(s \mid x) - \Gamma(s \mid x)\} }  \Lambda_{2}(\diff s \mid x),
  \\
  F(t, -1, x)
  & =
    P(\tilde{T} \leq t, \Delta=0 \mid X=x)
    = \int_0^t e^{\{-\Lambda_{1}(s \mid x)-\Lambda_{2}(s \mid x) - \Gamma(s \mid x)\} }  \Gamma(\diff s \mid x),
  \\
  F(t, 0, x)
  &
    = P(\tilde{T} > t \mid X= x)
    = 1- F(t, 1, x) - F(t, 2, x)- F(t, -1, x).
  \end{split}
\end{equation}
The state learner requires three libraries of learners,
\(\mathcal{A}_1\), \( \mathcal{A}_2 \), and \( \mathcal{B} \), where
\(\mathcal{A}_1\) and \( \mathcal{A}_2\) contain learners of the
conditional cause-specific cumulative hazard functions \(\Lambda_1\)
and \( \Lambda_2\), respectively, and \(\mathcal{B}\) contains
learners of the conditional cumulative hazard function of the
censoring distribution. % We further
% define \(\mathcal{H}\) as the set of all probability distributions on
% \( \mathcal{X} \) and
% \begin{equation*}
%   \mathcal{H}_n=\left\{h\colon\mathcal{O}^n\longrightarrow\mathcal{H} \midd
%     h(\data_n)=H_n = \frac 1 n \sum_{i=1}^n \delta_{X_i} \right\}
% \end{equation*}
% as the library of learners which consists of a single learner, the empirical
% distribution function.
Based on the Cartesian product of
libraries of learners for \((\Lambda_1,\Lambda_2,\Gamma)\) we construct a library
$\mathcal{F}$ of learners
for \( F \):
\begin{align*}
  \mathcal{F}(\mathcal{A}_1, \mathcal{A}_2, \mathcal{B})
  &= \{ \phi_{a_1,a_2, b} : a_1 \in \mathcal{A}_1, a_2 \in \mathcal{A}_2, b \in \mathcal{B}\},
    \intertext{where in correspondance with  the relations in equation \eqref{eq:transition},} 
    \phi_{a_1,a_2, b}(\data_n)(t,1,x) &= \int_0^t e^{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  a_1(\data_n)(\diff s \mid x),\\
  \phi_{a_1,a_2, b}(\data_n)(t,2,x) &= \int_0^t e^{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  a_2(\data_n)(\diff s \mid x),\\
  \phi_{a_1,a_2, b}(\data_n)(t,-1,x) &= \int_0^t e^{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  b(\data_n)(\diff s \mid x),\\
  \phi_{a_1,a_2, b}(\data_n)(t,0,x) &= 1
                                      - \phi_{a_1,a_2, b}(\data_n)(t,1,x)
                                      - \phi_{a_1,a_2, b}(\data_n)(t,2,x)
  \\
  & \qquad
    -\phi_{a_1,a_2, b}(\data_n)(t,-1,x)
\end{align*}
To evaluate how well a function \( F \) predicts the observed
multi-state process we use the integrated Brier score
\( \bar B_\tau( F,O) = \int_0^{\tau} B_t(F,O) \diff t \), where \( B_t \) is the
Brier score \citep{brier1950verification} at time \( t \in [0, \tau] \),
\begin{equation*}
  B_t(F,O) = \sum_{j=-1}^{2}
  \left(
      F(t,j,X) - \1{\{\eta(t)=j\}}
  \right)^2.
\end{equation*}
Based on a split of a data set \(\data_n\) into $K$ disjoint
approximately equally sized subsets (see Section \ref{sec:super-learning}), each learner
\( \phi_{a_1, a_2, b} \) in the library
\( \mathcal{F}(\mathcal{A}_1, \mathcal{A}_2, \mathcal{B}) \) is
evaluated using the cross-validated loss,
\begin{equation*}
  \hat{R}_{n}(\phi_{a_1,a_2,b} ; \bar{B}_{\tau}) =
  \frac{1}{K}\sum_{k=1}^{K}
  \frac{1}{| \data_n^{k} |}\sum_{O_i \in \data_n^{k}}
  \bar B_\tau
  {
    \left(
      \phi_{a_1,a_2,b}{ (\data_n^{-k})}
      , O_i
    \right)
  },
\end{equation*}
and the discrete state learner is given by
\begin{align*}\label{eq:discrete-state-learner}
  \hat{\phi}_n
  &=  \argmin_{(a_1,a_2,b)\in \mathcal{A}_1\times\mathcal{A}_2\times\mathcal{B}}
    \hat{R}_{n}(\phi_{a_1,a_2,b} ; \bar{B}_{\tau}).
\end{align*}

\section{Theoretical results for the state learner}
\label{sec:theor-results-prop}

In this section we establish theoretical guarantees for the state learner.
% We let \( \E_P \) denote
% expectation under \( P \) and use \( \Theta \) to denote the
% collection of all conditional state-occupation probability functions.
Proposition~\ref{prop:stric-prop} can be derived from the fact that
the integrated Brier score (also called the continuous ranked
probability score) is a strictly proper scoring rule
\citep{gneiting2007strictly}. This implies that if we minimise the
average loss of the integrated Brier score, we recover the parameters
of the data-generating distribution. Specifically, this implies that
the oracle of a state learner is consistent if the library of learners
contains at least one learner that is consistent for estimation of
\( F \). Recall that the function \(F\) implicitly depends on the data
generating probability measure \(P\in\mathcal P\) but that this was
suppressed in the notation. We now make this dependence explicit by
writing \(F_0\) for the function which is obtained by substituting a
specific \(P_0\in\mathcal{P}\) for \(P\) in equation
\eqref{eq:transition}.

\begin{proposition}
  \label{prop:stric-prop}
  For \(P_0\in\mathcal{P}\) define
  \begin{equation*}
    F^* = \argmin_{F} P_0{[\bar{B}_\tau(F, \blank)]},
  % F^* = \argmin_{(\Lambda_1,\Lambda_2,\Gamma) \in \mathcal{M}_{\tau}\otimes\mathcal{M}_{\tau}\otimes\mathcal{M}_{\tau}} P_0{[\bar B_\tau(\phi_{(\Lambda_1,\Lambda_2,\Gamma)}, \blank)]},
  \end{equation*}
  %ANDERS: would be good to define the set over which F is minimized with a symbol rather than only words
where the minimum is taken over all \( F \), such that \( F \) is a conditional
state occupation probability function for some measure \( P \) as defined in equation~(\ref{eq:F-def}).
Then \( F^*(t,j,\blank) = F_0(t,j,\blank) \) \( H \)-almost surely for
any \( j\in \{-1,0,1,2\} \) and almost any \( t \in [0, \tau]\).
\end{proposition}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

We now establish a finite sample oracle result for the state learner.
Our Corollary~\ref{cor:oracle-prop} is in essence a special case of Theorem 2.3
in \cite{vaart2006oracle}. % ANDERS: is Theo 2.3 correct?
We assume
that we split the data into equally sized folds, and for simplicity of
presentation we take \( n \) to be such that
\( |\data_n^{-k}| = n/K \) with \( K \) fixed. We will allow the
number of learners to grow with \( n \) and write
\( \mathcal{F}_n=\mathcal{F}(\mathcal{A}_{1,n}, \mathcal{A}_{2,n},
\mathcal{B}_n)\) as short-hand notation and to emphasise the
dependence on \( n \).
% The discrete super $\hat{\phi}_n$ and the
% oracle $\tilde{\phi}_n$ are defined as in Section~\ref{sec:framework}
% but now with respect to the library \( \mathcal{F}_n \) and the loss
% \( B \).
In the following we let \( \| \blank \|_{P_0} \) denote the norm
% ANDERS: norm on which space? 
\begin{equation}
  \label{eq:norm}
  \| F \|_{P_0} = 
  \left\{
    \sum_{j=-1}^{2}\int_{\mathcal{X}} \int_0^{\tau} F(t, j, x)^2 \diff t H_0( \diff x)
  \right\}^{1/2}.
\end{equation}
\begin{corollary}
  \label{cor:oracle-prop}
  For all \(P_0\in\mathcal{P}\), \( n \in \N \), \( k \in \{1, \dots, K\} \),
  and $\delta>0$,
  \begin{align*}
    \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    & \leq (1 + 2\delta)
      \E_{P_0}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    \\
    & \quad
      + (1+ \delta) 16   K \tau
      \left(
      13 + \frac{12}{\delta}
      \right)
      \frac{\log(1 + |\mathcal{F}_n|)}{n}.
  \end{align*}
\end{corollary}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

Corollary~\ref{cor:oracle-prop} has the following asymptotic consequences.

\begin{corollary}
  \label{cor:asymp-cons}
  Assume that \( |\mathcal{F}_n| = \bigO(n^q)\), for some \( q \in \N \)
  and that there exists a sequence \( \phi_n \in \mathcal{F}_n \),
  \( n \in \N \), such that
  \( \E_{P_0}{\left[ \Vert \phi_n(\data_n^{-k}) - F_{0} \Vert_{P_0}^2 \right]} =
  \bigO(n^{-\alpha}) \), for some \( \alpha\leq 1 \).
  \begin{enumerate}[label=(\roman*), topsep=0pt]
  \item If $\alpha=1$ then
    \( \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2
      \right]} = \bigO(\log(n)n^{-1}) \).
  \item If $\alpha<1$ then
    \( \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]} =
    \bigO(n^{-\alpha}) \).
  \end{enumerate}
\end{corollary}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

% ANDERS: would be good to separate 2-3 subsection in this section. here is a vague proposal. above could be under \subsection{Oracle property of the state learner}
\subsection{Transience of double-robustness}
% In Section~\ref{sec:targeted-learning} we demonstrate how the state learner can
% be used for targeted learning.
A targeted estimator is typically obtained from
estimators of the nuisance parameters $\Lambda_1$, $\Lambda_2$, and $\Gamma$. By
equations~(\ref{eq:lambdaj}) and~(\ref{eq:gamma}) and the definition of \( F \),
we have
\begin{equation}
  \label{eq:7}
  \Gamma(t , x) 
  = \int_0^t  \frac{F(\diff s, -1, x )}{F(s-, 0, x )},
  \quad \text{and} \quad
  \Lambda_j(t , x) 
  = \int_0^t  \frac{F(\diff s, j, x )}{F(s-, 0, x )},
  \quad j \in \{1,2\},
\end{equation}
and thus a targeted estimator can also be obtained from an estimator of $F$
using equation~(\ref{eq:7}). The key feature of a targeted estimator
% ANDERS: at this point we have not defined "targeted estimator" does fisher 2021 define this? Bickel does not ...
is that it
is asymptotically equivalent to a sum of i.i.d.\ random variables plus a second
order remainder term \citep{bickel1993efficient,fisher2021visually}. For the
setting with competing risks, the remainder term is dominated by terms
of the form
\begin{equation}
  \label{eq:dr-term}
  P{\left[
      \int_0^{\tau} w_n(s, \blank)
      \hat{M}_{1,n}(s \mid  \blank)
      \hat{M}_{2,n}(\diff s \mid  \blank)
    \right]},
\end{equation}
where \( (\hat{M}_{1,n}, \hat{M}_{2,n}) \) is any of the nine combinations of
\( \hat{M}_{1,n} \in \{[\Gamma -\hat{\Gamma}_n], [\Lambda_1
-\hat{\Lambda}_{1,n}], [\Lambda_2 -\hat{\Lambda}_{2,n}]\} \) and
\( \hat{M}_{2,n} \in \{[\Gamma -\hat{\Gamma}_n], [\Lambda_1
-\hat{\Lambda}_{1,n}], [\Lambda_2 -\hat{\Lambda}_{2,n}]\} \), and \( w_n \) is
some data-dependent function with domain \([0,\tau]\times\mathcal X \)
\citep{van2003unified}. In particular, a targeted estimator will be
asymptotically linear if the `products' of the estimation errors
\( \hat{M}_{1,n} \) and \( \hat{M}_{2,n} \) in equation~(\ref{eq:dr-term}) are
\( \smallO_P{(n^{-1/2})}\). Proposition~\ref{prop:dr-structure} states that if
equation~(\ref{eq:dr-term}) holds for the a targeted estimator based on
estimators $\hat{\Lambda}_{1,n}$, $\hat{\Lambda}_{2,n}$, and $\hat{\Gamma}_{n}$,
then a similar product structure holds for a targeted estimator based on
\( \hat{F}_n \). We state the result for the special case that
\(\hat{M}_{1,n}= \Gamma-\hat{\Gamma}_n \) and
\(\hat{M}_{2,n} =\Lambda_1-\hat{\Lambda}_{1,n} \), but similar results holds for
any combinations of \( \Gamma-\hat{\Gamma}_n\),
\( \Lambda_1-\hat{\Lambda}_{1,n} \), and \( \Lambda_2-\hat{\Lambda}_{2,n} \).
\begin{proposition}
  \label{prop:dr-structure}
  Assume that \( w(s,x)\leq c \), \( F(s, 0, x) \geq 1/c \) and
  \( \hat{F}_n(s, 0, x) \geq 1/c \) for some \( c>0 \) for all
  \( s \in [0, \tau] \) and \( x \in \mathcal{X} \). Then there are real-valued
  uniformly bounded functions \( w^a_n \), \( w^b_n \), \( w^c_n \), and
  \( w^d_n \) with domain \( [0,\tau]^2 \times \mathcal{X} \) such that
  \begin{align*}
    & P_0{\left[
      \int_0^{\tau} w(s, \blank)
      \left\{
      \Gamma_0(s,\blank) -\hat{\Gamma}_n(s,\blank)
      \right\}
      [\Lambda_0-\hat{\Lambda}_n]
      (\diff s, \blank)
      \right]}
    \\
    & =
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^a_n(s,u,\blank) [F_0 - \hat{F}_n](u-, 0, \blank)[F_0 - \hat{F}_n](s-, 0, \blank) F_0(\diff u, 2, \blank ) F_0 ( \diff s, 1, \blank)
      \right]}
    \\
    & \quad +
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^b_n(s,u,\blank) [F_0 - \hat{F}_n](u-, 0, \blank)
      F_0(\diff u, 2, \blank ) [F_0 - \hat{F}_n](\diff s, 1, \blank)
      \right]}
    \\
    & \quad +
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^c_n(s,u,\blank) [F_0 - \hat{F}_n](\diff u, 2, \blank)
      [F_0 - \hat{F}_n](s-, 0, \blank)
      F_0(\diff s, 1, \blank ) 
      \right]}
    \\
    & \quad +
      P_0{\left[
      \int_0^{\tau} \int_0^{s} w^d_n(s,u,\blank) [F_0 - \hat{F}_n](\diff u, 2, \blank)
      [F_0 - \hat{F}_n](\diff s, 1, \blank)
      \right]}.
  \end{align*}
\end{proposition}
\begin{proof}
  See Appendix~\ref{sec:state-learner-with}.
\end{proof}

\section{Targeted learning}
\label{sec:targeted-learning}

We are interested in a parameter
\( \theta \colon \mathcal{Q} \rightarrow \Theta \) that expresses a
property of the uncensored population governed by the measure
\( Q \in \mathcal{Q} \). The parameter space $\Theta$ can be a subset
of \(\R^d\) or a subset of a function space, for example a subset of
\(\mathcal{M}_{\tau}\). In subsection~\ref{sec:cause-spec-aver} we
discuss an example from causal inference where $\theta$ is the average
treatment effect and \( \Theta = [-1,1] \).
% ANDERS: need a sentence a la: Before we discuss the role of the
% state learner for targeted learning we review some results from
% semiparametric efficiency theory. The reader may want to skip this ...
Under the assumption of
conditional independent censoring and positivity, $\theta$ is
identifiable from \( \mathcal{P} \) which means that there exists an
operator \( \Psi \colon \mathcal{P} \rightarrow \Theta \) such that
\( \theta(Q) = \Psi(P_{Q, \Gamma}) \) for all
$\Gamma \in \mathcal{M}_{\tau}$. By equation~(\ref{eq:parametrizeP})
this implies that we may write
\begin{equation*}
  \theta(Q) = \Psi(P) = \tilde{\Psi}^0(\Lambda_1, \Lambda_2, H),
\end{equation*}
for some operator \( \tilde{\Psi}^0 \). The state learner provides a
ranking of tuples
\( (a_1, a_2, b) \in \mathcal{A}_1 \times \mathcal{A}_2 \times
\mathcal{B} \), and we use \( \hat{a}_{1,n} \), \( \hat{a}_{2,n} \),
and \( \hat{b}_n \) to denote the learners corresponding to the
discrete state learner \( \hat{\phi}_n \). Letting \( H(\data_n) \)
denote the empirical measure of \( \{X_1, \dots, X_n\} \), we obtain a
simple plug-in estimator of $\theta$:
\begin{equation}
  \label{eq:2}
  \hat{\Psi}^0(\data_n) =
  \tilde{\Psi}^0(\hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n), H(\data_n)). 
\end{equation}

The asymptotic distribution of \( \hat{\Psi}_{\tau, j}^0 \) is
difficult to analyze due to the cross-validated model selection
involved in the estimation of the nuisance parameters $\Lambda_1$ and
$\Lambda_2$. In addition, the estimator will typically have an
asymptotic bias that vanishes at a too slow 
% ANDERS: is sub-optimal = too slow?
rate. Using tools from semi-parametric efficiency theory, it is
possible to construct a so-called targeted or debiased estimator with
smaller asymptotic bias and a asymptotic distribution which we know
how to estimate
\citep{bickel1993efficient,van2011targeted,chernozhukov2018double}. A
targeted estimator is based on the efficient influence function for
the parameter $\tilde{\Psi}^0$ and relies on an estimator of $\Gamma$
in addition to estimators of \( \Lambda_1 \) and $\Lambda_2$. The
efficient influence function is a \( P \)-zero mean and square
integrable function indexed by the nuisance parameters
\( (\Lambda_1, \Lambda_2, \Gamma) \), which we denote by
\( \psi(\blank ; \Lambda_1, \Lambda_2, \Gamma) \). The name is
justified because any regular asymptotically linear estimator that has
\( \psi \) as its influence function is asymptotically efficient,
meaning that it has smallest asymptotic variance among all regular
asymptotically linear estimators.

An example of a targeted estimator is the one-step estimator, defined as
\begin{equation}
    \label{eq:one-step-def}
    \hat{\Psi}_{\text{OS}}(\data_n)
    =
    \tilde{\Psi}^0(\hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
    H(\data_n))
    + \empmeas{[\psi(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
      \hat{b}_n(\data_n) )]},
\end{equation}
where \( \empmeas \) is the empirical measure of a data set
\(\{O_i\}_{i=1}^n\).  Under a Donsker class condition
% ANDERS: Donsker class condition on what set?
we have the
following asymptotic expansion of the one-step estimator
\citep{pfanzagl1985contributions,van2003unified,fisher2021visually,kennedy2022semiparametric,rytgaard2022continuous},
\begin{equation*}
  \hat{\Psi}_{\text{OS}}(\data_n)- \Psi(P)
  =  \empmeas{[\psi(\blank ; \Lambda_1, \Lambda_2, \Gamma)]}
  +\mathrm{Rem}{(\hat{\Lambda}_{1,n},\hat{\Lambda}_{2,n},  \hat{\Gamma}_n, P)} + \smallO_{P}(n^{-1/2}),
\end{equation*}
where the remainder term has the form
\begin{equation}
  \label{eq:4}
  \mathrm{Rem}{(\hat{\Lambda}_{1,n},\hat{\Lambda}_{2,n},  \hat{\Gamma}_n, P)}
  = \mathcal{O}_P{
    \left\{
      \|\Lambda_1-\hat{\Lambda}_{1,n}\|^2
      +
      \|\Lambda_2-\hat{\Lambda}_{2,n}\|^2
      +
      \|\Gamma-\hat{\Gamma}_{n}\|^2
    \right\}
  },
\end{equation}
for some suitable norm \( \|\blank \| \), for instance the
\( \mathcal{L}_{P}^2 \)-norm. When equation~(\ref{eq:4}) holds and the
nuisance parameters $\Lambda_1$, $\Lambda_2$, and $\Gamma$ are
consistently estimated at rate \( \smallO_P{(n^{-1/4})} \), then
\begin{equation}
  \label{eq:3}
  \sqrt{n}(\hat{\Psi}_{\text{OS}}(\data_n)- \Psi(P)) \weakly \mathcal{N}(0,
  P{[\psi(\blank; \Lambda_1, \Lambda_2, \Gamma)^2]}),
\end{equation}
where we use \( \weakly \) to denote weak convergence \citep{van2000asymptotic}.
In particular, equation~(\ref{eq:3}) and Slutsky's lemma imply that we can
obtain asymptotically valid \((1-\alpha)\cdot100\%\) confidence intervals by
calculating
\begin{equation*}
  \left[
    \hat{\Psi}_{\text{OS}}(\data_n) - q_{\alpha/2} \hat{\sigma}(\data_n) ,
    \;
    \hat{\Psi}_{\text{OS}}(\data_n) + q_{\alpha/2} \hat{\sigma}(\data_n)
  \right],
\end{equation*}
where \( q_{\alpha} \) is the \( (1-\alpha) \)-quantile of the standard normal
distribution and
\( \hat{\sigma}(\data_n)^2 = \empmeas{ \left[ \psi(\blank;
    \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n), \hat{b}_n(\data_n))^2
  \right]} \).

\subsection{Average treatment effect on the absolute risk of an event}
\label{sec:cause-spec-aver}

In the following we demonstrate how the state learner and the general
estimation strategy outlined above can be used to construct an
estimator of the cause-specific average treatment effect. We assume
that the covariate vector \( X \in \R^p \) contains a binary treatment
indicator \( A \) and a vector of potential confounders,
\( W \in \mathcal{W}\subset \R^{p-1} \).
% ANDERS: why is this abuse of notation?
% We abuse notation slightly by writing
% \( \Lambda_j(t \mid x) = \Lambda_j(t \mid a, w) \) and
% \( S(t \mid x) = S(t \mid a, w) \) when \( x=(a, w) \).
We use $\mu$
to denote the marginal distribution of \( W \) and $\pi$ to denote the
conditional probability of treatment,
\begin{equation*}
  \pi(w) = P(A=1 \mid W=w).
\end{equation*}
We assume throughout that $\pi$ is uniformly bounded away from \( 0 \)
and \( 1 \) on \( \mathcal{W} \) and that both \( A \) and \( W \) are
fully observed for all individuals. We use an super learner to
estimate $\pi$ \citep{Polley_Ledell_Kennedy_Laan_2023_Superlearn}, and
we denote this estimator by $\hat{\pi}_n$. We use the empirical
measure of \( \{W_1, \dots, W_n\} \) to estimate $\mu$, and denote
this estimator by $\hat{\mu}_n$. As parameter of interest we consider
the standardised difference in the absolute risk of an event with
cause 1 at time $\tau$:
\begin{equation*}
  \theta_{{\tau}}(Q) = \int_{\mathcal{W}} 
  \left\{
    \int_0^{\tau}
    S(s- \mid w, 1)  \Lambda_1(\diff s \mid w, 1)
    -
    \int_0^{\tau}
    S(s- \mid w, 0)  \Lambda_1(\diff s \mid w, 0)
  \right\}
  \mu(\diff w).
\end{equation*}
Under a set of
% ANDERS: better to give these assumptions a name: positivity, ... and refer to Helene's paper?
additional assumptions, \( \theta_{{\tau}} \) can be given the causal
interpretation
\begin{equation*}
  \theta_{{\tau}}(Q) =
  P{(T^{A=1} \leq \tau, D^{A=1}=1)}-
  P{(T^{A=0} \leq \tau, D^{A=0}=1)},
\end{equation*}
where \( (T^A, D^A) \) denotes potential outcomes \citep{hernanRobinsWhatIf}. In
this case, the interpretation of $\theta_{\tau}$ is the difference in the
average risk of cause \( 1 \) occurring before time \( \tau \) in the population
if everyone had been given treatment (\( A=1 \)) compared to if no one had been
given treatment \( (A=0) \).

Using equation~(\ref{eq:surv-def}) we may write
\( \theta_{{\tau}}(Q) = \tilde{\Psi}_{t}^0(\Lambda_1, \Lambda_2, \mu) \), where
\begin{equation}
  \label{eq:1}    
  \begin{split}
  \tilde{\Psi}_{t}^0(\Lambda_1, \Lambda_2, \mu) & =
  \int_{\mathcal{W}} 
  \int_0^{\tau}
  e^{-\Lambda_1(s- \mid w, 1)-\Lambda_2(s- \mid w, 1)}  \Lambda_1(\diff s \mid
  w, 1)
  \mu(\diff w)
  \\
  &  \quad
  -\int_{\mathcal{W}} 
  \int_0^{\tau}
  e^{-\Lambda_1(s- \mid w, 0)-\Lambda_2(s- \mid w, 0)}  \Lambda_1(\diff s \mid w, 0)
  \mu(\diff w).
  \end{split}
\end{equation}
The efficient influence function for the parameter $\tilde{\Psi}_{\tau}$ depends
on the set \( (\Lambda_1, \Lambda_2, \Gamma, \pi) \) of nuisance parameters.
We define
\begin{align*}
  \omega_a(A,W; \pi)
  &=  \frac{% (-1)^{a+1}
    \1{\{A=a\}}}{\pi(W)^{a}(1-\pi(W))^{1-a}},
  \\
  g(t, A, W; \Lambda_1, \Lambda_2)
  & = \int_0^{t}
    e^{-\Lambda_1(s- \mid W, A)-\Lambda_2(s- \mid W, A)}  \Lambda_1(\diff s \mid
    W, A),
  \\  
  M_j(\diff t \mid A, W;  \Lambda_j  )
  & = N_j(\diff t) -
    \1{\{\tilde{T} \geq t\}} \Lambda_j(\diff t \mid W, A),
    \quad j \in \{1,2\},
  \intertext{and}
  M(\diff t \mid A, W;  \Lambda_1, \Lambda_2  )
  & = M_1(\diff t \mid A, W;  \Lambda_1  ) +
    M_2(\diff t \mid A, W;  \Lambda_2  ).
\end{align*}
The efficient influence function can now be written as
\citep{van2003unified,jewell2007non,rytgaard2022targeted},
\begin{equation*}
  \psi_{\tau}(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  = \psi_{\tau}^1(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  - \psi_{\tau}^0(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  -\tilde{\Psi}_{t}^0(\Lambda_1, \Lambda_2, \mu),
\end{equation*}
where
\begin{equation}
  \label{eq:5}
  \begin{split}
    & \psi_{\tau}^a(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
    \\
    & =
      \omega_a(A,W; \pi)
      \int_0^{\tau} e^{\Gamma(t- \mid A, W)}   
      M_1(\diff t \mid A, W; \Lambda_1)
    \\
    & \quad
      -
      \omega_a(A,W; \pi)
      g(\tau, A, W; \Lambda_1, \Lambda_2)
      \int_0^{\tau}
      e^{[\Gamma+\Lambda_1 + \Lambda_2](t- \mid A, W)}
      M(\diff t \mid A, W; \Lambda_1, \Lambda_2)
    \\
    & \quad
      +
      \omega_a(A,W; \pi)      
      \int_0^{\tau}
      g(t, A, W; \Lambda_1, \Lambda_2)
      e^{[\Gamma+\Lambda_1 + \Lambda_2](t- \mid A, W)}
      M(\diff t \mid A, W; \Lambda_1, \Lambda_2)
    \\
    & \quad + g(\tau, a, W; \Lambda_1, \Lambda_2).
  \end{split}
\end{equation}

Equations~(\ref{eq:1}) and~(\ref{eq:5}) allow us to construct a one-step
estimator by using the definition given in equation~(\ref{eq:one-step-def}),
which gives the estimator
\begin{equation}
  \label{eq:one-step-comp-ate}
  \begin{split}
    \hat{\Psi}_{t,\text{OS}}(\data_n)
    = &
        \tilde{\Psi}_{t}^0(\hat{a}_1(\data_n), \hat{a}_2(\data_n),
        \hat{\mu}_n(\data_n))
    \\
      &
        +
        \empmeas{[\psi_{\tau}(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}
    \\
    = &
        \empmeas{[\psi_{\tau}^1(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}
    \\
      &
        - \empmeas{[\psi_{\tau}^0(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}.
  \end{split}
\end{equation}

\section{Numerical experiments}
\label{sec:numer-exper}

As a proof of concept we conduct two small simulation studies in a setting
without competing risks. We generate data according to a distribution motivated
from a real dataset in which censoring depends on the baseline covariates. We
simulate data based on a prostate cancer study described in
\citep{kattan2000pretreatment}. The outcome of interest was the time to tumor
recurrence, and five baseline covariates were used to predict outcome:
prostate-specific antigen (PSA, ng/mL), Gleason score sum (GSS, values between 6
and 10), radiation dose (RD), hormone therapy (HT, yes/no) and clinical stage
(CS, six values). The study was designed such that a patient's radiation dose
depended on when the patient entered the study \citep{gerds2013estimating}. This
in turn implied that the time of censoring depended on the radiation dose. The
data were re-analyzed in \citep{gerds2013estimating} where a sensitivity
analysis was conducted based on simulated data. We use the same simulation
setup, where event and censoring times are generated according to parametric
Cox-Weibull models estimated from the original data, and the covariates are
generated according to either marginal Gaussian normal or binomial distributions
estimated from the original data
\citep[c.f.,][Section~4.6]{gerds2013estimating}. We refer to this simulation
setting as ``dependent censoring''. We also considered a simulation setting
where data were generated in the same way, except that censoring was generated
completely independently. We refer to this simulation setting as ``independent
censoring''.

For all super learners we use a library consisting of three learners: The
Kaplan-Meier estimator \citep{kaplan1958nonparametric,Gerds_2019prodlim}, a Cox
model with main effects \citep{cox1972regression, survival-package}, and a
random survival forest \citep{ishwaran2008random,rfsrc-paclage}. We use the same
library to learn the outcome and the censoring mechanism. To be precise, note
that given a learner for the cumulative hazard function of the outcome event, we
can typically use the same method to construct a learner of the cumulative
hazard function of the censoring distribution. This works by training the
learner on the data set \( \data_n^c \), where
\( \data_n^c = \{O_i^c\}_{i=1}^n \) with
\( O_i^c = (\tilde{T}_i, 1-\Delta, X_i) \). When we say that we use a learner
for the cumulative hazard function of the outcome to learn the cumulative hazard
function of the censoring time, we mean that the learner is trained on
\( \data_n^c \).

We first compare the state learner to two IPCW based super learners: The first
uses a Cox model with main effects to estimate the censoring probabilities
(IPCW(Cox)), while the second uses the Kaplan-Meier estimator to estimate the
censoring probabilities (IPCW(KM)). The Cox model for the censoring distribution
is correctly specified in both simulation settings while the Kaplan Meier
estimator only estimates the censoring model correctly in the simulation setting
where censoring is independent. Both IPCW super learners are fitted using the
\texttt{R}-package \texttt{riskRegression} \citep{Gerds_Ohlendorff_Ozenne_2023}.
Each super learner provides a learner for the cumulative hazard function for the
outcome of interest, and from this a risk prediction model can be obtained
(c.f., equation~(\ref{eq:surv-def}) with $\Lambda_2 = 0$). We measure
performance of each super learner by calculating the index of prediction
accuracy (IPA) \citep{kattan2018index} at a fixed time horizon (36 months) for
the risk prediction model provided by the super learner. The IPA is 1 minus the
ratio between the model's Brier score and the null model's Brier score, where
the null model is the model that does not use any covariate information. The IPA
is approximated using a large (\( n = 20,000 \)) independent data set of
uncensored data. As a benchmark we calculate the performance of the risk
prediction model chosen by the oracle selector. The oracle selector uses the
large data set of uncensored event times to select the learner with the highest
IPA.

The results are shown in Figure~\ref{fig:ipcw-fail}. We see that when censoring
is not independent, using the Kaplan-Meier estimator to estimate the censoring
probabilities provides a risk prediction model with an IPA that is lower than
the risk prediction model provided by the state learner. The performance of the
risk prediction model selected by the state learner is similar to the risk
prediction model selected by the IPCW(Cox) super learner which a priori uses a
correctly specified model for the censoring distribution. Both these risk
prediction models are close to the performance of the oracle, except for small
sample sizes.

\begin{figure}
  \centering %
  \includeFigCond[width=1\linewidth]{experiment-fig-sl-ipcw.pdf}
  \caption[]{For the risk prediction models provided by each of the super
    learners, the IPA is plotted against sample size. The results are based on
    1000 repetitions and the error bars are used to quantify the Monte Carlo
    uncertainty.}
  \label{fig:ipcw-fail}
\end{figure}

We next compare the state learner to the super learner \texttt{survSL}
\citep{westling2021inference}, as this is the only existing super learner we
know which works without a pre-specified censoring model. As both the state
learner and \texttt{survSL} also provides a prediction model for the probability
of being censored, we compare the performance of both the outcome and the
censoring modeling using the IPA.

The results are shown in Figure~\ref{fig:zelefski}. We see that the prediction
models selected for both censoring and outcome have similar or higher API
compared to the prediction models selected by \texttt{survSL} for most sample
sizes.

\begin{figure}
  \centering %
  \includeFigCond[width=1\linewidth]{experiment-fig-sl-survSL.pdf}
  \caption[]{For the risk prediction models of both the outcome and the
    censoring probability provided by each of the super learners, the IPA is
    plotted against sample size. The results are based on 1000 repetitions and
    the error bars are used to quantify the Monte Carlo uncertainty.}
  \label{fig:zelefski}
\end{figure}

\section{Example with prostate cancer study}
\label{sec:real-data-appl}

The prostate cancer data analyzed by \cite{kattan2000pretreatment}, which we
introduced in Section~\ref{sec:numer-exper}, include a competing event in the
form of death without tumor recurrence. For illustration, we fit the state
learner to the original data set consisting of 1,042 patients. We consider death
without tumor recurrence and recurrence of tumor as two competing events of
interest. We include the fives learner \texttt{KM}, \texttt{Cox strata CS},
\texttt{Lasso}, \texttt{Elastic}, and \texttt{RF} which are described in
Table~\ref{tab:zel-library}. We use the same library of learners to learn
\( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$. In this case, \( \Lambda_1 \)
denotes the cause-specific cumulative hazard function of tumor recurrence, and
\( \Lambda_2 \) denotes the cause-specific cumulative hazard function of death
without tumor recurrence.

\begin{table}
  \caption{\label{tab:zel-library}Overview of the five learners used by the
    state learner when applied to the prostate cancer data set. The Kaplan-Meier
    estimator was fitted using the package \texttt{prodlim}
    \citep{Gerds_2019prodlim}. All Cox models included all five covariates in
    the model and were fitted using the package \texttt{survival}
    \citep{survival-package}. All penalised Cox models included all five
    covariates as linear predictors and were fitted using the package
    \texttt{glmnet} \citep{glmnet-cox,glmnet-glm}. The random forest was fitted
    with the package \texttt{randomForestSRC} \citep{rfsrc-paclage}.}
  \centering
  \begin{tabular}{lll}
    \toprule
    Family & Model & Description \\
    \midrule
    Marginal & \texttt{KM} & The Kaplan-Meier estimator \\ 
    Cox & \texttt{Cox} & All five covariates included with additive effects \\
           & \( \texttt{Cox strata CS} \)  & Cox model stratified on CS  \\
           % & \( \texttt{Cox strata HT} \)  & Cox model stratified on HT \\
           % & \( \texttt{Cox spline} \)  & PSA and RD modeled with splines \\ 
    Penalised Cox & \texttt{Lasso} & Cox model with \( L_1 \)-norm penalty   \\
           % & \texttt{Ridge} & Cox model with \( L_2 \)-norm penalty \\
           & \texttt{Elastic} & Cox model with \( L_1 \)- and \( L_2 \)-norm penalty \\
    Random forest & \texttt{RF} & Random forest with default settings \\
    \bottomrule 
  \end{tabular}
\end{table}

This gives a library consisting of \( 5^3 = 125 \) learners for the conditional
state occupation probability function \( F \) defined in
equation~(\ref{eq:F-def}). We use five folds for training and testing the
models, and we repeat training and evaluation five times with different splits.
The integrated Brier score (defined in Section~\ref{sec:super-learner-simple})
for all learners are shown in Figure~\ref{fig:zelefski-real}, and the top 10
combinations of learners are displayed in Table~\ref{tab:zelefski-real}. We see
that the prediction performance is mostly affected by the choice of learner for
the censoring distribution. Several combinations of learners give similar
performance as measured by the integrated Brier score, as long as a random
forest is used to model the censoring distribution.

\begin{figure}
  \centering %
  \includeFigCond[width=1\linewidth]{zelefski-real-data.pdf}
  \caption[]{The results of applying the 125 combinations of learners to the
    prostate cancer data set. The learners are \texttt{KM} (KM), \texttt{Cox
      strata CS} (strata), \texttt{Lasso} (lasso), \texttt{Elastic} (elastic),
    and \texttt{RF} (RF) as described Table~\ref{tab:zel-library}. The error
    bars are based on five repetitions using different splits. We refer to
    learners of \( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$ as `Tumor
    learner', `Death learner', and `Censoring learner', respectively.}
  \label{fig:zelefski-real}
\end{figure}

\begin{table}
  \caption{\label{tab:zelefski-real}The 10 best performing models in terms of integrated Brier score. The
    reported standard errors are based on five repetitions using different
    splits. The models are described in Table~\ref{tab:zel-library}. We refer to
    learners of \( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$ as `Tumor
    learner', `Death learner', and `Censoring learner', respectively.}
  \centering
  \input{./zel-tab.tex}
\end{table}

As an illustration, we use the learners of the three cumulative hazard functions
selected by the state learner and the estimator defined in
Section~\ref{sec:cause-spec-aver}, equation~(\ref{eq:one-step-comp-ate}), to
estimate the average treatment effect of hormone therapy on risk of tumor
recurrence and death. The propensity score is estimated with a lasso model that
includes all levels of interaction. The results are shown in
Figure~\ref{fig:zelefski-real-target} for 6 month intervals after baseline with
point wise 95\% confidence intervals. We see that hormone therapy appears to
decrease the risk of tumor recurrence and increase the risk of death without
tumor recurrence, but that none of the estimated effects are statistically
significant from zero.

\begin{figure}
  \centering%
  \includeFigCond[width=1\linewidth]{zelefsky-data-target-par.pdf}
  \caption[]{Estimates of the average treatment effect of hormone therapy on the
    risk of tumor recurrence and death obtained from the prostate cancer study
    analyzed by \cite{kattan2000pretreatment}. The estimates are based on the
    estimator defined in equation~(\ref{eq:one-step-comp-ate}) with point wise
    95\% confidence intervals calculated as described in
    Section~\ref{sec:targeted-learning}. The estimates of the nuisance
    parameters are provided by the state learner.}
  \label{fig:zelefski-real-target}
\end{figure}

\section{Discussion}
\label{sec:discussion}

We have proposed a new super learner that can be used with right-censored data
and competing events. Compared to existing methods, the advantage of the state
learner is that it does not depend on a pre-specified estimator of the censoring
distribution, but selects one automatically based on the provided library
\( \mathcal{B} \). Furthermore, the state learner neither requires evaluation of
the density of \( \Lambda_j \) nor assumes a particular semi-parametric
structure for $\Lambda_j$. In the remainder of this section we point to some
limitations of our proposal and discuss avenues for further research.

A major advantage of the state learner is that performance of each combination
of learners is measured in terms of observable quantities. This means that no
additional nuisance parameters need to be estimated to evaluate the loss. The
drawback with this approach is that we are rarely interested in features of the
observed data distribution when the data are right-censored. The finite sample
oracle inequality in Corollary~\ref{cor:oracle-prop} concerns the function
\( F \), which is a feature of \( P \in \mathcal{P} \), while what we are
typically interested in is \( \Lambda_j \) or \( S \), which are features of
\( Q \in \mathcal{Q} \). We emphasise that while the state learner provides us
with estimates of \( \Lambda_j \) and $\Gamma$ based on libraries
\( \mathcal{A}_j \) and \( \mathcal{B} \), performance are not assessed directly
for these parameters, but only jointly for estimation of the parameter \( F \).
For settings without a competing risk, our numerical studies suggest that
measuring performance with respect to estimation of \( F \) also leads to good
performance for estimation of \( S \). Further research on this topic, both
numerical and theoretical, is warranted.

% In the context of targeted learning, a drawback of the state learner is that the
% doubly robustness property defined in equation~(\ref{eq:dr-property}) seem to be
% lost because we only use a single nuisance parameter estimator. As defined here,
% however, the state learner is build using libraries for the conditional
% cause-specific cumulative hazard functions, so some doubly robustness might be
% preserved.

Our proposed super learner can be implemented with a broad library of learners
using existing software, for instance the \texttt{R}-package
\texttt{riskRegression} \citep{Gerds_Ohlendorff_Ozenne_2023}. Furthermore, while
the library \( \mathcal{F}(\mathcal{A}_1,\mathcal{A}_2,\mathcal{B}) \) consists
of \( |\mathcal{A}_1||\mathcal{A}_2||\mathcal{B}| \) learners, as long as we
have sufficient memory we need only fit
\( |\mathcal{A}_1| +|\mathcal{A}_2| + |\mathcal{B}| \) learners in each fold. To
evaluate the performance of each learner we need to perform
\( |\mathcal{A}_1||\mathcal{A}_2||\mathcal{B}| \) operations to calculate the
integrated Brier score in each hold-out sample, one for each combination of the
fitted models, but these operations are often negligible compared to fitting the
models. Hence the state learner is essentially no more computationally demanding
than any procedure that uses super learning to learn $\Lambda_1$, $\Lambda_2$,
and $\Gamma$ separately. While our proposal is based on constructing the library
\( \mathcal{F} \) from libraries for learning \( \Lambda_1 \), $\Lambda_2$, and
$\Gamma$, it would also be of interest to consider learners that estimate
\( F \) directly.

In our numerical studies, we only considered learners of $\Lambda_j$ and
$\Gamma$ that provide cumulative hazard functions which are piece-wise constant
in the time argument. This simplifies the calculation of \( F \) as the
integrals in equation~(\ref{eq:transition}) reduce to sums. When $\Lambda_j$ or
\( \Gamma \) are absolutely continuous in the time argument, calculating \( F \)
is more involved, but we expect that a good approximation can be achieved by
discretisation. In the future, we intend to investigate the performance of the
state learner when using a broader library of learners in more comprehensive
simulation studies.

\appendix

\section{Theoretical guarantees for the state learner}
\label{sec:proof-proposition}

In this section we provide proofs of the results stated in
Section~\ref{sec:theor-results-prop}.

Define
\( \bar{B}_{\tau,0}(F, o) = \bar{B}_{\tau}(F, o) - \bar{B}_{\tau}(F_0, o) \) and
\( R_{0}(F) = P_0{[\bar{B}_{\tau,0}(F, \blank)]} \).
\begin{lemma}
  \label{lemma:norm}
  \( R_{0}(F) = \Vert F - F_0 \Vert_{P_0}^2 \), where \( \Vert \blank \Vert_{P_0}\) is defined
  in equation~(\ref{eq:norm}).
\end{lemma}
\begin{proof}
  For any \( t \in [0, \tau] \) and \( j\in \{-1,0,1,2\} \) we have
  \begin{align*}
    & \E_{P_0}{\left[ (F(t, j, X) - \1{\{\eta(t) = j \}})^2 \right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, j, X) - F_0(t, j, X) + F_0(t, j, X) - \1{\{\eta(t) = j
      \}})^2 \right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, j, X) - F_0(t, j, X))^2\right]}
      + \E_{P_0}{\left[ (F_0(t, j, X) - \1{\{\eta(t) = j \}})^2\right]}
    \\
    & \quad
      + 2\E_{P_0}{\left[ (F(t, j, X) - F_0(t, j, X))(F_0(t, j, X) - \1{\{\eta(t) = j
      \}})\right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, j, X) - F_0(t, j, X))^2\right]}
      + \E_{P_0}{\left[ (F_0(t, j, X) - \1{\{\eta(t) = j \}})^2\right]},
  \end{align*}
  where the last equality follows from the tower property. Hence, using Fubini,
  we have
  \begin{equation*}
    P{[\bar{B}_{\tau}(F, \blank)]}
    = \Vert F - F_0 \Vert_{P_0}^2 + P_0{[\bar{B}_{\tau}(F_0, \blank)]}.
  \end{equation*}
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:stric-prop}]
  The result follows from Lemma~\ref{lemma:norm}.
\end{proof}

In the following, let $\Theta$ denote the function space consisting of all
conditional state occupation probability functions for some measure \( P \) as
defined in equation~(\ref{eq:F-def}).

\begin{proof}[Proof of Corollary~\ref{cor:oracle-prop}]
  First note that minimising the loss \( \bar{B}_{\tau} \) is equivalent to
  minimising the loss \( \bar{B}_{\tau,0} \), so the discrete super learner and
  oracle according to \( \bar{B}_{\tau} \) and \( \bar{B}_{\tau,0} \) are
  identical. By Lemma~\ref{lemma:norm}, \( R_0(F) \geq 0 \) for any
  \( F \in \Theta \), and so using Theorem 2.3 from \citep{vaart2006oracle} with
  \( p=1 \), we have that for all \( \delta >0 \),
\begin{align*}
  \E_{P_0}{\left[ R_0(\hat{\phi}_n(\data_n^{-k})) \right]}
  \leq
  &(1+2\delta)\E_{P_0}{\left[ R_0(\tilde{\phi}_n(\data_n^{-k})) \right]}
  \\
  & \quad + (1+\delta) \frac{16 K}{n}
    \log(1 + |\mathcal{F}_n|)\sup_{F \in \Theta}
    \left\{
    M(F) + \frac{v(F)}{R_0(F)}
    \left(
    \frac{1}{\delta} + 1
    \right)
    \right\}
\end{align*}
where for each \( F \in \Theta \), \( (M(F), v(F)) \) is some Bernstein pair for
the function \(o \mapsto \bar{B}_{\tau,0}(F, o) \). As
\( \bar{B}_{\tau,0}(F, \blank) \) is uniformly bounded by \( \tau \) for any
\( F \in \Theta \), it follows from section 8.1 in \citep{vaart2006oracle} that
\( (\tau, 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}) \) is a Bernstein pair for
\( \bar{B}_{\tau,0}(F, \blank) \). Now, for any \( a,b,c \in \R \) we have
\begin{align*}
  (a-c)^2 - (b-c)^2
  & = (a-b+b-c)^2 - (b-c)^2
  \\
  & = (a-b)^2 + (b-c)^2 +2(b-c)(a-b) - (b-c)^2
  \\
  & = (a-b)
    \left\{
    (a-b) +  2(b-c)
    \right\}
  \\
  & = (a-b)
    \left\{
     a + b -2c
    \right\},
\end{align*}
so using this with \( a=F(t, j, x) \), \( b=F_0(t, j, x) \), and
\( c = \1{\{\eta(t) = j\}} \), we have by Jensen's inequality
\begin{align*}
  & P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}
  \\
  & \leq
    2\tau\E_{P_0}{\left[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left\{
    \left(
    F(t, j, X) - \1{\{\eta(t) = j\}}
    \right)^2
    -
    \left(
    F_0(t, j, X) - \1{\{\eta(t) = j\}}
    \right)^2
    \right\}^2
    \diff t 
    \right]}
  \\
  & =2\tau
    \E_{P_0}\Bigg[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left(
    F(t, j, X) - F_0(t, j, X)
    \right)^2
  \\
  & \quad \quad \quad\quad \quad \quad \times
    \left\{
    F(t, j, X) +  F_0(t, j, X)-2 \1{\{\eta(t) = j\}}
    \right\}^2
    \diff t 
    \Bigg]
  \\
  & \leq
    8\tau \E_{P_0}{\left[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left(
    F(t, j, X) - F_0(t, j, X)
    \right)^2
    \diff t 
    \right]}.
  \\
  & =
    8\tau \Vert F - F_0 \Vert_{P_0}^2.
\end{align*}
Thus when \( v(F) = 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]} \) we have by
Lemma~\ref{lemma:norm}
\begin{equation*}
  \frac{v(F)}{R_0(F)}
  = 1.5 \frac{P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}}{P_0{[\bar{B}_{\tau,0}(F, \blank)]}}
  \leq 12 \tau,
\end{equation*}
and so using the Bernstein pairs \( (\tau, 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}) \) we have
\begin{equation*}
  \sup_{F \in \Theta}
  \left\{
    M(F) + \frac{v(F)}{R_0(F)}
    \left(
      \frac{1}{\delta} + 1
    \right)
  \right\}
  \leq \tau
  \left(
    13 + \frac{12}{\delta}
  \right),
\end{equation*}
For all $\delta>0$ we thus have
\begin{align*}
  \E_{P_0}{\left[ R_0(\hat{\phi}_n(\data_n^{-k})) \right]}
  \leq
  &(1+2\delta)\E_{P_0}{\left[ R_0(\tilde{\phi}_n(\data_n^{-k})) \right]}
  \\
  & \quad
    + (1+\delta)\log(1 + |\mathcal{F}_n|) \tau \frac{16 K}{n}
    \left(
    13 + \frac{12}{\delta}
    \right),
\end{align*}
and then the final result follows from Lemma~\ref{lemma:norm}.
\end{proof}

\begin{proof}[Proof of Corollary~\ref{cor:asymp-cons}]
  By definition of the oracle and Lemma~\ref{lemma:norm},
  \begin{equation*}
    \E_{P_0}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2
      \right]} \leq \E_{P_0}{\left[ \Vert \phi_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2
      \right]}  
  \end{equation*}
  for all \( n \in \N \). The results then follows from
  Corollary~\ref{cor:oracle-prop}.
\end{proof}

\section{The state learner with targeted learning}
\label{sec:state-learner-with}

In this section show that a product structure is preserved when an estimator
$\bar\Psi(\hat{F}_n, \hat{H}_n)$ is used instead of
$\tilde\Psi(\hat{\Lambda}_n, \hat{\Gamma}_n, \hat{H}_n)$.

\begin{proof}[Proof of Proposition~\ref{prop:dr-structure}]
  For notational convenience we suppress \( X \) in the following. The final
  result can be obtained by adding the argument \( X \) to all functions and
  averaging. We use the relations from equation~(\ref{eq:7}) to write
  \begin{align*}
    & \int_0^{\tau} w(s) 
      \left\{
      \Gamma(s) - \hat{\Gamma}_n(s)
      \right\}
      [\Lambda - \hat{\Lambda}_n](\diff s)
    \\
    & =
      \int_0^{\tau} w(s) 
      \left\{
      \int_0^s \frac{F(\diff u, 2)}{F(u-, 0)} -
      \int_0^s \frac{\hat{F}_n(\diff u, 2)}{\hat{F}_n(u-, 0)}  -
      \right\}
      \left[
      \frac{F(\diff s, 1)}{F(s-, 0)}
      - \frac{\hat{F}_n(\diff s, 1)}{\hat{F}_n(s-, 0)}
      \right]
    \\
    & =
      \int_0^{\tau} w(s) 
      \Bigg\{
      \int_0^s 
      \left(
      \frac{1}{F(u-, 0)} -  \frac{1}{\hat{F}_n(u-, 0)}
      \right) F(\diff u, 2)
    \\
    & \qquad\qquad \qquad
      +
      \int_0^s \frac{1}{\hat{F}_n(u-, 0)} 
      \left[
      F(\diff u, 2) - \hat{F}_n(\diff u, 2)
      \right]
      \Bigg\}
    \\
    & \qquad\qquad \times
      \left[
      \left(
      \frac{1}{F(s-, 0)} -
      \frac{1}{\hat{F}_n(s-, 0)}
      \right)F(\diff s, 1)
       + \frac{1}{\hat{F}_n(s-, 0)}
      \left(
      F(\diff s, 1) -
      \hat{F}_n(\diff s, 1)
      \right)
      \right]
    \\
    &
      = \int_0^{\tau} 
      \int_0^s
      w(s) 
      \left(
      \frac{1}{F(u-, 0)} -  \frac{1}{\hat{F}_n(u-, 0)}
      \right) 
      \left(
      \frac{1}{F(s-, 0)} -
      \frac{1}{\hat{F}_n(s-, 0)}
      \right)F(\diff u, 2)F(\diff s, 1)
    \\
    & \quad +
      \int_0^{\tau}
      \int_0^s
      w(s) 
      \left(
      \frac{1}{F(u-, 0)} -  \frac{1}{\hat{F}_n(u-, 0)}
      \right) \frac{F(\diff u, 2) }{\hat{F}_n(u-,0)}
      \left(
      F(\diff s, 1) -
      \hat{F}_n(\diff s, 1)
      \right)
    \\
    & \quad +
      \int_0^{\tau} 
      \int_0^s      
      \frac{w(s) }{\hat{F}_n(u-, 0)} 
      \left[
      F(\diff u, 2) - \hat{F}_n(\diff u, 2)
      \right]
      \left(
      \frac{1}{F(s-, 0)} -
      \frac{1}{\hat{F}_n(s-, 0)}
      \right)F(\diff s, 1)
    \\
    & \quad +
      \int_0^{\tau} 
      \int_0^s      
      \frac{w(s) }{\hat{F}_n(u-, 0)} 
      \left[
      F(\diff u, 2) - \hat{F}_n(\diff u, 2)
      \right]
      \frac{1}{\hat{F}_n(s-, 0)}
      \left(
      F(\diff s, 1) -
      \hat{F}_n(\diff s, 1)
      \right).
  \end{align*}
  Consider the first term on the right hand side. Defining
  \begin{equation*}
    w_n^*(t)  = 
    \left(
      F(t-, 0)
      - \hat{F}_n(t-, 0)
    \right)
    \left(
      \frac{1}{F(t-, 0)}
      - \frac{1}{\hat{F}_n(t-, 0)}
    \right),
  \end{equation*}
  we can write
  \begin{align*}
    & \int_0^{\tau} 
      \int_0^s
      w(s) 
      \left(
      \frac{1}{F(u-, 0)} -  \frac{1}{\hat{F}_n(u-, 0)}
      \right)      
      \left(
      \frac{1}{F(s-, 0)} -
      \frac{1}{\hat{F}_n(s-, 0)}
      \right)F(\diff u, 2)F(\diff s, 1)
    \\
    & =
      \int_0^{\tau} 
      \int_0^s
      w(s)
      w_n^*(u) 
      \left(
      F(u-, 0) - \hat{F}_n(u-, 0)
      \right)
    \\
    & \qquad \qquad \quad
      \times
      w_n^*(s) 
      \left(
      F(s-, 0) - \hat{F}_n(s-, 0)
      \right)       
      F(\diff u, 2)F(\diff s, 1)
    \\
    & =
      \int_0^{\tau} 
      \int_0^s
      w_n^a(s,u)
      \left(
      F(u-, 0) - \hat{F}_n(u-, 0)
      \right)
      \left(
      F(s-, 0) - \hat{F}_n(s-, 0)
      \right)       
      F(\diff u, 2)F(\diff s, 1),
  \end{align*}
  where we have defined \( w_n^a(s,u) = w(s)w^*_n(s)w^*_n(u) \). By assumption,
  \( w_n^a(s,u) \) is uniformly bounded. The same approach can be applied to the
  three remaining terms which gives the result.
\end{proof}

\bibliography{bib.bib}

\end{document}