### 2025-01-13 manuscript/suppl.tex

\documentclass[11pt]{article}
\usepackage[a4paper,left=2.8cm, right=2.8cm, top=3cm, bottom=3cm]{geometry}

%% Packages
\usepackage{amsthm, dsfont, amssymb,soul,xcolor, enumitem, amsmath,verbatim}

\usepackage[authoryear]{natbib}
\bibliographystyle{abbrvnat}

\definecolor{linkcolor}{rgb}{0, 0, 0.54}
\usepackage[colorlinks,allcolors=linkcolor]{hyperref}

\theoremstyle{thmstyleone}%\
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%
\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}

%% Line spacing
\linespread{2}

% \usepackage{lineno}
% \linenumbers

\usepackage{endfloat}

%% Only for comments:
\usepackage[author=]{fixme}
\fxusetheme{color}
\definecolor{fxtarget}{rgb}{.5,.5,.5}
\definecolor{fxnote}{rgb}{.5,.5,.5}
\fxsetup{status=draft}

%% Toggle with or without figures
% \def\nofig{1}

\RequirePackage{graphicx}
\usepackage[font=small]{caption}

\newcommand{\includeFigCond}[2][]{
  \ifx\nofig\undefined %
    \includegraphics[#1]{#2} %
  \else %
    \texttt{#2} %
  \fi %
}

% New operators and commands
\DeclareMathOperator{\E}{\mathbb{E}} % expectation
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\C}{\mathbb{C}}
\renewcommand{\S}{\mathbb{S}}
\newcommand{\blank}{\makebox[1ex]{\textbf{$\cdot$}}}
\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
\renewcommand{\phi}{\varphi}
\renewcommand{\epsilon}{\varepsilon}
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand{\weakly}{\rightsquigarrow}
\newcommand\smallO{\textit{o}}
\newcommand\bigO{\textit{O}}
\newcommand{\midd}{\; \middle|\;}
\newcommand{\1}{\mathds{1}}
\usepackage{ifthen} %% Empirical process with default argument
\newcommand{\G}[2][n]{
{\ensuremath{\mathbb{G}_{#1}}{\left[#2\right]}}
}
\DeclareMathOperator*{\argmin}{\arg\!\min}
\DeclareMathOperator*{\argmax}{\arg\!\max}
\newcommand{\empmeas}{\ensuremath{\mathbb{P}_n}} % empirical measure
\newcommand{\data}{\ensuremath{\mathcal{D}}}

%% Title
\title{\Large Supplementary material: The state learner -- a super learner for right-censored data}
\date{\small \today}

\usepackage{authblk}
\author[1,*]{\normalsize Anders Munch}
\author[1]{\normalsize Thomas~A.~Gerds}
\affil[1]{\normalsize Section of Biostatistics, University of Copenhagen}
\affil[*]{\normalsize Address for correspondence: \href{email:email-id.com}{a.munch@sund.ku.dk}}

\usepackage{fancyhdr}
\fancyhf{}
\fancyhead[l]{\textit{Munch et al.}}
\fancyhead[r]{\textit{The state learner}}
\fancyfoot[c]{\thepage}
\renewcommand\headrulewidth{0pt}
\pagestyle{fancy}

\renewcommand{\thesection}{S\arabic{section}}

\begin{document}

\maketitle

This document contains proofs of the results stated in the main paper.
Section~\ref{sec:consistency} contains a proof of the consistency result from
Section~5.1 of the main paper; Section~\ref{sec:oracle-inequalities} contains
proofs of the oracle inequalities from Section~5.2 of the main paper; and
Section~\ref{sec:state-learner-with} demonstrates transience of the second order
remainder structure stated in Section~5.3 of the main paper.

We start by recalling some notation. We use \( F \) to denote a function that
describes the conditional state occupation probabilities of the observed data,
as defined in equation~(5) of the main paper. We denote the integrated Brier
score as \( \bar B_\tau( F,O) = \int_0^{\tau} B_t(F,O) \diff t \), where
\( B_t \) is defined
\begin{equation*}
  B_t(F,O) = \sum_{j=-1}^{2}
  \left(
      F(t,j,X) - \1{\{\eta(t)=j\}}
  \right)^2.
\end{equation*}
Finally, recall that we let $\mathcal{H}_{\mathcal{P}}$ denote the function
space consisting of all conditional state occupation probability functions for
some measure \( P \in \mathcal{P} \), and that we equipped this space with the
norm
\begin{equation}
  \label{eq:norm}
  \| F \|_{P_0} = 
  \left\{
    \sum_{j=-1}^{2}\int_{\mathcal{X}} \int_0^{\tau} F(t, j, x)^2 \diff t H_0( \diff x)
  \right\}^{1/2},
\end{equation}
for some fixed measure \( P_0 \in \mathcal{P} \). We use \( F_0 \) to denote the
conditional state occupation probability function associated with the measure
\( P_0 \).

\section{Consistency}
\label{sec:consistency}

Define
\( \bar{B}_{\tau,0}(F, o) = \bar{B}_{\tau}(F, o) - \bar{B}_{\tau}(F_0, o) \) and
\( R_{0}(F) = P_0{[\bar{B}_{\tau,0}(F, \blank)]} \).
\begin{lemma}
  \label{lemma:norm}
  \( R_{0}(F) = \Vert F - F_0 \Vert_{P_0}^2 \), where \( \Vert \blank \Vert_{P_0}\) is defined
  in equation~(\ref{eq:norm}).
\end{lemma}
\begin{proof}[Proof]
  For any \( t \in [0, \tau] \) and \( j\in \{-1,0,1,2\} \) we have
  \begin{align*}
    & \E_{P_0}{\left[ (F(t, j, X) - \1{\{\eta(t) = j \}})^2 \right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, j, X) - F_0(t, j, X) + F_0(t, j, X) - \1{\{\eta(t) = j
      \}})^2 \right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, j, X) - F_0(t, j, X))^2\right]}
      + \E_{P_0}{\left[ (F_0(t, j, X) - \1{\{\eta(t) = j \}})^2\right]}
    \\
    & \quad
      + 2\E_{P_0}{\left[ (F(t, j, X) - F_0(t, j, X))(F_0(t, j, X) - \1{\{\eta(t) = j
      \}})\right]}
    \\
    & =    \E_{P_0}{\left[ (F(t, j, X) - F_0(t, j, X))^2\right]}
      + \E_{P_0}{\left[ (F_0(t, j, X) - \1{\{\eta(t) = j \}})^2\right]},
  \end{align*}
  where the last equality follows from the tower property. Hence, using Fubini,
  we have
  \begin{equation*}
    P{[\bar{B}_{\tau}(F, \blank)]}
    = \Vert F - F_0 \Vert_{P_0}^2 + P_0{[\bar{B}_{\tau}(F_0, \blank)]}.
  \end{equation*}
\end{proof}

\begin{proof}[Proof of Proposition~1 of the main paper]
  The result follows from Lemma~\ref{lemma:norm}.
\end{proof}

\section{Oracle inequalities}
\label{sec:oracle-inequalities}

Recall that we use \( \mathcal{F}_n \) to denote a library of learners for the
function \( F \), and that \( \hat{\phi} \) and \( \tilde{\phi} \) denotes,
respectively, the discrete super learner and the oracle learner for the library
\( \mathcal{F}_n \), c.f., Section~4 of the main paper.

\begin{proof}[Proof of Corollary~1 of the main paper]
  First note that minimising the loss \( \bar{B}_{\tau} \) is equivalent to
  minimising the loss \( \bar{B}_{\tau,0} \), so the discrete super learner and
  oracle according to \( \bar{B}_{\tau} \) and \( \bar{B}_{\tau,0} \) are
  identical. By Lemma~\ref{lemma:norm}, \( R_0(F) \geq 0 \) for any
  \( F \in \mathcal{H}_{\mathcal{P}} \), and so using Theorem 2.3 from \citep{vaart2006oracle} with
  \( p=1 \), we have that for all \( \delta >0 \),
\begin{align*}
  & \E_{P_0}{\left[ R_0(\hat{\phi}_n(\data_n^{-k})) \right]}
  \\
  &  \quad \leq
    (1+2\delta)\E_{P_0}{\left[ R_0(\tilde{\phi}_n(\data_n^{-k})) \right]}
  \\
  & \qquad + (1+\delta) \frac{16 K}{n}
    \log(1 + |\mathcal{F}_n|)\sup_{F \in \mathcal{H}_{\mathcal{P}}}
    \left\{
    M(F) + \frac{v(F)}{R_0(F)}
    \left(
    \frac{1}{\delta} + 1
    \right)
    \right\}
\end{align*}
where for each \( F \in \mathcal{H}_{\mathcal{P}} \), \( (M(F), v(F)) \) is some Bernstein pair for
the function \(o \mapsto \bar{B}_{\tau,0}(F, o) \). As
\( \bar{B}_{\tau,0}(F, \blank) \) is uniformly bounded by \( \tau \) for any
\( F \in \mathcal{H}_{\mathcal{P}} \), it follows from section 8.1 in \citep{vaart2006oracle} that
\( (\tau, 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}) \) is a Bernstein pair for
\( \bar{B}_{\tau,0}(F, \blank) \). Now, for any \( a,b,c \in \R \) we have
\begin{align*}
  (a-c)^2 - (b-c)^2
  & = (a-b+b-c)^2 - (b-c)^2
  \\
  & = (a-b)^2 + (b-c)^2 +2(b-c)(a-b) - (b-c)^2
  \\
  & = (a-b)
    \left\{
    (a-b) +  2(b-c)
    \right\}
  \\
  & = (a-b)
    \left\{
     a + b -2c
    \right\},
\end{align*}
so using this with \( a=F(t, j, x) \), \( b=F_0(t, j, x) \), and
\( c = \1{\{\eta(t) = j\}} \), we have by Jensen's inequality
\begin{align*}
  & P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}
  \\
  & \leq
    2\tau\E_{P_0}{\left[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left\{
    \left(
    F(t, j, X) - \1{\{\eta(t) = j\}}
    \right)^2
    -
    \left(
    F_0(t, j, X) - \1{\{\eta(t) = j\}}
    \right)^2
    \right\}^2
    \diff t 
    \right]}
  \\
  & =2\tau
    \E_{P_0}\Bigg[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left(
    F(t, j, X) - F_0(t, j, X)
    \right)^2
  \\
  & \quad \quad \quad\quad \quad \quad \times
    \left\{
    F(t, j, X) +  F_0(t, j, X)-2 \1{\{\eta(t) = j\}}
    \right\}^2
    \diff t 
    \Bigg]
  \\
  & \leq
    8\tau \E_{P_0}{\left[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left(
    F(t, j, X) - F_0(t, j, X)
    \right)^2
    \diff t 
    \right]}.
  \\
  & =
    8\tau \Vert F - F_0 \Vert_{P_0}^2.
\end{align*}
Thus when \( v(F) = 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]} \) we have by
Lemma~\ref{lemma:norm}
\begin{equation*}
  \frac{v(F)}{R_0(F)}
  = 1.5 \frac{P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}}{P_0{[\bar{B}_{\tau,0}(F, \blank)]}}
  \leq 12 \tau,
\end{equation*}
and so using the Bernstein pairs \( (\tau, 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}) \) we have
\begin{equation*}
  \sup_{F \in \mathcal{H}_{\mathcal{P}}}
  \left\{
    M(F) + \frac{v(F)}{R_0(F)}
    \left(
      \frac{1}{\delta} + 1
    \right)
  \right\}
  \leq \tau
  \left(
    13 + \frac{12}{\delta}
  \right),
\end{equation*}
For all $\delta>0$ we thus have
\begin{align*}
  \E_{P_0}{\left[ R_0(\hat{\phi}_n(\data_n^{-k})) \right]}
  \leq
  &(1+2\delta)\E_{P_0}{\left[ R_0(\tilde{\phi}_n(\data_n^{-k})) \right]}
  \\
  & \quad
    + (1+\delta)\log(1 + |\mathcal{F}_n|) \tau \frac{16 K}{n}
    \left(
    13 + \frac{12}{\delta}
    \right),
\end{align*}
and then the final result follows from Lemma~\ref{lemma:norm}.
\end{proof}

\begin{proof}[Proof of Corollary~2 of the main paper]
  By definition of the oracle and Lemma~\ref{lemma:norm},
  \begin{equation*}
    \E_{P_0}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2
      \right]} \leq \E_{P_0}{\left[ \Vert \phi_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2
      \right]}  
  \end{equation*}
  for all \( n \in \N \). The results then follows from
  Corollary~1 of the main paper.
\end{proof}

\section{Transience of the second order remainder structure}
\label{sec:state-learner-with}

Recall that we let $\Lambda_1$ denote the conditional cumulative hazard function
for one of the event times of interest and $\Gamma$ the conditional cumulative
hazard function for censoring, c.f., Section~2 of the main paper.

\begin{proof}[Proof of Proposition~2 of the main paper]
  For notational convenience we suppress \( X \) in the following. The final
  result can be obtained by adding the argument \( X \) to all functions and
  averaging. We use the relations from equation~(8) in the main paper to write
  \begingroup %
  \allowdisplaybreaks
    \begin{align*}
      & \int_0^{\tau} w(s) 
        \left\{
        \Gamma(s) - \hat{\Gamma}_n(s)
        \right\}
        [\Lambda_1 - \hat{\Lambda}_{1,n}](\diff s)
      \\
      & =
        \int_0^{\tau} w(s) 
        \left\{
        \int_0^s \frac{F(\diff u, -1)}{F(u-, 0)} -
        \int_0^s \frac{\hat{F}_n(\diff u, -1)}{\hat{F}_n(u-, 0)}  -
        \right\}
        \left[
        \frac{F(\diff s, 1)}{F(s-, 0)}
        - \frac{\hat{F}_n(\diff s, 1)}{\hat{F}_n(s-, 0)}
        \right]
      \\
      & =
        \int_0^{\tau} w(s) 
        \Bigg\{
        \int_0^s 
        \left(
        \frac{1}{F(u-, 0)} -  \frac{1}{\hat{F}_n(u-, 0)}
        \right) F(\diff u, -1)
      \\
      & \qquad\qquad \qquad
        +
        \int_0^s \frac{1}{\hat{F}_n(u-, 0)} 
        \left[
        F(\diff u, -1) - \hat{F}_n(\diff u, -1)
        \right]
        \Bigg\}
      \\
      & \qquad\qquad \times
        \left[
        \left(
        \frac{1}{F(s-, 0)} -
        \frac{1}{\hat{F}_n(s-, 0)}
        \right)F(\diff s, 1)
        + \frac{1}{\hat{F}_n(s-, 0)}
        \left(
        F(\diff s, 1) -
        \hat{F}_n(\diff s, 1)
        \right)
        \right]
      \\
      &
        = \int_0^{\tau} 
        \int_0^s
        w(s) 
        \left(
        \frac{1}{F(u-, 0)} -  \frac{1}{\hat{F}_n(u-, 0)}
        \right) 
        \left(
        \frac{1}{F(s-, 0)} -
        \frac{1}{\hat{F}_n(s-, 0)}
        \right)F(\diff u, -1)F(\diff s, 1)
      \\
      & \quad +
        \int_0^{\tau}
        \int_0^s
        w(s) 
        \left(
        \frac{1}{F(u-, 0)} -  \frac{1}{\hat{F}_n(u-, 0)}
        \right) \frac{F(\diff u, -1) }{\hat{F}_n(u-,0)}
        \left(
        F(\diff s, 1) -
        \hat{F}_n(\diff s, 1)
        \right)
      \\
      & \quad +
        \int_0^{\tau} 
        \int_0^s      
        \frac{w(s) }{\hat{F}_n(u-, 0)} 
        \left[
        F(\diff u, -1) - \hat{F}_n(\diff u, -1)
        \right]
        \left(
        \frac{1}{F(s-, 0)} -
        \frac{1}{\hat{F}_n(s-, 0)}
        \right)F(\diff s, 1)
      \\
      & \quad +
        \int_0^{\tau} 
        \int_0^s      
        \frac{w(s) }{\hat{F}_n(u-, 0)} 
        \left[
        F(\diff u, -1) - \hat{F}_n(\diff u, -1)
        \right]
        \frac{1}{\hat{F}_n(s-, 0)}
        \left(
        F(\diff s, 1) -
        \hat{F}_n(\diff s, 1)
        \right).
    \end{align*}
    \endgroup %
    Consider the first term on the right hand side. By the mean value theorem,
    \begin{equation*}
      \frac{1}{F(t-, 0)}
      - \frac{1}{\hat{F}_n(t-, 0)}
      = \frac{-1}{\tilde{r}_n(t)^2}
      \left[
        F(t-, 0)
        - \hat{F}_n(t-, 0)
      \right],
    \end{equation*}
    where \( \tilde{r}_n(t) \) is some value between \( \hat{F}(t-, 0) \) and
    \( \hat{F}_n(t-, 0) \). Letting \( w_n^*(t) = -\tilde{r}_n(t)^{-2} \), we
    may write
  \begin{align*}
    & \int_0^{\tau} 
      \int_0^s
      w(s) 
      \left(
      \frac{1}{F(u-, 0)} -  \frac{1}{\hat{F}_n(u-, 0)}
      \right)      
      \left(
      \frac{1}{F(s-, 0)} -
      \frac{1}{\hat{F}_n(s-, 0)}
      \right)F(\diff u, -1)F(\diff s, 1)
    \\
    & =
      \int_0^{\tau} 
      \int_0^s
      w(s)
      w_n^*(u) 
      \left(
      F(u-, 0) - \hat{F}_n(u-, 0)
      \right)
    \\
    & \qquad \qquad \quad
      \times
      w_n^*(s) 
      \left(
      F(s-, 0) - \hat{F}_n(s-, 0)
      \right)       
      F(\diff u, -1)F(\diff s, 1)
    \\
    & =
      \int_0^{\tau} 
      \int_0^s
      w_n^a(s,u)
      \left(
      F(u-, 0) - \hat{F}_n(u-, 0)
      \right)
      \left(
      F(s-, 0) - \hat{F}_n(s-, 0)
      \right)       
      F(\diff u, -1)F(\diff s, 1),
  \end{align*}
  where we have defined \( w_n^a(s,u) = w(s)w^*_n(s)w^*_n(u) \). By the
  assumption that \( F(\blank,0) \) and \( F_n(\blank,0) \) are uniformly
  bounded away from zero on \( [0,\tau] \), it follows that \( w_n^* \) is
  uniformly bounded, and hence \( w_n^a(s,u) \) is also uniformly bounded,
  because \( w \) was assumed uniformly bounded. The same approach can be
  applied to the three remaining terms which gives the result.
\end{proof}

\bibliography{bib.bib}

\end{document}