### 2025-05-28 manuscript/manuscript-one.tex
rephrasing

\documentclass[a4paper,danish]{article}
\usepackage[margin=3.5cm]{geometry}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{natbib, dsfont, enumitem,
  amssymb,soul,xcolor,amsmath,amsthm,graphicx,subcaption,verbatim,pgfplots,tikz,prodint}
\usetikzlibrary{calc,patterns,angles,quotes,automata,
  positioning,arrows,shapes}

\definecolor{linkcolor}{rgb}{0, 0, 0.54}
\usepackage[colorlinks,allcolors=linkcolor]{hyperref}

\usepackage[author=]{fixme}
\fxusetheme{color}
\definecolor{fxtarget}{rgb}{.5,.5,.5}
\definecolor{fxnote}{rgb}{.5,.5,.5}

\fxsetup{status=draft}

% Ref and bibliography style 
\bibliographystyle{abbrvnat}

% Handling new lines
\setlength{\parskip}{1em}
\setlength{\parindent}{0em}

% Handling space after sections
\usepackage{titlesec}
\titlespacing*{\section}{0em}{2em}{0em}
\titlespacing*{\subsection}{0em}{2em}{0em}
\titlespacing*{\subsubsection}{0em}{2em}{0em}

% No spacing after in start of list
\setlist[itemize]{topsep=0pt}
\setlist[enumerate]{topsep=0pt}

% Theorem environments 
\theoremstyle{plain} % plain italic style
\newtheorem{theorem}{Theorem}
\numberwithin{theorem}{section}
\newtheorem*{theorem*}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition} % non-italic
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem*{assumption*}{Assumption}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{note}[theorem]{Note}

% Notation
\DeclareMathOperator{\E}{\mathbb{E}} % expectation
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\C}{\mathbb{C}}
\renewcommand{\S}{\mathbb{S}}
\newcommand{\blank}{\makebox[1ex]{\textbf{$\cdot$}}}
\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
\renewcommand{\phi}{\varphi}
\renewcommand{\epsilon}{\varepsilon}
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand{\weakly}{\rightsquigarrow}
\newcommand\smallO{\textit{o}}
\newcommand\bigO{\textit{O}}
\newcommand{\midd}{\; \middle|\;}
\newcommand{\1}{\mathds{1}}
\usepackage{ifthen} %% Empirical process with default argument
\newcommand{\G}[2][n]{
{\ensuremath{\mathbb{G}_{#1}}{\left[#2\right]}}
}
\DeclareMathOperator*{\argmin}{\arg\!\min}
\DeclareMathOperator*{\argmax}{\arg\!\max}
\newcommand{\empmeas}{\ensuremath{\mathbb{P}_n}} % empirical measure
\newcommand{\data}{\ensuremath{\mathcal{D}}}

\title{The state learner -- a super learner for right-censored
  data}
\author{Anders Munch \& Thomas A. Gerds}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
  In survival analysis, prediction models are needed as
  stand-alone tools and in applications of causal inference to
  estimate nuisance parameters. The super learner is a machine
  learning algorithm which combines a library of prediction
  models into a meta learner based on cross-validated loss. In
  right-censored data, the choice of the loss function and the
  estimation of the expected loss need careful consideration.
  We introduce the state learner, a new super learner for
  survival analysis, which simultaneously evaluates libraries
  of prediction models for the event of interest and the
  censoring distribution. The state learner places no
  restrictions on the algorithms that can be included in the
  library, works in the presence of competing risks, and does
  not require a single pre-specified estimator of the
  conditional censoring distribution. We establish an oracle
  inequality for the state learner and investigate its
  performance through numerical experiments. We illustrate the
  application of the state learner with prostate cancer data.
\end{abstract}

\textit{\textbf{Keywords:} Competing risks, cross-validation,
  loss based estimation, right-censored data, super learner}

\section{Introduction}
\label{sec:introduction}

Accurate risk prediction from time-to-event data is a
important problem in many areas of medical research and public
health. Super learning \citep{van2007super}, also known as
ensemble learning or stacked regression
\citep{wolpert1992stacked,breiman1996stacked}, provides a
powerful approach to this problem by combining multiple
candidate prediction models. In the context of survival
analysis, a super learner can for instance combine Cox
regression models and random survival forests \citep[][Section
8.4]{gerds2021medical}.

In this paper we propose the state learner, a new super
learner designed to handle the specific challenges of ensemble
learning with right-censored data. The state learner estimates
the performance of a library of candidate models for both the
event-time and censoring distributions simultaneously. This
joint estimation framework avoids the need for pre-specifying
a censoring model, which is a limitation of many existing
methods. Our approach also allows for full flexibility in the
choice of models in the library, contrasting with earlier
proposals that restrict the library to specific model classes
\citep{polley2011-sl-cens,golmakani2020super}. Furthermore,
unlike iterative approaches such as those in
\cite{han2021inverse} and \cite{westling2021inference}, we can
provide theoretical guarantees for the performance of the
state learner.

The state learner is based on the simple idea of considering
right-censored competing risks data as an artificial
multi-state system where censoring is included as a state of
its own. Candidate models are then assessed based on how well
they predict the state occupations of this system across time,
based on baseline covariate information. We demonstrate how to construct
a library for the state learner using existing survival
models, and how to obtain risk predictions from the resulting
ensemble. To analyse the theoretical properties of the state
learner we focus on the discrete super learner which picks the
model in the library with best estimated performance
\citep{van2007super}. We show that the discrete state learner
is consistent under natural conditions and prove a finite-sample
oracle inequality.

\subsection{Related work}
\label{sec:related-work}

Super learning requires a library of candidate models, or
\textit{learners}, and a loss function for evaluating
performance on hold-out samples. Machine learning based on
right-censored data commonly uses the partial log-likelihood
as a loss function
\citep[e.g.,][]{li2016regularized,yao2017deep,lee2018deephit,katzman2018deepsurv,gensheimer2019scalable,lee2021boosted,kvamme2021continuous}.
This loss function is also suggested for super learning with
right-censored data by \cite{polley2011-sl-cens}, where data
is assumed to be observed in discrete time. However, the
partial log-likelihood loss does not work well as a measure of
performance in hold-out samples observed in continuous time.
The reason is that the partial log-likelihood loss assigns an
infinite value for any learner that predicts piece-wise
constant cumulative hazard functions when there are event
times in the test set which do not occur in the learning set.
This problem occurs with prominent survival learners including
the Kaplan-Meier estimator, random survival forests, and
semi-parametric Cox regression models and these learners
cannot be included in the library of the super learner
proposed by \cite{polley2011-sl-cens}. When a proportional
hazards model is assumed, the baseline hazard function can be
profiled out of the likelihood \citep{cox1972regression}. The
cross-validated partial log-likelihood loss
\citep{verweij1993cross} has therefore been suggested as a
loss function for super learning by \cite{golmakani2020super}.
This choice of loss function restricts the library of learners
to include only Cox proportional hazards models, and hence
excludes many learners such as, e.g., random survival forests,
additive hazards models, and accelerated failure time models.

Alternative approaches for super learning with right-censored
data use an inverse probability of censoring weighted (IPCW)
loss function
\citep{graf1999assessment,van2003unicv,molinaro2004tree,keles2004asymptotically,hothorn2006survival,gerds2006consistent,gonzalez2021stacked},
censoring unbiased transformations
\citep{fan1996local,steingrimsson2019censoring}, or
pseudo-values
\citep{andersen2003generalised,mogensen2013random,sachs2019ensemble}.
All these methods rely on an estimator of the censoring
distribution, and their drawback is that this estimator has to
be pre-specified.

More recent work by \cite{han2021inverse} and
\cite{westling2021inference} circumvents the need to
pre-specify a censoring model by iterating
between estimation of the outcome and censoring models. However, it seems that this procedure is in general not
guaranteed to converge to the true data-generating mechanism
\citep[][Appendix~A.4]{munch2024thesis}.

\subsection{Outline of the paper}
\label{sec:outline-paper}

We introduce our notation and framework in
Section~\ref{sec:framework}. Section~\ref{sec:super-learning}
defines general super learning for right-censored data. In
Section~\ref{sec:super-learner-simple} we define the state
learner, while Section~\ref{sec:theor-results-prop} provides
theoretical guarantees. In Section~\ref{sec:use-cases-state}
we demonstrate how the state learner can be used to obtain
risk predictions and how it can be applied in the context of
causal inference. Section~\ref{sec:numer-exper} reports the
results of numerical experiments, and
Section~\ref{sec:real-data-appl} illustrates the method on
prostate cancer data. We conclude with a discussion in
Section~\ref{sec:discussion}. Proofs are collected in
Appendix~\ref{sec:proofs}. Code and an implementation of the
state learner are available at
\url{https://github.com/amnudn/statelearner}.

% A super learner is a machine learning algorithm that combines
% a finite set of learners into a meta learner by estimating
% prediction performance in hold-out samples using a
% pre-specified loss function \citep{van2007super}. When the aim
% is to make a prediction model, super learners combine strong
% learners, such as Cox regression models and random survival
% forests \citep[][Section 8.4]{gerds2021medical}. While the
% general idea of combining strong learners based on
% cross-validation stems from earlier work
% \citep{wolpert1992stacked,breiman1996stacked}, the name super
% learner is justified by an oracle inequality
% \citep{van2003unicv,vaart2006oracle}.

% We define the state learner, a new super learner for
% right-censored data, which simultaneously estimates the
% expected loss of learners of the event time distribution and
% the censoring distribution. Compared to the super learners of
% \cite{polley2011-sl-cens} and \cite{golmakani2020super}, the
% state learner proposed here poses no restrictions on the type
% of learners that can be included in the library. Compared to
% super learners that rely on inverse probability of censoring
% weighting, censoring unbiased transformations, or
% pseudo-values, the state learner does not require a
% pre-specified censoring distribution. Compared to the super
% learners by \cite{han2021inverse} and
% \cite{westling2021inference}, the state learner has
% theoretical guarantees which these methods do not have.

% The loss function underlying the state learner operates on
% coarsened data. In the right-censored survival setting the
% coarsened data consist of the minimum and the order of the
% censoring time and the event time as well as the baseline
% covariates. The state learner works by simultaneously
% assessing how well learners of the event time distribution and
% the censoring distribution predict the states of the coarsened
% data. To analyse the theoretical properties of the state
% learner we focus on the discrete super learner which combines
% the library of learners by picking the one that minimises the
% cross-validated loss \citep{van2007super}. In the presence of
% competing risks, our algorithm uses separate libraries of
% learners for the cumulative hazard functions for each of the
% competing risks and for the censoring distribution. We show
% that the oracle selector of the state learner is consistent if
% all libraries contain a consistent learner and prove a finite
% sample oracle inequality.

% The state learner algorithm can output a medical risk
% prediction model \citep{gerds2021medical} which predicts the
% probability of an event based on covariates in the presence of
% competing risks. Another important application, which
% motivated our contribution, is in targeted learning where
% conditional event probabilities occur as high-dimensional
% nuisance parameters which need to be estimated at a certain
% rate \citep{van2011targeted, rytgaard2021estimation,
%   rytgaard2022targeted}. The state learner is particularly
% well-suited for this application as it simultaneously provides
% estimates for both the outcome and the censoring, both of
% which are needed for targeted learning. For the asymptotic
% bias term of a targeted estimator, which uses the state
% learner to estimate nuisance parameters, we show that a second
% order product structure holds. We illustrate both applications
% of the state learner with prostate cancer data
% \citep{kattan2000pretreatment}.

\section{Notation and framework}
\label{sec:framework}

In a competing risk framework \citep{andersen2012statistical},
let \( T\) be a time to event variable, \(D\in\{1,2\}\) the
cause of the event, and $X \in \mathcal{X}$ a vector of
baseline covariates taking values in a bounded subset
\( \mathcal{X} \subset \R^p \), \( p\in\N \). Let
$\tau< \infty$ be the prediction horizon. We use
\( \mathcal{Q} \) to denote the collection of all probability
measures on \( [0,\tau] \times \{1,2\}\times \mathcal{X} \)
such that \( (T, D, X) \sim Q \) for some unknown
\( Q \in \mathcal{Q} \). For \(j\in\{1,2\}\), the
cause-specific conditional cumulative hazard functions
\( \Lambda_{j} \colon [0, \tau] \times \mathcal{X} \rightarrow
\R_+ \) are defined as
\begin{equation*}
  % \label{eq:cum-haz}
  \Lambda_{j}(t \mid x) = \int_0^t\frac{  Q(T \in \diff s, D=j \mid X=x )}{Q(T \geq s \mid X=x )}.
\end{equation*}
For ease of presentation we assume throughout that the map
\( t\mapsto \Lambda_j(t \mid x) \) is continuous for all
\( x \) and \( j \). All arguments carry over directly to the
general case. We denote by \(S\) the conditional event-free
survival function,
\begin{equation}
  \label{eq:surv-def}
  S(t \mid x)=\exp\left\{-\Lambda_{1}(t \mid x)-\Lambda_{2}(t \mid x)\right\}.
\end{equation}
Let \( \mathcal{M}_{\tau}\) denote the space of all conditional cumulative hazard
functions on \( [0,\tau] \times\mathcal{X}\). Any distribution
\( Q \in \mathcal{Q} \) can be characterised by
\begin{equation*}
  \label{eq:parametrizeQ}
  \begin{split}
    Q(\diff t,j,\diff x)=& \left\{S(t- \mid x)\Lambda_1(\diff t \mid x)H(\diff x)\right\}^{\1{\{j=1\}}}\\
                         &  \left\{S(t- \mid x)\Lambda_2(\diff t \mid x)H(\diff x)\right\}^{\1{\{j=2\}}},
  \end{split}
\end{equation*}
where \(\Lambda_{j} \in \mathcal{M}_{\tau}\) for \(j=1,2\) and \(H\) is the marginal
distribution of the covariates.

We consider the right-censored setting in which we observe
\(O = (\tilde{T},\tilde D, X)\), where $\tilde T = \min(T,C)$
for a right-censoring time \(C\), $\Delta = \1{\{T \leq C\}}$,
and \(\tilde D=\Delta D\). Let \(\mathcal{P}\) denote a set of
probability measures on the sample space
\(\mathcal{O} = [0, \tau] \times \{0, 1, 2\} \times
\mathcal{X}\) such that \(O \sim P \) for some unknown
\(P\in \mathcal{P}\). We assume that the event times and the
censoring times are conditionally independent given
covariates, \( T \independent C \mid X \). This implies that
any distribution \( P \in \mathcal{P} \) is characterised by a
distribution \( Q \in \mathcal{Q} \) and a conditional
cumulative hazard function for \( C \) given \( X \)
\citep[c.f.,][]{begun1983information,gill1997coarsening}. We
use \(\Gamma\in\mathcal{M}_{\tau}\) to denote the conditional
cumulative hazard function for censoring. For ease of
presentation we now also assume that
\( \Gamma(\blank \mid x) \) is continuous for all \( x \). We
let
\((t,x)\mapsto G(t \mid x)=\exp\left\{-\Gamma(t \mid
  x)\right\}\) denote the survival function of the conditional
censoring distribution. The distribution \( P \) is
characterised by
\begin{equation}\label{eq:parametrizeP}
  \begin{split}
    P(\diff t, j, \diff x) =& \left\{G(t- \mid x)S(t- \mid x)\Lambda_1(\diff t \mid x)H(\diff x)\right\}^{\1{{\{j=1\}}}}\\
                            & \left\{G(t- \mid x)S(t- \mid x)\Lambda_2(\diff t \mid x)H(\diff x)\right\}^{\1{{\{j=2\}}}}\\
                            & \left\{G(t- \mid x)S(t- \mid x)\Gamma(\diff t \mid x)H(\diff x)\right\}^{\1{{\{j=0\}}}}\\
    = & \left\{G(t- \mid x)Q(\diff t,j,\diff x)\right\}^{\1{{\{j\ne 0\}}}}\\    
                            & \left\{G(t- \mid x)S(t- \mid x)\Gamma(\diff t \mid x)H(\diff x)\right\}^{\1{{\{j=0\}}}}.
  \end{split}
\end{equation}
Hence, we may write
\( \mathcal{P} = \{ P_{Q, \Gamma} : Q \in \mathcal{Q}, \Gamma \in
\mathcal{G} \} \) for some \( \mathcal{G} \subset \mathcal{M}_{\tau} \). We
also have \(H\)-almost everywhere
\begin{equation*}
P(\tilde T>t \mid X=x) = S(t \mid x)G(t \mid x) = \exp\left\{-\Lambda_{1}(t \mid x)-\Lambda_{2}(t \mid x)-\Gamma(t \mid x) \right\}.
\end{equation*}
We further assume that there exists \(\kappa<\infty\) such that
\(\Lambda_{j}(\tau- \mid x)<\kappa \), for \(j\in\{1,2\}\), and
\(\Gamma(\tau- \mid x)<\kappa\) for almost all \(x\in\mathcal X\). Note that this
implies that \(G(\tau- \mid x)\) is bounded away from zero for almost all \(x\in\mathcal X\).
Under these assumptions, the conditional cumulative hazard functions
\(\Lambda_{j}\) and \(\Gamma\) can be identified from \(P\) by
\begin{align}
  \Lambda_{j}(t \mid x) &= \int_0^t\frac{  P(\tilde T \in \diff s, \tilde D=j \mid X=x )}{P(\tilde T \geq s \mid X=x )}, \label{eq:lambdaj}\\
  \Gamma(t \mid x) &= \int_0^t\frac{  P(\tilde T \in \diff s, \tilde D=0 \mid X=x )}{P(\tilde T \geq s \mid X=x )}\label{eq:gamma}.
\end{align}
Thus, we can consider $\Lambda_j$ and \(\Gamma\) as operators which map from
\( \mathcal{P} \) to \(\mathcal M_{\tau}\).

\section{Super learning}
\label{sec:super-learning}

In survival analysis, a super learner can be used to estimate
a parameter $\theta \in \Theta$ which can be identified from
the observed data distribution \(P\in\mathcal P\). A super
learner typically estimates a function, and thus the parameter
space $\Theta$ is most often a class of functions. For
example, $\theta$ could be a conditional survival function or
a cumulative hazard function. In
Section~\ref{sec:super-learner-simple} we consider a specific
choice of parameter as target for our super learner, but in
this section we describe super learning for a general
parameter $\theta$ with values in some parameter space
$\Theta$.

As input to a super learner we need a data set
\( \data_n=\{O_i\}_{i=1}^n \) of i.i.d.\ observations from
\( P \in \mathcal{P} \) and a collection of candidate models
or learners $\mathcal{A}$. Each learner \(a \in \mathcal{A}\)
is a map \( a \colon \mathcal{O}^n \rightarrow \Theta \) which
takes a data set as input and returns an estimate
$a(\data_n) \in \Theta$ of $\theta$.
% Formally, the domain of $a$ depends on \(n\) but we suppress this in the notation.
In what follows, we use the short-hand notation
\(P[f] = \int f(o) P(\diff o) \). A super learner evaluates the
performance of \(a \in \mathcal{A}\) with a loss function
\(L\colon \Theta \times \mathcal{O} \rightarrow \R_+\) and
estimates the expected loss \(P[L(a(\data_n), \blank)]\) using
cross-validation. Specifically, the expected loss of $a\in\mathcal A$
is estimated by splitting the data set $\data_n$ into $K$ disjoint
approximately equally sized subsets
\(\data_n^1, \data_n^2, \dots, \data_n^K \) and then calculating the
cross-validated loss
\begin{equation*}
  % \label{eq:cv-risk-est}
  \hat{R}_n(a; L) =
  \frac{1}{K}\sum_{k=1}^{K}
  % \empmeas^k{[L {(a{ (\data_n^{-k})} , \blank) }]},
  \frac{1}{| \data_n^{k} |}\sum_{O_i \in \data_n^{k}}
  L
  {
    \left(
      a{ (\data_n^{-k})}
      , O_i
    \right)
  },
  \quad \text{with} \quad
  \data_n^{-k} = \data_n \setminus \data_n^{k}.
\end{equation*}
% where \( \empmeas^{k} \) is the empirical measure of \( \data_n^{k} \). 
The subset \(\data_n^{-k}\) is referred to as the \(k\)'th training
sample, while \(\data_n^{k}\) is referred to as the \(k\)'th test or
hold-out sample.
The discrete super learner is defined as
\begin{equation*}
\hat{a}_n = \argmin_{a\in\mathcal A}\hat{R}_n(a; L).
\end{equation*}
The oracle learner is defined as the learner that minimises the
expected loss under the data-generating distribution \( P \),
i.e.,
\begin{equation*}
  \tilde{a}_n =
  \argmin_{a \in \mathcal{A}}
  \tilde{R}_n(a; L),
  \quad \text{with} \quad 
  \tilde{R}_n(a; L)=
  \frac{1}{K}\sum_{k=1}^{K} 
  P{
    \left[
      L
      {
        \left(
          a{ (\data_n^{-k})}
          , \blank
        \right)
      }
    \right]}
  .
\end{equation*}
Note that both the discrete super learner and the oracle learner
depend on the library of learners and on the number of folds \(K\),
and that the oracle learner is a function of the data and the unknown
data-generating distribution. However, these dependencies are
suppressed in the notation.

\section{The state learner}
\label{sec:super-learner-simple}

The main idea of the state learner is to jointly use learners
for \( \Lambda_1 \), \( \Lambda_2 \), and \( \Gamma \), and
the relations in equation~(\ref{eq:parametrizeP}), to learn a
feature of the observed data distribution \( P \). The
discrete state learner ranks a tuple of learners for the tuple
of the cumulative hazard functions
\( (\Lambda_1, \Lambda_2, \Gamma) \) based on how well they
jointly model the observed data. To formally introduce the
state learner, we define the artificial multi-state process
\begin{equation*}
  \eta(t) = \1\{\tilde{T} \leq t, \tilde D=1\} + 2\,\1\{\tilde{T} \leq t, \tilde
  D=2\} - \1\{\tilde{T} \leq t, \tilde D=0\},
  \quad \text{for} \quad t \in [0, \tau].
\end{equation*}
We use the phrase `artificial' because multi-state models
should not usually include censoring as a state.
At time \(t\), we observe that each individual is in one of four mutually
exclusive states: \( 0 \), \( 1 \), \( 2 \), or \( -1 \).
The conditional distribution of the process \( \eta(t) \) given
baseline covariates \( X \) is determined by the function
\begin{equation}
  \label{eq:F-def}
  F(t, k, x) = P(\eta(t) = k \mid X=x),
\end{equation}
for all \( t \in [0,\tau] \), \( k \in \{-1,0,1,2\} \), and
\( x \in \mathcal{X} \). The function \( F \) describes the
conditional state occupation probabilities of the process
\(\eta\).

We define the state learner as a super learner for the
function-valued parameter $\theta(P) = F$ which is identified
through equation~(\ref{eq:F-def}). Under conditional
independent censoring each tuple
$(\Lambda_{1}, \Lambda_{2}, \Gamma, H)$ characterises a
distribution \(P\in\mathcal P\), c.f.\
equation~\eqref{eq:parametrizeP}, which in turn determines
\( (F, H) \). Hence, a learner for \( F \) can be constructed
from learners for \( \Lambda_1 \), \( \Lambda_2 \), and
$\Gamma$ as follows:
\begin{equation}\label{eq:transition}
  \begin{split}
    F(t, 0, x)
    &
      = P(\tilde{T} > t \mid X= x)
      =
      \exp{{\{-\Lambda_{1}(t \mid x)-\Lambda_{2}(t \mid x) - \Gamma(t \mid x)\}
      }},
    \\
    F(t, 1, x)
    &
      = P(\tilde{T} \leq t, \tilde{D}=1 \mid X=x)
      = \int_0^t F(s-, 0, x)  \Lambda_{1}(\diff s \mid x),
    \\
    F(t, 2, x)
    &
      = P(\tilde{T} \leq t, \tilde{D}=2 \mid X=x)
      = \int_0^t  F(s-, 0, x)  \Lambda_{2}(\diff s \mid x),
    \\
    F(t, -1, x)
    &
      = P(\tilde{T} \leq t, \tilde{D}=0 \mid X=x)
      = \int_0^t F(s-, 0, x)  \Gamma(\diff s \mid x).
  \end{split}
\end{equation}
Equation~\eqref{eq:transition} implies that a library for the
state learner can by build from three libraries of learners:
\(\mathcal{A}_1\), \( \mathcal{A}_2 \), and \( \mathcal{B} \),
where \(\mathcal{A}_1\) and \( \mathcal{A}_2\) contain
learners for the conditional cause-specific cumulative hazard
functions \(\Lambda_1\) and \( \Lambda_2\), respectively, and
\(\mathcal{B}\) contains learners for the conditional
cumulative hazard function of the censoring distribution.
Based on the Cartesian product of libraries of learners for
\((\Lambda_1,\Lambda_2,\Gamma)\) we construct a library
$\mathcal{F}$ of learners for \( F \):
\begin{align*}
  \mathcal{F}(\mathcal{A}_1, \mathcal{A}_2, \mathcal{B})
  &= \{ \phi_{a_1,a_2, b} : a_1 \in \mathcal{A}_1, a_2 \in \mathcal{A}_2, b \in \mathcal{B}\},
    \intertext{where in correspondence with  the relations in equation
    \eqref{eq:transition},}
    \phi_{a_1,a_2, b}(\data_n)(t,0,x)
  &= \exp{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid
    x)\} },
  \\
  \phi_{a_1,a_2, b}(\data_n)(t,1,x)
  &= \int_0^t
    \phi_{a_1,a_2, b}(\data_n)(s-,0,x)  a_1(\data_n)(\diff s \mid x),
  \\
  \phi_{a_1,a_2, b}(\data_n)(t,2,x)
  &= \int_0^t\phi_{a_1,a_2, b}(\data_n)(s-,0,x)  a_2(\data_n)(\diff s \mid x),
  \\
  \phi_{a_1,a_2, b}(\data_n)(t,-1,x)
  &= \int_0^t \phi_{a_1,a_2, b}(\data_n)(s-,0,x)  b(\data_n)(\diff s \mid x).
\end{align*}
Importantly, the libraries \( \mathcal{A}_1 \),
\( \mathcal{A}_2 \), and \( \mathcal{B} \) can be constructed
using standard survival analysis software.

To evaluate how well a function \( F \) predicts an observed
process $\eta$ we use the integrated Brier score
\( \bar B_\tau( F,O) = \int_0^{\tau} B_t(F,O) \diff t \),
where \( B_t \) is the Brier score
\citep{brier1950verification} at time \( t \in [0, \tau] \),
\begin{equation*}
  B_t(F,O) = \sum_{j=-1}^{2}
  \left(
      F(t,j,X) - \1{\{\eta(t)=j\}}
  \right)^2.
\end{equation*}
Based on a split of a data set \(\data_n\) into $K$ disjoint
approximately equally sized subsets (c.f., Section
\ref{sec:super-learning}), each learner
\( \phi_{a_1, a_2, b} \) in the library
\( \mathcal{F}(\mathcal{A}_1, \mathcal{A}_2, \mathcal{B}) \)
is evaluated using the cross-validated loss,
\begin{equation*}
  \hat{R}_{n}(\phi_{a_1,a_2,b} ; \bar{B}_{\tau}) =
  \frac{1}{K}\sum_{k=1}^{K}
  \frac{1}{| \data_n^{k} |}\sum_{O_i \in \data_n^{k}}
  \bar B_\tau
  {
    \left(
      \phi_{a_1,a_2,b}{ (\data_n^{-k})}
      , O_i
    \right)
  },
\end{equation*}
and the discrete state learner is given by
\begin{align*}\label{eq:discrete-state-learner}
  \hat{\phi}_n
  &=  \argmin_{(a_1,a_2,b)\in \mathcal{A}_1\times\mathcal{A}_2\times\mathcal{B}}
    \hat{R}_{n}(\phi_{a_1,a_2,b} ; \bar{B}_{\tau}).
\end{align*}

\section{Theoretical results for the state learner}
\label{sec:theor-results-prop}

In this section we establish theoretical guarantees for the
state learner. We show that the state learner is consistent if
its library contains a consistent learner, and we establish a
finite sample inequality for the excess risk of the state
learner compared to the oracle.

\subsection{Consistency}
\label{sec:consistency}

% We let \( \E_P \) denote
% expectation under \( P \) and use \( \Theta \) to denote the
% collection of all conditional state-occupation probability functions.
The following result can be derived from the fact
that the integrated Brier score (also called the continuous
ranked probability score) is a strictly proper scoring rule
\citep{gneiting2007strictly}. This implies that if we minimise
the average loss of the integrated Brier score, we recover the
parameters of the data-generating distribution. Specifically,
this implies that the oracle of a state learner is consistent
if the library of learners contains at least one learner that
is consistent for estimation of \( F \). Recall that the
function \(F\) implicitly depends on the data-generating
probability measure \(P\in\mathcal P\) but that this was
suppressed in the notation. We now make this dependence
explicit by writing \(F_P\) for the function determined by a
given \(P \in\mathcal{P}\) in accordance with equation
equation~(\ref{eq:F-def}). In the following we let
\( \mathcal{H}_{\mathcal{P}} = \{F_P : P \in \mathcal{P}\} \).

\begin{proposition}
  \label{prop:stric-prop}
  If \(P \in\mathcal{P}\) then
  \begin{equation*}
    F_P = \argmin_{F \in \mathcal{H}_{\mathcal{P}}} P{[\bar{B}_\tau(F, \blank)]},
  \end{equation*}
  \( P \)-almost surely.
\end{proposition}
\begin{proof}
  See Appendix~\ref{sec:consistency-proof}.
\end{proof}

\subsection{Oracle inequalities}
\label{sec:finite-sample-oracle}

We establish a finite sample oracle result for the state
learner. Our Proposition~\ref{prop:oracle-prop} is in essence a
special case of Theorem 2.3 in \citep{vaart2006oracle}. We
assume that we split the data into equally sized folds, and
for simplicity of presentation we take \( n \) to be such that
\( |\data_n^{-k}| = n/K \) with \( K \) fixed. We will allow
the number of learners to grow with \( n \) and write
\( \mathcal{F}_n=\mathcal{F}(\mathcal{A}_{1,n},
\mathcal{A}_{2,n}, \mathcal{B}_n)\) as short-hand notation and
to emphasise the dependence on \( n \). In the following we
let the space \( \mathcal{H}_{\mathcal{P}} \) be equipped with
the norm \( \| \blank \|_{P} \) defined as
\begin{equation}
  \label{eq:norm}
  \| F \|_{P} = 
  \left\{
    \sum_{j=-1}^{2}\int_{\mathcal{X}} \int_0^{\tau} F(t, j, x)^2 \diff t P( \diff x)
  \right\}^{1/2}.
\end{equation}

\begin{proposition}
  \label{prop:oracle-prop}
  For all \(P\in\mathcal{P}\), \( n \in \N \), \( k \in \{1, \dots, K\} \),
  and $\delta>0$,
  \begin{align*}
    \E_{P}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_P \Vert_{P}^2 \right]}
    & \leq (1 + 2\delta)
      \E_{P}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_P \Vert_{P}^2 \right]}
    \\
    & \quad
      + (1+ \delta) 16   K \tau
      \left(
      13 + \frac{12}{\delta}
      \right)
      \frac{\log(1 + |\mathcal{F}_n|)}{n}.
  \end{align*}
\end{proposition}
\begin{proof}
  See Appendix~\ref{sec:oracle-inequalities}.
\end{proof}

Proposition~\ref{prop:oracle-prop} has the following asymptotic consequences.

\begin{corollary}
  \label{cor:asymp-cons}
  Assume that \( |\mathcal{F}_n| = \bigO(n^q)\), for some \( q \in \N \) and
  that there exists a sequence \( \phi_n \in \mathcal{F}_n \), \( n \in \N \),
  such that
  \( \E_{P}{\left[ \Vert \phi_n(\data_n^{-k}) - F_P \Vert_{P}^2 \right]} =
  \bigO(n^{-\alpha}) \), for some \( \alpha\leq 1 \).
  \begin{enumerate}[label=(\alph*)]
  \item If $\alpha=1$ then
    \( \E_{P}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_P \Vert_{P}^2
      \right]} = \bigO(\log(n)n^{-1}) \).
  \item If $\alpha<1$ then
    \( \E_{P}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_P \Vert_{P}^2 \right]} =
    \bigO(n^{-\alpha}) \).
  \end{enumerate}
\end{corollary}
\begin{proof}
  See Appendix~\ref{sec:oracle-inequalities}.
\end{proof}

Proposition~\ref{prop:oracle-prop} provides a precise bound on
the average price we pay for using cross-validation to select
a learner from a library of candidates as compared to knowing
the oracle learner in advance. Corollary~\ref{cor:asymp-cons}
states that this price is asymptotically vanishing as long as
the number of learners in the library grows with sample size
at a polynomial rate.

\section{Use cases for the state learner}
\label{sec:use-cases-state}

The state learner estimates the parameter \( F \) which is a
feature of the observed right-censored data distribution
\( P \). In particular, \( F \) depends on the censoring
distribution and is typically not of direct interest in
itself. In the following, we describe how an estimate of
\( F \), as provided by the state learner, can be used to
estimate more relevant parameters of interest, such as, e.g.,
survival probabilities.

\subsection{Survival and risk predictions}
\label{sec:surv-risk-pred}

Event-free survival probabilities and risk predictions can be
obtained from the state-occupation function \( F \) under the
assumption of conditional independent censoring and
positivity, as introduced in Section~\ref{sec:framework}. By
equations~(\ref{eq:lambdaj}) and~(\ref{eq:gamma}) and the
definition of \( F \), we have
\begin{equation}
  \label{eq:7}
  % \Gamma(t , x) = \int_0^t \frac{F(\diff s, -1, x )}{F(s-,
  % 0, x )}, \quad \text{and} \quad
  \Lambda_j(t , x) 
  = \int_0^t  \frac{F(\diff s, j, x )}{F(s-, 0, x )},
  \quad j \in \{1,2\}.
\end{equation}
Equation~\eqref{eq:surv-def} provides a formula for obtaining
event-free survival probabilities from the cause-specific
cumulative hazard functions. Similarly, cause-specific risk
predictions can be obtained from \( \Lambda_1 \) and
$\Lambda_2$ using the formula
\citep[e.g.,][]{benichou1990estimates,
  ozenne2017riskregression},
\begin{equation}
  \label{eq:cs-risk-def}
  Q(T \leq t, D = j \mid X=x) =
  \int_0^t \exp\left\{-\Lambda_{1}(u \mid x)-\Lambda_{2}(u
    \mid x)\right\}  \Lambda_j(\diff u \mid x),
  \quad j \in \{1,2\}.
\end{equation}
Hence, given the state learner's estimate of \( F \) we can
use equation~\eqref{eq:7} to obtain estimates of the
cause-specific cumulative hazard functions $\Lambda_j$, which
can in turn be used to obtain estimates of the event-free
survival probabilites and cause-specific risks through
equations~\eqref{eq:surv-def} and~\eqref{eq:cs-risk-def}.

In Section~\ref{sec:super-learner-simple} we suggested to
implement the state learner by building a library using
learners of the cause-specific cumulative hazard functions,
$\Lambda_j$, and the cumulative hazard function for censoring,
$\Gamma$. With this implementation we can directly input the
highest ranked tuple of cause-specific hazard functions
$(\Lambda_1, \Lambda_2)$ provided by the state learner as
input to equations~\eqref{eq:surv-def}
and~\eqref{eq:cs-risk-def}.

\subsection{Targeted and debiased machine learning}
\label{sec:targeted-learning}

The framework of \textit{targeted} or \textit{debiased}
machine learning is a general methodology for combining
data-adaptive modeling, such as machine and super learning,
with asymptotically valid statistical inference for
interpretable target parameters \citep{van2006targeted,
  chernozhukov2018double}. \fxnote*{add more references?}{This
  framework has been applied extensively for causal inference
  based on observational data
  \citep[e.g.,][]{van2011targeted,kennedy2016semiparametric,rytgaard2022targeted}.}
A targeted or debiased estimator is constructed by first
estimation high-dimensional nuisance parameters using machine
learning, and then employing a debiasing or targeting step by
using tools from semi-parametric efficiency theory
\citep{pfanzagl1985contributions,bickel1993efficient,van2003unified,tsiatis2007semiparametric,kennedy2022semiparametric}.
In the context of right-censored competing risk data, the
high-dimensional nuisance parameters needed to be estimated
will typically include the cause-specific cumulative hazard
functions and the cumulative hazard function for censoring
\citep[e.g.,][]{van2003unified,rytgaard2021estimation,rytgaard2022targeted}.
The state learner is hence particularly well suited for
targeted and debiased machine learning as it immediately
provides estimates of these nuisance parameters. Though it
served as our original motivation for developing the state
learner, we leave the study of the state learner in the
context of targeted and debiased machine learning for a future
paper.

\section{Numerical experiments}
\label{sec:numer-exper}

In this section we report results from a simulation study where we consider
estimation of the conditional survival function. In the first part, we compare
the state learner to two IPCW based discrete super learners that use either the
Kaplan-Meier estimator or a Cox model to estimate the censoring probability
\citep{gonzalez2021stacked}. In the second part we compare the state learner to
the super learner proposed by \cite{westling2021inference}.

In both parts we use the same data-generating mechanism. We generate data
according to a distribution motivated from a real data set in which censoring
depends on the baseline covariates. We simulate data based on the prostate
cancer study of \cite{kattan2000pretreatment}. The outcome of interest is the
time to tumor recurrence, and five baseline covariates are used to predict
outcome: prostate-specific antigen (PSA, ng/mL), Gleason score sum (GSS, values
between 6 and 10), radiation dose (RD), hormone therapy (HT, yes/no) and
clinical stage (CS, six values). The study was designed such that a patient's
radiation dose depended on when the patient entered the study
\citep{gerds2013estimating}. This in turn implies that the time of censoring
depends on the radiation dose. The data were re-analysed in
\citep{gerds2013estimating} where a sensitivity analysis was conducted based on
simulated data. Here we use the same simulation setup, where event and censoring
times are generated according to parametric Cox-Weibull models estimated from
the original data, and the covariates are generated according to either marginal
Gaussian normal or binomial distributions estimated from the original data
\citep[c.f.,][Section~4.6]{gerds2013estimating}. We refer to this simulation
setting as `dependent censoring'. We also considered a simulation setting where
data were generated in the same way, except that censoring was generated
completely independently. We refer to this simulation setting as `independent
censoring'.

For all super learners we use a library consisting of three
learners: The Kaplan-Meier estimator
\citep{kaplan1958nonparametric,Gerds_2019prodlim}, a Cox model
with main effects \citep{cox1972regression, survival-package},
and a random survival forest
\citep{ishwaran2008random,rfsrc-paclage}. We use the same
library to learn the outcome distribution and the censoring
distribution. Note that the three learners in our library of
learners can be used to learn the cumulative hazard functions
of the outcome and the censoring distribution. The latter
works by training the learner on the data set \( \data_n^c \),
where \( \data_n^c = \{O_i^c\}_{i=1}^n \) with
\( O_i^c = (\tilde{T}_i, 1-\Delta_i, X_i) \). When we say that
we use a learner for the cumulative hazard function of the
outcome to learn the cumulative hazard function of the
censoring time, we mean that the learner is trained on
\( \data_n^c \).

We compare the state learner to two IPCW based super learners: The
first super learner, called IPCW(Cox), uses a Cox model with main
effects to estimate the censoring probabilities, while the second
super learner, called IPCW(KM), uses the Kaplan-Meier estimator to
estimate the censoring probabilities. The Cox model for the censoring
distribution is correctly specified in both simulation settings while
the Kaplan Meier estimator only estimates the censoring model
correctly in the simulation setting where censoring is
independent. Both IPCW super learners are fitted using the
\texttt{R}-package \texttt{riskRegression}
\citep{Gerds_Ohlendorff_Ozenne_2023}.
%
% | time | sim_setting | true_events | true_cens | at_risk |
% |------+-------------+-------------+-----------+---------|
% |   36 | original    |      24.619 |    61.853 |  25.774 |
% |   36 | indep_cens  |      24.674 |    38.740 |  46.141 |
%
The IPCW super learners use the integrated Brier score up to a fixed time
horizon (36 months). The marginal risk of the event before this time horizon is
\(\approx 24.6\)\%. Under the `dependent censoring' setting the marginal
censoring probability before the time horizon is \(\approx 61.9\)\%. Under the
`independent censoring' setting the marginal censoring probability before this
time horizon is \( \approx 38.7 \)\%.

Each super learner provides a learner for the cumulative
hazard function for the outcome of interest. From the
cumulative hazard function a risk prediction model can be
obtained as described in Section~\ref{sec:surv-risk-pred}
(with the special case of $\Lambda_2 = 0$). We measure the
performance of each super learner by calculating the index of
prediction accuracy (IPA) \citep{kattan2018index} at a fixed
time horizon (36 months) for the risk prediction model
provided by the super learner. The IPA is 1 minus the ratio
between the model's Brier score and the null model's Brier
score, where the null model is the model that does not use any
covariate information. The IPA is approximated using a large
(\( n = 20,000 \)) independent data set of uncensored data. As
a benchmark we calculate the performance of the risk
prediction model chosen by the oracle selector, which uses the
large data set of uncensored event times to select the learner
with the highest IPA.

The results are shown in Figure~\ref{fig:ipcw-fail}. We see that in
the scenario where censoring depends on the covariates, using the
Kaplan-Meier estimator to estimate the censoring probabilities
provides a risk prediction model with an IPA that is lower than the
risk prediction model provided by the state learner. The performance
of the risk prediction model selected by the state learner is similar
to the risk prediction model selected by the IPCW(Cox) super learner
which a priori uses a correctly specified model for the censoring
distribution. Both these risk prediction models are close to the
performance of the oracle, except for small sample sizes.

\begin{figure}
  \centering %
  \includegraphics[width=1\linewidth]{experiment-fig-sl-ipcw.pdf}
  \caption[]{For the risk prediction models provided by each of the super
    learners, the IPA is plotted against sample size. The results are averages across 
    1000 simulated data sets and the error bars are used to quantify the Monte Carlo
    uncertainty.
  }
  \label{fig:ipcw-fail}
\end{figure}

We next compare the state learner to the super learner \texttt{survSL}
\citep{westling2021inference}. This is another super learner which
like the state learner works without a pre-specified censoring
model. Note that both the state learner and \texttt{survSL} provide a
prediction model for the event time outcome and also for the
probability of being censored. Hence, we compare the performance of
these methods with respect to both the outcome and the censoring
distribution. Again we use the IPA to quantify the predictive
performance.

The results are shown in Figures~\ref{fig:zelefski-out}
and~\ref{fig:zelefski-cens}. We see that for most sample sizes, the state
learner selected prediction models for both censoring and outcome which have
similar or higher IPA compared to the prediction models selected by
\texttt{survSL}.
\begin{figure}
  \centering %
  \includegraphics[width=1\linewidth]{experiment-fig-sl-survSL-out.pdf}
  \caption[]{For the risk prediction models of the outcome provided by each
    of the super learners, the IPA at the fixed time horizon is plotted against
    sample size. The results are averages across 1000 repetitions and the error
    bars are used to quantify the Monte Carlo uncertainty.}
  \label{fig:zelefski-out}
\end{figure}

\begin{figure}
  \centering %
  \includegraphics[width=.7\linewidth]{experiment-fig-sl-survSL-cens.pdf}
  \caption[]{For the risk prediction models of the censoring model provided by
    each of the super learners, the IPA at the fixed time horizon is plotted
    against sample size. The results are averages across 1000 repetitions and
    the error bars are used to quantify the Monte Carlo uncertainty.}
  \label{fig:zelefski-cens}
\end{figure}

\section{Prostate cancer study}
\label{sec:real-data-appl}

In this section we use the prostate cancer data of
\cite{kattan2000pretreatment} to illustrate the use of the
state learner in the presence of competing risks. We have
introduced the data in Section~\ref{sec:numer-exper}. The data
consists of 1,042 patients who are followed from start of
followup until tumor recurrence, death without tumor
recurrence or end of followup (censored) whatever came first.
We use the state learner to rank libraries of learners for the
cause-specific cumulative hazard functions of tumor
recurrence, death without tumor recurrence, and censoring. The
libraries of learners each include five learners: the
Nelson-Aalen estimator, three Cox regression models
(unpenalized, Lasso, Elastic net) each including additive
effects of the 5 covariates (c.f., Section~\ref{sec:numer-exper}),
and a random survival forest. We use the same set of learners
to learn the cumulative hazard function of tumor recurrence
\( \Lambda_1 \), the cumulative hazard function of death
without tumor recurrence \( \Lambda_2 \), and the cumulative
hazard function of the conditional censoring distribution
$\Gamma$.

This gives a library consisting of \( 5^3 = 125 \) learners for the
conditional state occupation probability function \( F \) defined in
equation~(\ref{eq:F-def}). We use five folds for training and testing
the models, and we repeat training and evaluation five times with
different splits.  The integrated Brier score (defined in
Section~\ref{sec:super-learner-simple}) for all learners are shown in
Figure~\ref{fig:zelefski-real}. We see that the prediction
performance is mostly affected by the choice of learner for the
censoring distribution. Several combinations of learners give similar
performance as measured by the integrated Brier score, as long as a
random forest is used to model the censoring distribution.

\begin{figure}
  \centering %
  \includegraphics[width=1\linewidth]{real-data-state-learner.pdf}
  \caption[]{The results of applying the 125 combinations of learners to the
    prostate cancer data set. The error bars are based on five repetitions using
    different splits. We refer to learners of \( \Lambda_1 \), \( \Lambda_2 \),
    and $\Gamma$ as `Tumor learner', `Mortality learner', and `Censoring
    learner', respectively.}
  \label{fig:zelefski-real}
\end{figure}

\section{Discussion}
\label{sec:discussion}

The state learner is a new super learner that can be used with
right-censored data and competing events. Compared to existing
IPCW-based methods, the advantage of the state learner is that
it does not depend on a pre-specified estimator of the
censoring distribution, but selects one automatically based on
a library of candidate learners. Furthermore, the state
learner neither requires that the cause-specific cumulative
hazard functions \( \Lambda_j \) can be written as integrals
with respect to Lebesgue measure, nor does it assume a
(semi)parametric formula. In the remainder of this section
we discuss the limitations of our proposal and avenues for
further research.

A major advantage of the state learner is that the performance of each
combination of learners can be estimated without additional nuisance
parameters. A potential drawback of our approach is that we are
evaluating the loss of the learners on the level of the observed data
distribution while the target of the analysis is either the event time
distribution, or the censoring distribution, or both.
Specifically, the finite sample oracle inequality in
Proposition~\ref{prop:oracle-prop} concerns the function \( F \), which
is a feature of \( P \in \mathcal{P} \), while what we are typically
interested in is \( \Lambda_j \) or \( S \), which are features of
\( Q \in \mathcal{Q} \). We emphasise that while the state learner
provides us with estimates of \( \Lambda_j \) and $\Gamma$ based on
libraries \( \mathcal{A}_j \) and \( \mathcal{B} \), the performance
of these learners is not assessed directly for their respective target
parameters, but only indirectly via the performance of \( F \).  For
settings without competing risks, our numerical studies suggest that
measuring the performance of \( F \) also leads to good performance
for estimation of \( S \).

Our proposed super learner can be implemented with a broad library of learners
and using existing software.
% For instance the \texttt{R}-package
% \texttt{riskRegression} \citep{Gerds_Ohlendorff_Ozenne_2023} and the \texttt{R}-package by westling
Furthermore, while
the library \( \mathcal{F}(\mathcal{A}_1,\mathcal{A}_2,\mathcal{B}) \) consists
of \( |\mathcal{A}_1||\mathcal{A}_2||\mathcal{B}| \) many learners, we only need to fit
\( |\mathcal{A}_1| +|\mathcal{A}_2| + |\mathcal{B}| \) many learners in each fold. To
evaluate the performance of each learner we need to perform
\( |\mathcal{A}_1||\mathcal{A}_2||\mathcal{B}| \) many operations to calculate the
integrated Brier score in each hold-out sample, one for each combination of the
fitted models, but these operations are often negligible compared to fitting the
models. Hence the state learner is essentially not more computationally demanding
than any procedure that uses super learning to learn $\Lambda_1$, $\Lambda_2$,
and $\Gamma$ separately. While our proposal is based on constructing the library
\( \mathcal{F} \) from libraries for learning \( \Lambda_1 \), $\Lambda_2$, and
$\Gamma$, it could also be of interest to consider learners that estimate
\( F \) directly.

In our numerical studies, we only considered learners of $\Lambda_j$ and
$\Gamma$ that provide cumulative hazard functions which are piece-wise constant
in the time argument. This simplifies the calculation of \( F \) as the
integrals in equation~(\ref{eq:transition}) reduce to sums. When $\Lambda_j$ or
\( \Gamma \) are absolutely continuous in the time argument, calculating \( F \)
is more involved, but we expect that a good approximation can be achieved by
discretisation.

\section*{Conflict of interest}

The authors declare that they have no conflict of interest.

\bibliography{bib.bib}

\appendix

\section{Proofs}
\label{sec:proofs}

This section contains proofs of the results stated in the
paper. Section~\ref{sec:consistency-proof} contains a proof of
the consistency result from Section~\ref{sec:consistency}, and
Section~\ref{sec:oracle-inequalities} contains proofs of the
oracle inequalities from
Section~\ref{sec:finite-sample-oracle}.

\subsection{Consistency}
\label{sec:consistency-proof}

Define
\( \bar{B}_{\tau,P}(F, o) = \bar{B}_{\tau}(F, o) -
\bar{B}_{\tau}(F_P, o) \) and
\( R_{P}(F) = P{[\bar{B}_{\tau,P}(F, \blank)]} \), where the
integrated Brier score \( \bar{B}_{\tau} \) was defined in
Section~\ref{sec:super-learner-simple}. Recall the norm
\( \Vert \blank \Vert_{P}\) defined in
equation~(\ref{eq:norm}).

\begin{lemma}
  \label{lemma:norm}
  \( R_{P}(F) = \Vert F - F_P \Vert_{P}^2 \).
\end{lemma}
\begin{proof}[Proof]
  For any \( t \in [0, \tau] \) and \( j\in \{-1,0,1,2\} \) we have
  \begin{align*}
    & \E_{P}{\left[ (F(t, j, X) - \1{\{\eta(t) = j \}})^2 \right]}
    \\
    & =    \E_{P}{\left[ (F(t, j, X) - F_P(t, j, X) + F_P(t, j, X) - \1{\{\eta(t) = j
      \}})^2 \right]}
    \\
    & =    \E_{P}{\left[ (F(t, j, X) - F_P(t, j, X))^2\right]}
      + \E_{P}{\left[ (F_P(t, j, X) - \1{\{\eta(t) = j \}})^2\right]}
    \\
    & \quad
      + 2\E_{P}{\left[ (F(t, j, X) - F_P(t, j, X))(F_P(t, j, X) - \1{\{\eta(t) = j
      \}})\right]}
    \\
    & =    \E_{P}{\left[ (F(t, j, X) - F_P(t, j, X))^2\right]}
      + \E_{P}{\left[ (F_P(t, j, X) - \1{\{\eta(t) = j \}})^2\right]},
  \end{align*}
  where the last equality follows from the tower property. Hence, using Fubini,
  we have
  \begin{equation*}
    P{[\bar{B}_{\tau}(F, \blank)]}
    = \Vert F - F_P \Vert_{P}^2 + P{[\bar{B}_{\tau}(F_P, \blank)]}.
  \end{equation*}
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:stric-prop}]
  The result follows from Lemma~\ref{lemma:norm}.
\end{proof}

\subsection{Oracle inequalities}
\label{sec:oracle-inequalities}

Recall that we use \( \mathcal{F}_n \) to denote a library of learners for the
function \( F \), and that \( \hat{\phi} \) and \( \tilde{\phi} \) denotes,
respectively, the discrete super learner and the oracle learner for the library
\( \mathcal{F}_n \), c.f., Section~\ref{sec:super-learner-simple}.

\begin{proof}[Proof of Proposition~\ref{prop:oracle-prop}]
  First note that minimising the loss \( \bar{B}_{\tau} \) is equivalent to
  minimising the loss \( \bar{B}_{\tau,P} \), so the discrete super learner and
  oracle according to \( \bar{B}_{\tau} \) and \( \bar{B}_{\tau,P} \) are
  identical. By Lemma~\ref{lemma:norm}, \( R_P(F) \geq 0 \) for any
  \( F \in \mathcal{H}_{\mathcal{P}} \), and so using Theorem 2.3 from
  \citep{vaart2006oracle} with \( p=1 \), we have that for all \( \delta >0 \),
\begin{align*}
  & \E_{P}{\left[ R_P(\hat{\phi}_n(\data_n^{-k})) \right]}
  \\
  &  \quad \leq
    (1+2\delta)\E_{P}{\left[ R_P(\tilde{\phi}_n(\data_n^{-k})) \right]}
  \\
  & \qquad + (1+\delta) \frac{16 K}{n}
    \log(1 + |\mathcal{F}_n|)\sup_{F \in \mathcal{H}_{\mathcal{P}}}
    \left\{
    M(F) + \frac{v(F)}{R_P(F)}
    \left(
    \frac{1}{\delta} + 1
    \right)
    \right\}
\end{align*}
where for each \( F \in \mathcal{H}_{\mathcal{P}} \), \( (M(F), v(F)) \) is some Bernstein pair for
the function \(o \mapsto \bar{B}_{\tau,P}(F, o) \). As
\( \bar{B}_{\tau,P}(F, \blank) \) is uniformly bounded by \( \tau \) for any
\( F \in \mathcal{H}_{\mathcal{P}} \), it follows from section 8.1 in \citep{vaart2006oracle} that
\( (\tau, 1.5 P{[\bar{B}_{\tau,P}(F, \blank)^2]}) \) is a Bernstein pair for
\( \bar{B}_{\tau,P}(F, \blank) \). Now, for any \( a,b,c \in \R \) we have
\begin{align*}
  (a-c)^2 - (b-c)^2
  & = (a-b+b-c)^2 - (b-c)^2
  \\
  & = (a-b)^2 + (b-c)^2 +2(b-c)(a-b) - (b-c)^2
  \\
  & = (a-b)
    \left\{
    (a-b) +  2(b-c)
    \right\}
  \\
  & = (a-b)
    \left\{
     a + b -2c
    \right\},
\end{align*}
so using this with \( a=F(t, j, x) \), \( b=F_P(t, j, x) \), and
\( c = \1{\{\eta(t) = j\}} \), we have by Jensen's inequality
\begin{align*}
  & P{[\bar{B}_{\tau,P}(F, \blank)^2]}
  \\
  & \leq
    2\tau\E_{P}{\left[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left\{
    \left(
    F(t, j, X) - \1{\{\eta(t) = j\}}
    \right)^2
    -
    \left(
    F_P(t, j, X) - \1{\{\eta(t) = j\}}
    \right)^2
    \right\}^2
    \diff t 
    \right]}
  \\
  & =2\tau
    \E_{P}\Bigg[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left(
    F(t, j, X) - F_P(t, j, X)
    \right)^2
  \\
  & \quad \quad \quad\quad \quad \quad \times
    \left\{
    F(t, j, X) +  F_P(t, j, X)-2 \1{\{\eta(t) = j\}}
    \right\}^2
    \diff t 
    \Bigg]
  \\
  & \leq
    8\tau \E_{P}{\left[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left(
    F(t, j, X) - F_P(t, j, X)
    \right)^2
    \diff t 
    \right]}.
  \\
  & =
    8\tau \Vert F - F_P \Vert_{P}^2.
\end{align*}
Thus when \( v(F) = 1.5 P{[\bar{B}_{\tau,P}(F, \blank)^2]} \) we have by
Lemma~\ref{lemma:norm}
\begin{equation*}
  \frac{v(F)}{R_P(F)}
  = 1.5 \frac{P{[\bar{B}_{\tau,P}(F, \blank)^2]}}{P{[\bar{B}_{\tau,P}(F, \blank)]}}
  \leq 12 \tau,
\end{equation*}
and so using the Bernstein pairs \( (\tau, 1.5 P{[\bar{B}_{\tau,P}(F, \blank)^2]}) \) we have
\begin{equation*}
  \sup_{F \in \mathcal{H}_{\mathcal{P}}}
  \left\{
    M(F) + \frac{v(F)}{R_P(F)}
    \left(
      \frac{1}{\delta} + 1
    \right)
  \right\}
  \leq \tau
  \left(
    13 + \frac{12}{\delta}
  \right).
\end{equation*}
For all $\delta>0$ we thus have
\begin{align*}
  \E_{P}{\left[ R_P(\hat{\phi}_n(\data_n^{-k})) \right]}
  \leq
  &(1+2\delta)\E_{P}{\left[ R_P(\tilde{\phi}_n(\data_n^{-k})) \right]}
  \\
  & \quad
    + (1+\delta)\log(1 + |\mathcal{F}_n|) \tau \frac{16 K}{n}
    \left(
    13 + \frac{12}{\delta}
    \right),
\end{align*}
and then the final result follows from Lemma~\ref{lemma:norm}.
\end{proof}

\begin{proof}[Proof of Corollary~\ref{cor:asymp-cons}]
  By definition of the oracle and Lemma~\ref{lemma:norm},
  \begin{equation*}
    \E_{P}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_P \Vert_{P}^2
      \right]} \leq \E_{P}{\left[ \Vert \phi_n(\data_n^{-k}) - F_P \Vert_{P}^2
      \right]}  
  \end{equation*}
  for all \( n \in \N \). The results then follows from
  Proposition~\ref{prop:oracle-prop}.
\end{proof}

\end{document}
% end of file template.tex